---
title: "Untitled"
output: html_document
date: '2022-09-27'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
load("seuratData.37.RData")

s0=seuratData[,seuratData$celltype=="neutrophil"]

for(i in seq(5,50,by=5)){
  print(i)
  s=s0
  s <- FindVariableFeatures(s, selection.method = "vst", nfeatures = 5000)
  s <- ScaleData(object = s, vars.to.regress = c("S.Score", "G2M.Score",'batcheffect','percent.mt','percent.ercc','nFeature_RNA'))
  
  #Dimension reduction
  s <- RunPCA(s, npcs = 50, verbose = FALSE,features =VariableFeatures(s) )
  dims=i
  s <- RunUMAP(s, dims = 1:dims, verbose = FALSE,perplexity=20)

  s <- FindNeighbors(s,reduction = 'pca', verbose = FALSE, dims = 1:dims)
  s <- FindClusters(s, algorithm = 1, random.seed = 256, resolution = 0.3)
  p1=DimPlot(s,label = T)+scale_color_brewer(palette = 'Set1')
  p2=DimPlot(s,group.by = "orig")+scale_color_brewer(palette = 'Set1')
  p3=DimPlot(s,group.by = "celltype2")
  pdf(paste0("neutrophil",i,".pdf"))
  print(p1)
  print(p2)
  print(p3)
  dev.off()
  
}



##rm GF sample
```{r}
dataname0="20220927"
load("F:/Edata/yaobingqing/20220922_GF/4sample/20220922seuratdata.RData")
seuratData=seuratData[,!seuratData$orig%in%"GF"]
seuratData$orig=droplevels(seuratData$orig)

seuratData=seuratData[,!seuratData$celltype2%in%"other"]

seuratData$celltype=plyr::mapvalues(seuratData$celltype,from = "epithelial",to="macrophage")
seuratData$celltype2=plyr::mapvalues(seuratData$celltype2,from = "epithelial",to="alveolar_mac")
seuratData$celltype2=plyr::mapvalues(seuratData$celltype2,from = "macrophage",to="tissue_mac")

```

```{r}
s.genes=first2big(cc.genes.updated.2019$s.genes)
g2m.genes=first2big(cc.genes.updated.2019$g2m.genes)

seuratData <- CellCycleScoring(seuratData, s.features = s.genes, g2m.features =g2m.genes, set.ident = TRUE)
seuratData=NormalizeData(object = seuratData, normalization.method = "LogNormalize", scale.factor = 10000)
seuratData <- FindVariableFeatures(seuratData, selection.method = "vst", nfeatures = 5000)
seuratData <- ScaleData(object = seuratData, vars.to.regress = c("S.Score", "G2M.Score",'batcheffect','percent.mt','percent.ercc','nFeature_RNA'))


#Dimension reduction
seuratData <- RunPCA(seuratData, npcs = 50, verbose = FALSE,features =VariableFeatures(seuratData) )
 # ElbowPlot(seuratData, ndims = 70)
# 
# seuratData=JackStraw(seuratData, num.replicate = 100,dims = 100)
# seuratData <- ScoreJackStraw(seuratData, dims = 1:70)
# JackStrawPlot(seuratData, dims = 1:50)


# dims=which(seuratData@reductions$pca@jackstraw$overall.p.values[,2]>0.05)[1]-1
# if(dims>20){dims=20}
dims=30
if(ncol(seuratData)<70){
  seuratData <- RunTSNE(seuratData, dims = 1:dims, verbose = FALSE,perplexity=20)
}else{
  seuratData <- RunTSNE(seuratData, dims = 1:dims, verbose = FALSE)
}
seuratData <- FindNeighbors(seuratData,reduction = 'pca', verbose = FALSE, dims = 1:dims)
seuratData <- FindClusters(seuratData, algorithm = 1, random.seed = 256, resolution = 1)
pdf(paste0(dataname0,"cluster.pdf"))
DimPlot(seuratData,label = T)
DimPlot(seuratData,group.by = "orig")
DimPlot(seuratData,group.by = "celltype",label = T)
DimPlot(seuratData,group.by = "celltype2",label = T)

dev.off()

```

```{r}
seuratData  <- RunUMAP(seuratData , dims = 1:50)

# ElbowPlot(object = seuratData, ndims = 100)
seuratData  <- FindNeighbors(object = seuratData , dims = 1:50, verbose = FALSE)
seuratData  <- FindClusters(object = seuratData , resolution = 1, verbose = FALSE)

p5=DimPlot(seuratData,label = T)+scale_color_manual(values = colpalettes)
p7=DimPlot(seuratData,group.by = "orig",label = T)+scale_color_manual(values = colpalettes)
p8=DimPlot(seuratData,group.by = "celltype",label = T)+scale_color_manual(values = colpalettes)
p8.2=DimPlot(seuratData,group.by = "celltype2",label = T)+scale_color_manual(values = colpalettes)
pdf(paste0(dataname0,"cluster2-umap",".pdf"))
print(p5)
print(p7)
print(p8)
p8.2
dev.off()

```

```{r}
genes_to_check = c('Esr1','Pgr','Bcl11b','Tspan8','Tspan18','Tspan1','Procr','Axin2','Krt8','Krt18','Krt19','Krt5','Krt14','Krt17','Lgr5','Dll1','Elf5','Cd14','Cd200','Cd34','Aldh1a1','Sox9','Mycn','Tbx2','Sema3a','Mki67','Top2a','Ccnd1','Ccnd2','Ccnd3','Hes1','Gli1','Gli2','Gli3','Hey1','Spp1','Cd47','Slc6a2','Bcl11b','Pecam1','Ptprc','Col1a1','Itgb3','Itga2','Ly6a','Trpm3','Brca1','Lepr','Ptger3','Tmem97','Cd14','Prom1', "Erbb2","Esr1","Epcam","Cd274","Ar","Cdh1","Cd3e","Cd68","Cd8a","Cd4",'Krt14','Krt5',"Dpp4","Dlk1","Tgfb1","Tgfb2","Ifng","Ctla4","S100a4","Tnf","Itgam","Itgax","Pecam1","Acta2","Lyve1","Cd19","Cd78","FoxP3","Cdh5","Ly6g","Ly6c",'Klrk1', 'Pdpn','Siglecf', 'Cd24',"Cd24a", 'Fcgr1a',"Cd64a","Cd64b","Cd64c","H2-Q1",  "H2-Q2",  "H2-Q4" , "H2-Q6" , "H2-Q7" , "H2-Q10",'Cd200','Krt14','Krt8','Epcam','Acta2','Pdpn','Emb')

genes_to_check=first2big(genes_to_check)
names(genes_to_check)=NULL
genes_to_check=sort(intersect(genes_to_check,rownames(seuratData)))
# VlnPlot(object = seuratData, features= genes_to_check,pt.size = 0.2)

jpeg(paste0(dataname0,"all_marker","_Vln.jpg"),width = 4000,height = 1600)
b=VlnPlot(object = seuratData,genes_to_check,pt.size = 0, ncol = 10)+NoLegend()
print(b)
dev.off()
jpeg(paste0(dataname0,"all_marker","_fea.jpg"),width = 2000,height = 1600)
b=FeaturePlot(object = seuratData,genes_to_check,pt.size = 0.3, ncol = 10)+NoLegend()
print(b)
dev.off()


epithelial_marker=c("EPCAM","KRT8","KRT18","KRT19","KRT14","KRT17","KRT5","ERBB2","ESR1","PGR")
lung_epithelial=c("EPCAM","CDH1","AGER","HOPX","SFTPC","SLC34A2","ABCA3","SOX2","PAX9","TP63","KRT5","KRT14","MUC5B","SCGB1A1")
fibroblast_marker=c("ACTA2","PDGFRA","DCN","THY1","COL1A1","FAP","PDPN","COL1A2","COL3A1","COL6A1")
endo_marker=c("PECAM1","CLDN5","FLT1","RAMP2","VWF","ENG","CDH5","KDR","RNASE1","LYVE1","CD34")
Bcell_marker=c("CD79A","IGHM","IGHG3","IGHA2","CD19")
Tcell_marker=c("CD3D","CD3E","CD3G","TRAC","CD2","CD8B","CD8A")
NK_marker=c("XCL1","KLRD1","KLRF1","FCGR3A","KLRB1","NCAM1","NKG7","GNLY" )
macrophage_marker=c("CD14","CD163","CD68","FCGR2A","CSF1R")
DC_marker=c("CD1A","CD1B","FCER1A","CD40","CD80","CD83","CCR7","LAMP3")
neutrophil_marker=c("S100a8","S100a9","CD14","CXCR2","ITGA4","ITGAM","ITGAX","ITGB2","SIGLEC1","SIGLECF")

myeloid_marker=c("LYZ","AIF1","C1QB","CD14","CD163","CD68","FCGR2A","CSF1R","CD1A","CD1B","FCER1A","CD80","LAMP3","S100a8","S100a9","ITGA4","ITGAM","ITGAX","ITGB2")

celltypemarker=unique(c(epithelial_marker,lung_epithelial,fibroblast_marker,endo_marker,Bcell_marker,Tcell_marker,NK_marker,macrophage_marker,DC_marker,myeloid_marker))

celltypemarker=first2big(celltypemarker)
jpeg(paste0(dataname0,"celltype_marker","_fea.jpg"),width = 2000,height = 1000)
b=FeaturePlot(object = seuratData,sort(celltypemarker),pt.size = 0.3, ncol = 10)+NoLegend()
print(b)
dev.off()
pdf("marker.pdf",width = 10)
DotPlot(object=seuratData,assay ="RNA",features = c(celltypemarker))+
  theme(axis.text.x = element_text(size=5,vjust=0.5,angle = 90))
DotPlot(object=seuratData,assay ="RNA",group.by = "celltype",features = c(celltypemarker))+
  theme(axis.text.x = element_text(size=5,vjust=0.5,angle = 90))
dev.off()

celltypemarker=list(nuetrophil=c("S100a8","S100a9","CD14","CXCR2","ITGAM"),
                 macrophage=c("ITGAX","SIGLEC1","SIGLECF",
                 "CD68","CSF1R"),
                 DC=c("CD40","CD80","CD83","CCR7"),#,"LAMP3"
                 Tcell=c("CD3D","CD3E","CD3G","TRAC","CD2"),
                 Bcell=c("CD79A","IGHM","CD19"),
                 NK=c("KLRD1","NKG7"),
                 endo=c("PECAM1","CLDN5","FLT1","RAMP2","ENG","CDH5","KDR","LYVE1","CD34"),
                 fibroblast=c("PDGFRA","DCN","PDPN","COL1A1","COL1A2","COL3A1","COL6A1"),
                  epithelial=c("EPCAM","KRT19")
                 )
celltypemarker=do.call("c",celltypemarker)
names(celltypemarker)=NULL

celltypemarker=first2big(celltypemarker)
celltypemarker=celltypemarker[!duplicated(celltypemarker)]
pdf("marker2.pdf",width = 10)
DotPlot(object=seuratData,assay ="RNA",group.by = "celltype",features = c(celltypemarker))+
  theme(axis.text.x = element_text(size=5,vjust=0.5,angle = 90))
dev.off()

save(seuratData,file = "seuratData.RData")
```



# subcluster
```{r}
markerlist=list()
seuratData_blood10x_extract=list()
for(celltype in c("neutrophil","macrophage","DC","Tcell","Bcell","NK")){#
  print(celltype)
  s=seuratData[,seuratData$celltype==celltype]
  s <- FindVariableFeatures(s, selection.method = "vst", nfeatures = 5000)
  s <- ScaleData(object = s, vars.to.regress = c("S.Score", "G2M.Score",'batcheffect','percent.mt','percent.ercc','nFeature_RNA'))
  
  #Dimension reduction
  s <- RunPCA(s, npcs = 50, verbose = FALSE,features =VariableFeatures(s) )
  dims=10
  if(ncol(s)<70){
    s <- RunTSNE(s, dims = 1:dims, verbose = FALSE,perplexity=20)
  }else{
    s <- RunTSNE(s, dims = 1:dims, verbose = FALSE)
  }
  s <- FindNeighbors(s,reduction = 'pca', verbose = FALSE, dims = 1:dims)
  s <- FindClusters(s, algorithm = 1, random.seed = 256, resolution = 0.3)
  p1=DimPlot(s,label = T)
  p2=DimPlot(s,group.by = "orig")
  p3=DimPlot(s,group.by = "celltype2")
  
  # t=c("KLRF1","KLRD1","CD4","CD3D","CD3G","CD3E","CD8A","CD8B")
  g=first2big(c(do.call("c",celltype_marker_list[[celltype]]),celltypemarker))
  g=g[!duplicated(g)]
  names(g)=NULL
  p4=DotPlot(s,features = g)+coord_flip()
  p4.1=DotPlot(s,features = g,group.by = "celltype2")+coord_flip()
  
  aveexp=AverageExpression(s,features = g,return.seurat = T,group.by = "seurat_clusters")
  aveexp$celltype=rownames(aveexp@meta.data)
  Idents(aveexp)="celltype"
  p5=DoHeatmap(aveexp,features = g,draw.lines = F)+NoLegend()+
  scale_fill_gradient2(low = scales::muted("blue"),high = scales::muted("red"))+
  theme(axis.text.y = element_text(size=5))

  pdf(paste0(dataname0,"_",celltype,".pdf"))
  print(p1)
  print(p2)
  print(p3)
  print(p4)
  print(p4.1)
  print(p5)
  dev.off()
  
  seuratData_blood10x_extract[[celltype]]=s
  if(length(unique(s$seurat_clusters))==1){next()}
  marker=FindAllMarkers(s,only.pos = T)
  t=do.call("c",lapply(unique(sort(as.numeric(as.character(s$seurat_clusters)))),function(x){head(marker$gene[marker$cluster==x],10)}))
  t=t[!duplicated(t)]
  p3=FeaturePlot(s,features = c(t),ncol = 10)
  tiff(paste0(dataname0,celltype,".markergene.tiff"),res = 300,width = 10000,height = 1000*length(unique(s$seurat_clusters)))
  print(p3)
  dev.off()
  markerlist[[celltype]]=marker
  
  
}

save(seuratData_blood10x_extract,markerlist,file="seuratData_blood10x_extract.markerlist.RData")
```

```{r}
seuratData$cd45=seuratData$celltype
seuratData$cd45=plyr::mapvalues(seuratData$celltype,from = c("neutrophil","macrophage","DC","Tcell","Bcell","NK"),to=rep("cd45",6))
seuratData$ly_my=as.character(seuratData$celltype)
seuratData$ly_my[seuratData$celltype2%in%do.call("c",lapply(celltype_marker_list[c("Tcell","Bcell","NK")],names))]="lymphocyte"
seuratData$ly_my[seuratData$celltype2%in%do.call("c",lapply(celltype_marker_list[c("neutrophil","macrophage","DC","myeloid")],names))]="myeloid"

for(i in c("celltype","celltype2","cd45","ly_my")){
  print(i)
  
  data=data.frame(table(seuratData@meta.data[,c(i,"orig")]))
  colnames(data)[1]="celltype"
  colnames(data)[2]="orig"
  data=data%>%group_by(orig)%>%mutate(combined_freq=Freq/sum(Freq))
  data=data%>%group_by(celltype)%>%mutate(combined_freq2=Freq/sum(Freq))
  data=data%>%arrange(desc(celltype))%>%group_by(orig)%>%mutate(ypos = cumsum(combined_freq)- 0.5*combined_freq )
  
  p1=ggplot(data,aes(orig,combined_freq,fill=celltype))+geom_bar(stat="identity")+theme_classic()+
    theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0))+scale_fill_manual(values = colpalettes)
  p2=ggplot(data,aes(orig,combined_freq,fill=celltype))+geom_bar(stat="identity")+
    facet_wrap(~celltype,scales = "free_y")+theme_classic()+
    theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0))+scale_fill_manual(values = colpalettes)
  pdf(paste0(dataname0,"_blood_cellproportion_",i,".pdf"))
  print(p1)
  print(p2)
  dev.off()
}

data0=seuratData@meta.data
datalist=list(lymphocyte=data0[data0$ly_my%in%"lymphocyte",],myeloid=data0[data0$ly_my%in%"myeloid",],cd45=data0[data0$cd45=="cd45",])
for(i in names(datalist)){
  print(i)
  data=datalist[[i]]
  data=data.frame(table(data[,c("celltype","orig")]))
  colnames(data)[1]="celltype"
  colnames(data)[2]="orig"
  data=data[data$Freq>0,]
  data=data%>%group_by(orig)%>%mutate(combined_freq=Freq/sum(Freq))
  data=data%>%group_by(celltype)%>%mutate(combined_freq2=Freq/sum(Freq))
  data=data%>%arrange(desc(celltype))%>%group_by(orig)%>%mutate(ypos = cumsum(combined_freq)- 0.5*combined_freq )
  p1=ggplot(data,aes(orig,combined_freq,fill=celltype))+geom_bar(stat="identity")+
    theme_classic()+theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0))+scale_fill_manual(values = colpalettes)
  p2=ggplot(data,aes(orig,combined_freq,fill=celltype))+geom_bar(stat="identity")+
    facet_wrap(~celltype,scales = "free_y")+theme_classic()+theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0))+
    scale_fill_manual(values = colpalettes)
  
  p3=ggplot(data, aes(x="", y=combined_freq, fill=celltype)) +
    geom_bar(stat="identity", width=1, color="white") +
    coord_polar("y", start=0) +
    theme_void() + facet_wrap(~orig)+
    theme(legend.position="none") +
    geom_text(aes(y=ypos,label = celltype), color = "black", size=2) +scale_fill_manual(values = colpalettes)
  p4=ggplot(data, aes(x="", y=combined_freq, fill=celltype)) +
    geom_bar(stat="identity", width=1, color="white") +
    coord_polar("y", start=0) +
    theme_void() + facet_wrap(~orig)+
    theme(legend.position="none") +
    geom_text(aes(y=ypos,label = signif(combined_freq,2)), color = "black", size=2)  +scale_fill_manual(values = colpalettes)
  
  pdf(paste0("20230304-blood-",i,".pdf"))
  print(p1)
  print(p2)
  print(p3)
  print(p4)
  dev.off()
}


```


```{r}
# neutrophil_genelist=list()
# 
# N2_vs_N1=xlsx::read.xlsx("E:/YaoBingQing/20210529_blood_singlecell/celltype_recluster/publicdata/1/koni-05-11-1232221-s001/KONI_A_1232221_s04.xlsx",sheetIndex = 1,startRow = 6)
# N2vsN1=data.frame(gene=c(N2_vs_N1$Gene,N2_vs_N1$Gene.1,N2_vs_N1$Gene.2),N2VSN1=c(N2_vs_N1$Fold.change.N2.vs..N1,N2_vs_N1$Fold.change.N2.vs..N1.1,N2_vs_N1$Fold.change.N2.vs..N1.2))
# N2vsN1$N1N2="N1"
# N2vsN1$N1N2[N2vsN1$N2VSN1>0]="N2"
# N2vsN1=N2vsN1[,c(3,1,2)]
# N1genes=N2vsN1$gene[N2vsN1$N1N2=="N1"]
# neutrophil_genelist[["N1_staph_control"]]=list(N1gene=N1genes)
# 
# 
# f="E:/YaoBingQing/20210529_blood_singlecell/celltype_recluster/publicdata/2/NN_gene.xlsx"
# N1N2gene=list(GMDSC_NN=xlsx::read.xlsx(f,sheetName = "GMDSC_NN"),
#               TAN_NN=xlsx::read.xlsx(f,sheetName = "TAN_NN"),
#               TAN_GMDSC=xlsx::read.xlsx(f,sheetName = "TAN_GMDSC"))
# N1N2gene=lapply(N1N2gene,function(x){x[-1,]})
# plosone_N1N2=sapply(N1N2gene,function(x){x$Gene.Symbol})
# neutrophil_genelist[["plosone"]]=plosone_N1N2
# 
# datasig=read.table("E:/YaoBingQing/20210529_blood_singlecell/celltype_recluster/sig_gene.txt",header = T,as.is = T)
# N1N2gene=list(N1=rownames(datasig)[datasig$N1_N.log2FoldChange>1&datasig$N1_N.padj<0.05&datasig$N1_N2.log2FoldChange>1&datasig$N1_N2.padj<0.05],
#               N2=rownames(datasig)[datasig$N2_N.log2FoldChange>1&datasig$N2_N.padj<0.05&datasig$N1_N2.log2FoldChange< -1&datasig$N1_N2.padj<0.05])
# N1N2gene=lapply(N1N2gene,function(x){x[!is.na(x)]})
# neutrophil_genelist[["frontiers_in_immunology"]]=N1N2gene
# 
# GMDSC=xlsx::read.xlsx("E:/YaoBingQing/20210529_blood_singlecell/celltype_recluster/publicdata/5_MDSC/GMDSC.xlsx",sheetIndex = 1)
# GMDSC=GMDSC[GMDSC$avg_logFC>0.25,]
# MMDSC=xlsx::read.xlsx("E:/YaoBingQing/20210529_blood_singlecell/celltype_recluster/publicdata/5_MDSC/MMDSC.xlsx",sheetIndex = 1)
# MMDSC=MMDSC[MMDSC$avg_logFC>0.25,]
# MDSC=list(GMDSC=GMDSC$gene,MMDSC=MMDSC$gene)
# neutrophil_genelist[["MDSC"]]=MDSC
# 
# N2=xlsx::read.xlsx("E:/YaoBingQing/20210529_blood_singlecell/celltype_recluster/publicdata/3combine_multi_paper/N2_gene_combined.xlsx",sheetIndex = 1,header = F)
# neutrophil_genelist[["N2_comb"]]=list(N2_comb=sapply(N2$X1,first2big))
save(neutrophil_genelist,file = "neutrophil_genelist.RData")
load("X:/analyze_data/FEdata/yaobingqing/neutrophil_genelist.RData")

s=seuratData_blood10x_extract$neutrophil


s=s[,s$celltype2=="neutrophil"]

alllist=do.call("c",neutrophil_genelist)
s=AddModuleScore(s,features = alllist,seed = 100)
colnames(s@meta.data)[grep("Cluster1",colnames(s@meta.data)):ncol(s@meta.data)]=names(alllist)
pdf("neutrophil_score.pdf")
for(genelist in names(alllist)){
  print(genelist)
  d=s@meta.data[,c("orig",genelist)]
  colnames(d)[ncol(d)]="exp"
  p2.1=ggplot(d,aes(orig,exp,col=orig))+geom_violin()+geom_boxplot(width=.1,outlier.colour = NA)+
    ggpubr::stat_compare_means(comparisons = as.list(data.frame(combn(levels(d$orig),2))))+theme_classic()+
    labs(title = genelist)+geom_hline(yintercept = median(d$exp[d$orig=="GF"]))
    

  print(p2.1)
}
dev.off()



### macrophage score
s=seuratData_blood10x_extract$macrophage

alllist=do.call("c",neutrophil_genelist)
s=AddModuleScore(s,features = alllist,seed = 100)
colnames(s@meta.data)[grep("Cluster1",colnames(s@meta.data)):ncol(s@meta.data)]=names(alllist)
pdf("macrophage_score.pdf")
for(genelist in names(alllist)){
  print(genelist)
  d=s@meta.data[,c("orig","celltype2",genelist)]
  colnames(d)[ncol(d)]="exp"
  p2.1=ggplot(d,aes(orig,exp,col=orig))+geom_violin()+geom_boxplot(width=.1,outlier.colour = NA)+
    ggpubr::stat_compare_means(comparisons = as.list(data.frame(combn(levels(d$orig),2))))+theme_classic()+
    labs(title = genelist)+geom_hline(yintercept = median(d$exp[d$orig=="GF"]))+facet_wrap(~celltype2)
  
  
  print(p2.1)
}
dev.off()


gsea_stage_list2=list()
for(celltype in c(names(seuratData_blood10x_extract))){
  for(celltype2 in unique(seuratData_blood10x_extract[[celltype]]$celltype2)){
    print(celltype)
    s=seuratData_blood10x_extract[[celltype]]
    s=s[,s$celltype2%in%celltype2]
    Idents(s)="orig"
    pairstocomp=list(control_vs_WT=c("control","WT"),  control_vs_ATBX=c("control","ATBX"), ATBX_vs_WT=c("ATBX","WT"))
    for(i in names(pairstocomp)){
      print(paste(celltype,celltype2,i))
      id1=pairstocomp[[i]][1]
      id2=pairstocomp[[i]][2]
      
      genes <- FindMarkers(object = s, ident.1 =id1,ident.2 = id2, only.pos = F, min.pct = 0.01, logfc.threshold = 0,test.use = 'wilcox')
      genes=genes[order(genes$avg_log2FC,decreasing = T),]
      geneList=genes$avg_log2FC
      names(geneList)=rownames(genes)
      gsea <- clusterProfiler::GSEA(geneList, TERM2GENE = pathwaylist)
      gsea_stage_list2[[celltype]][[celltype2]][[i]]=gsea
    }
  }  
}

neutrophil_gsealist=gsea_stage_list2$neutrophil$neutrophil
writexl::write_xlsx(lapply(neutrophil_gsealist,function(x){x@result}),path = "neutrophil_gsea.xlsx")
neutrophil_gsealist_df=lapply(neutrophil_gsealist,function(x){x@result})
save(neutrophil_gsealist,file = "neutrophil_gsealist.RData")

gs=list(Natural_killer_cell_mediated_cytotoxicity=mmu_kegg_list_des$`Natural killer cell mediated cytotoxicity`)
s=seuratData_blood10x_extract$NK
s=AddModuleScore(s,features = gs,seed = 100)
colnames(s@meta.data)[grep("Cluster1",colnames(s@meta.data)):ncol(s@meta.data)]=names(gs)
pdf("NK_score.pdf")
for(genelist in names(gs)){
  print(genelist)
  d=s@meta.data[,c("orig",genelist)]
  colnames(d)[ncol(d)]="exp"
  p2.1=ggplot(d,aes(orig,exp,col=orig))+geom_violin()+geom_boxplot(width=.1,outlier.colour = NA)+
    ggpubr::stat_compare_means(comparisons = as.list(data.frame(combn(levels(d$orig),2))))+theme_classic()+
    labs(title = genelist)+geom_hline(yintercept = median(d$exp[d$orig=="control"]))
  
  
  print(p2.1)
}
genes=c(celltype_marker_list$Tcell$Tcytotoxic,celltype_marker_list$NK$NK)
genes=genes[grep("KL",genes,invert = T)]
genes=first2big(genes)
genes=genes[!duplicated(genes)]
VlnPlot(seuratData_blood10x_extract$NK,features = c(sort(genes)),group.by = "orig",pt.size = 0)
d=seuratData_blood10x_extract$NK@assays$RNA@data[match(genes,rownames(seuratData_blood10x_extract$NK@assays$RNA@data),nomatch = 0),]

d=data.frame(t(d),orig=seuratData_blood10x_extract$NK$orig)
df=reshape2::melt(d)
ggplot(df,aes(orig,value,col=orig))+geom_violin()+geom_boxplot(width=.1,outlier.colour = NA)+
  facet_wrap(~variable)+
  ggpubr::stat_compare_means(comparisons = as.list(data.frame(combn(levels(d$orig),2))))+
  theme_classic()

dev.off()

```


#diff gene & pathway
```{r}
library(clusterProfiler)
library(org.Mm.eg.db)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(ggplot2)
k=keys(org.Mm.eg.db,keytype = "SYMBOL")
list=select(org.Mm.eg.db,keys=k,columns = c("ENTREZID","SYMBOL"), keytype="SYMBOL")


markerlist=list()
enrichfunciton=list()
for(celltype in c("neutrophil","macrophage","DC","Tcell","Bcell","NK")){
  print(celltype)
  s0=seuratData_blood10x_extract[[celltype]]
  Idents(s0)="orig"
  pairstocomp=list(control_vs_WT=c("control","WT"),  control_vs_ATBX=c("control","ATBX"), ATBX_vs_WT=c("ATBX","WT"))
  for(celltype2 in unique(s0$celltype2)){
    s=s0[,s0$celltype2%in%celltype2]
    for(i in names(pairstocomp)){
    print(i)
    id1=pairstocomp[[i]][1]
    id2=pairstocomp[[i]][2]
    f=FindMarkers(s,ident.1 = id1,ident.2 = id2)
    f=f[f$p_val<0.05,]
    markerlist[[celltype]][[celltype2]][[i]]=f
    }
  }
  }


for(celltype in names(markerlist)){
  print(celltype)
  for(celltype2 in names(markerlist[[celltype]])){
    print(celltype2)
    
    for(i in names(markerlist[[celltype]][[celltype2]])){
  print(i)
  for(j in c("high","low")){
    print(j)
    genes=rownames(markerlist[[celltype]][[celltype2]][[i]])[markerlist[[celltype]][[celltype2]][[i]]$avg_log2FC>0]
    if(j=="low"){
      genes=rownames(markerlist[[celltype]][[celltype2]][[i]])[markerlist[[celltype]][[celltype2]][[i]]$avg_log2FC<0]
    }
    
    geneID=gene2id$ENTREZID[match(genes,gene2id$SYMBOL,nomatch=0)]
  ego=enrichGO(gene = geneID, OrgDb = org.Mm.eg.db, ont = "BP", pAdjustMethod = "BH", qvalueCutoff = 0.05, readable = TRUE)
  ekegg <-enrichKEGG(gene = geneID, organism = 'mmu', pAdjustMethod = "BH", qvalueCutoff = 0.05)
  
  if(!is.null(ego)){
    ego=ego@result[ego@result$pvalue<0.05,]
    enrichfunciton[[celltype]][[celltype2]][[i]][[j]][["ego"]]=ego
  }
  if(!is.null(ekegg)){
    ekegg=ekegg@result[ekegg@result$pvalue<0.05,]
    ekegg$genesymbol=sapply(ekegg$geneID,function(x){a=unlist(strsplit(x,"/"));b=gene2id$SYMBOL[match(a,gene2id$ENTREZID,nomatch = 0)];return(paste0(b,collapse = "/"))})
    enrichfunciton[[celltype]][[celltype2]][[i]][[j]][["ekegg"]]=ekegg
    
  }
  }
  
}
  }
  
}
save(enrichfunciton,markerlist,file = "enriched_function.RData")
```

```{r}
for(celltype in names(markerlist)){
  for(celltype2 in names(markerlist[[celltype]])){
    for(i  in names(markerlist[[celltype]][[celltype2]])){
  print(i)
  xlsx::write.xlsx(markerlist[[celltype]][[celltype2]][[i]],file = paste0("20220927_",celltype,"_marker_pval.xlsx"),sheetName = paste0(celltype2,"_",i),append = T)
}
  }
  
}

for(celltype in names(enrichfunciton)){
  print(celltype)
  for(celltype2 in names(enrichfunciton[[celltype]])){
    t=enrichfunciton[[celltype]][[celltype2]]
    a=do.call("c",lapply(t,function(x){do.call("c",x)}))
    writexl::write_xlsx(a,path =paste0("20220927_enrichedfunction_",celltype,celltype2,".xlsx") )
  #   for(i in names(enrichfunciton[[celltype]][[celltype2]])){
  #   print(i)
  #   for(j in c("high","low")){
  #     
  #     if(nrow(enrichfunciton[[celltype]][[celltype2]][[i]][[j]][["ego"]])>0){
  #       xlsx::write.xlsx(enrichfunciton[[celltype]][[celltype2]][[i]][[j]][["ego"]],file=paste0("20220923_enrichedfunction_",celltype,celltype2,".xlsx"),append = T,sheetName = paste0(i,j,"ego"))
  #       
  #       
  #     }
  #     if(nrow(enrichfunciton[[celltype]][[celltype2]][[i]][[j]][["ekegg"]])>0){
  #       xlsx::write.xlsx(enrichfunciton[[celltype]][[celltype2]][[i]][[j]][["ekegg"]],file=paste0("20220923_enrichedfunction_",celltype,celltype2,".xlsx"),append = T,sheetName = paste0(i,j,"ekegg"))
  #       
  #     }
  #     
  #   }
  # }
  }
  
}


```
#show pathway
```{r}
pdf("20220927.function-dot.pdf")
for(celltype in names(enrichfunciton)){
  print(celltype)
  t0=enrichfunciton[[celltype]]
  t=do.call("rbind",lapply(t0,function(x){do.call("rbind",lapply(x,function(y){y$high$ekegg}))}))
  
  t$celltype2=gsub("[.].*","",rownames(t))
  t$comp=sapply(rownames(t),function(x){unlist(strsplit(x,"[.]"))[2]})
  t$GeneRatio=sapply(t$GeneRatio,function(x){t=unlist(strsplit(x,'/'));as.numeric(t[1])/as.numeric(t[2])})
  t=t[order(t$comp),]
  t$Description=factor(t$Description,levels = t$Description[!duplicated(t$Description)])
  
  for(celltype2 in unique(t$celltype2)){
    d=t[t$celltype2%in%celltype2,]
    p=ggplot(d)+geom_point(aes(x=comp,y=Description,fill=pvalue,size=GeneRatio),shape=21)+
      scale_fill_continuous(name="pvalue",low="red",high="blue")+theme_bw()+
      theme(axis.ticks.y = element_blank(),axis.text.x = element_text(angle = 45,hjust = 1),
            axis.text.y = element_text(size = 5),plot.title = element_text(hjust = 0.5))+labs(title=paste(celltype,celltype2))
    print(p)
  }
}
dev.off()

load("E:/databases/pathway_genesets/02_kegg_API_download/kegg_api_geneSets20221019.RData")
names(mmu_kegg_list_des)=gsub(" - Mus musculus \\(mouse\\)","",names(mmu_kegg_list_des))
pdf("20220927.function-heatmap.pdf")
for(celltype in names(enrichfunciton)){
  print(celltype)
  t0=enrichfunciton[[celltype]]
  for(celltype2 in names(t0)){
    t=t0[[celltype2]]
    showpathway=unique(do.call("c",lapply(t,function(x){head(x$high$ekegg$Description,10)})))
    
    k=mmu_kegg_list_des[match(showpathway,names(mmu_kegg_list_des),nomatch = 0)]
    s=seuratData_blood10x_extract[[celltype]] ##neutrophil
    s=AddModuleScore(s,features = k,seed = 100)
    colnames(s@meta.data)[match("Cluster1",colnames(s@meta.data)):ncol(s@meta.data)]=names(k)
    
    d=s@meta.data[,names(k)]
    d=aggregate(d,by=list(orig=s$orig),median)
    rownames(d)=d$orig
    d=d[,-1]
    d=t(d)
    t=t(scale(t(d)))
    
    annot_df_col=data.frame(orig=colnames(t))
    column_color2=rainbow(length(unique(annot_df_col$orig)))
    names(column_color2)=levels(s$orig)
    anno_col=HeatmapAnnotation(df=annot_df_col,col=list(orig=column_color2))
    
    heatmap.score=Heatmap(t, name = paste(celltype,celltype2), 
                          cluster_rows = T, 
                          cluster_columns = F, 
                          # top_annotation = anno_col, use_raster=T,
                          show_column_dend = F, show_row_dend = F, 
                          clustering_distance_rows='euclidean', show_row_names = T,show_column_names = T,
                          row_km = 1)
    print(heatmap.score)
  }
  
  
}
dev.off()


```
#gsea
```{r}
load("E:/databases/mm10_geneloc/20210422_human2mice.RData")#human2mice gene symbol conversion

load("E:/databases/pathway_genesets/01_msigdb/20220428_msigdblist_human&mice.RData")#msigdb
load("E:/databases/pathway_genesets/02_kegg_API_download/kegg_api_geneSets20221019.RData") #hsa_kegg_list_des
# used_pathway=c(hsa_kegg_list_des,msigdblist$c2_01_all_canonical_pathway)
used_pathway=mmu_kegg_list_des
names(used_pathway)=gsub(" - Mus musculus \\(mouse\\)","",names(used_pathway))
pathwaylist=data.frame(pathway=rep(names(used_pathway),sapply(used_pathway,length)),gene=do.call("c",used_pathway),stringsAsFactors = F)
# pathwaylist$gene=human2mice$m[match(pathwaylist$gene,human2mice$h)]
pathwaylist=pathwaylist[complete.cases(pathwaylist),]
# pathwaylist$pathway=gsub(" - Homo sapiens \\(human\\)","",pathwaylist$pathway)

gsea_stage_list=list()
for(celltype in c(names(seuratData_blood10x_extract))){
  for(celltype2 in unique(seuratData_blood10x_extract[[celltype]]$celltype2)){
    print(celltype)
    s=seuratData_blood10x_extract[[celltype]]
    s=s[,s$celltype2%in%celltype2]
    Idents(s)="orig"
    pairstocomp=list(control_vs_WT=c("control","WT"),  control_vs_ATBX=c("control","ATBX"), ATBX_vs_WT=c("ATBX","WT"))
    for(i in names(pairstocomp)){
      print(paste(celltype,celltype2,i))
      id1=pairstocomp[[i]][1]
      id2=pairstocomp[[i]][2]
      
      genes <- FindMarkers(object = s, ident.1 =id1,ident.2 = id2, only.pos = F, min.pct = 0.1, logfc.threshold = 0,test.use = 'wilcox')
      genes=genes[order(genes$avg_log2FC,decreasing = T),]
      geneList=genes$avg_log2FC
      names(geneList)=rownames(genes)
      gsea <- clusterProfiler::GSEA(geneList, TERM2GENE = pathwaylist)
      gsea_stage_list[[celltype]][[celltype2]][[i]]=gsea
    }
  }  
}
save(gsea_stage_list,file = "gsea_stage_list.RData")
```

```{r}
for(celltype in names(gsea_stage_list)){
  print(celltype)
  for(celltype2 in names(gsea_stage_list[[celltype]])){
    t=gsea_stage_list[[celltype]][[celltype2]]
    t=lapply(t,function(x){x@result})
    writexl::write_xlsx(t,path =paste0("20220927_gsea_",celltype,celltype2,".xlsx") )

  }
  
}
```

#cellchat
# run on cluster
```{r}
library(CellChat)
cellchat_list=list()
cellchat_df_list=list()
Sys.time()
for(orig in c("WT", "control", "ATBX")){
  print(orig)
  s=seuratData[,seuratData@meta.data$orig==orig]
  cellchat=createCellChat(object = s,meta=s@meta.data,group.by = "celltype")
  rm(s)
  cellchat@DB=CellChatDB.mouse
  cellchat=subsetData(cellchat)
  future::plan("multiprocess",workers=4)
  cellchat=identifyOverExpressedGenes(cellchat)
  cellchat=identifyOverExpressedInteractions(cellchat)
  cellchat=projectData(cellchat,PPI.mouse)
  cellchat=computeCommunProb(cellchat)
  cellchat= filterCommunication(cellchat, min.cells = 4)
  df.net <- subsetCommunication(cellchat)
  
  
  cellchat <- computeCommunProbPathway(cellchat)
  cellchat <- aggregateNet(cellchat)
  cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")
  
  cellchat_df_list[[orig]]=df.net
  cellchat_list[[orig]] =cellchat
  
}

save(cellchat_df_list,cellchat_list,file = "cellchat_celltype.RData")
Sys.time()

library(CellChat)
cellchat_list=list()
cellchat_df_list=list()
Sys.time()
for(orig in c("WT", "control", "ATBX")){
  print(orig)
  s=seuratData[,seuratData@meta.data$orig==orig]
  cellchat=createCellChat(object = s,meta=s@meta.data,group.by = "celltype2")
  rm(s)
  cellchat@DB=CellChatDB.mouse
  cellchat=subsetData(cellchat)
  future::plan("multiprocess",workers=4)
  cellchat=identifyOverExpressedGenes(cellchat)
  cellchat=identifyOverExpressedInteractions(cellchat)
  cellchat=projectData(cellchat,PPI.mouse)
  cellchat=computeCommunProb(cellchat)
  cellchat= filterCommunication(cellchat, min.cells = 4)
  df.net <- subsetCommunication(cellchat)
  
  
  cellchat <- computeCommunProbPathway(cellchat)
  cellchat <- aggregateNet(cellchat)
  cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")
  
  cellchat_df_list[[orig]]=df.net
  cellchat_list[[orig]] =cellchat
  
}

save(cellchat_df_list,cellchat_list,file = "cellchat_celltype2.RData")
Sys.time()
```


```{r}
load("X:/analyze_data/FEdata/yaobingqing/20220922_GF/3sample/20220928_cellchat/cellchat_celltype2.RData")
cellchat <- mergeCellChat(cellchat_list, add.names = names(cellchat_list))

pdf(paste0(dataname0,"_interaction_count_strength2.pdf"))
netVisual_diffInteraction(cellchat, weight.scale = T, measure = "weight",comparison = c(2,1),title.name = "WT_vs_con")
netVisual_diffInteraction(cellchat, weight.scale = T, measure = "weight",comparison = c(2,3),title.name = "ATBX_vs_con")

for(i in colnames(cellchat@net[[2]]$weight)){
  print(i)
  # if(i%in%c("Tresident","cDC2")){next()}
  netVisual_diffInteraction(cellchat, weight.scale = T,sources.use = i,comparison = c(2,1),title.name = "WT_vs_con")
  netVisual_diffInteraction(cellchat, weight.scale = T,targets.use  = i,comparison = c(2,1),title.name = "WT_vs_con")
    netVisual_diffInteraction(cellchat, weight.scale = T,sources.use = i,comparison = c(2,3),title.name = "ATBX_vs_con")
  netVisual_diffInteraction(cellchat, weight.scale = T,targets.use  = i,comparison = c(2,3),title.name = "ATBX_vs_con")
  
}
dev.off()

```

```{r}
load("X:/analyze_data/FEdata/yaobingqing/20220922_GF/enriched_function.RData")
# markerlist

restore_list=list()
restore_list_all=list()
origchange_list=list()
origchange_list_all=list()
for(i in names(markerlist)){
  for(j in names(markerlist[[i]])){
    t1=markerlist[[i]][[j]]$control_vs_WT
    t2=markerlist[[i]][[j]]$control_vs_ATBX
    t=intersect(rownames(t1)[t1$avg_log2FC>0],rownames(t2)[t2$avg_log2FC>0])
    origchange_list[[i]][[j]][["high"]]=rownames(t1)[t1$avg_log2FC>0]
    restore_list[[i]][[j]][["high"]]=t
    t=intersect(rownames(t1)[t1$avg_log2FC<0],rownames(t2)[t2$avg_log2FC<0])
    restore_list[[i]][[j]][["low"]]=t
    origchange_list[[i]][[j]][["low"]]=rownames(t1)[t1$avg_log2FC<0]
    restore_list_all[[i]][[j]]=c(restore_list[[i]][[j]][["high"]],restore_list[[i]][[j]][["low"]])
    # restore_list_all[[i]][[j]]=c(intersect(rownames(t1),rownames(t2)))
    origchange_list_all[[i]][[j]]=rownames(t1)
  }
}

t1=do.call("c",lapply(restore_list,function(x){do.call("c",x)}))
t2=do.call("c",lapply(origchange_list,function(x){do.call("c",x)}))
data=data.frame(origchange=sapply(t2,length),restore=sapply(t1,length))
data$orig=rownames(data)
data$rate=data$restore/data$origchange
rownames(data)=names(t1)
data$celltype2=gsub(".high|.low","",rownames(data))
data$celltype=gsub("[.].*","",rownames(data))
data$highlow="low"
data$highlow[grep("high",rownames(data))]="high"
ggplot(data,aes(orig,rate,fill=highlow))+geom_bar(stat = "identity")+theme_classic()+
  theme(axis.text.x = element_text(hjust = 1,vjust=0.5,angle = 90))

t1=do.call("c",restore_list_all)
t2=do.call("c",origchange_list_all)
data=data.frame(origchange=sapply(t2,length),restore=sapply(t1,length))
data$orig=rownames(data)
data$rate=data$restore/data$origchange
rownames(data)=names(t1)
data$celltype2=rownames(data)
data$celltype=gsub("[.].*","",rownames(data))
data=data[order(data$rate,decreasing = T),]
data$orig=factor(data$orig,levels = data$orig[!duplicated(data$orig)])
pdf("gene.restoration_rate.pdf")
ggplot(data,aes(orig,rate,fill=celltype))+geom_bar(stat = "identity")+theme_classic()+
  theme(axis.text.x = element_text(hjust = 1,vjust=0.5,angle = 90))
dev.off()



scicolors=ggsci::pal_lancet("lanonc")(9)
cellchat3=mergeCellChat(cellchat_list,add.names = names(cellchat_list),cell.prefix = F)
p1=netVisual_heatmap(cellchat3,comparison = c(1,2),title.name = "conVSWT-num")
p2=netVisual_heatmap(cellchat3,comparison = c(2,3),title.name = "ATBXVScon-num")

p3=netVisual_heatmap(cellchat3,comparison = c(1,2),measure = "weight",title.name = "conVSWT-strength")
p4=netVisual_heatmap(cellchat3,comparison = c(2,3),measure = "weight",title.name = "ATBXVScon-strength")
pdf("difference_interaction.pdf")
p1+p2
p3+p4
dev.off()

object.list=cellchat_list
num.link <- sapply(object.list, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(num.link), max(num.link)) # control the dot size in the different datasets
gg <- list()
for (i in 1:length(object.list)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(object.list[[i]], 
                                               title = names(object.list)[i], 
                                               weight.MinMax = weight.MinMax)+xlim(c(0,20))+ylim(c(0,15))
}
pdf("outgong&incoming1.pdf",width = 15)
gg[[1]]+gg[[2]]+gg[[3]]
dev.off()

pathway.union <- sort(unique(do.call("c",lapply(object.list,function(x){x@netP$pathways}))))
gg <- list()
gg2=list()
gg3=list()
for(i in 1:length(object.list)){
  gg[[i]]=netAnalysis_signalingRole_heatmap(object.list[[i]],font.size = 5,pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i], width = 5, height = 10)
  gg2[[i]]=netAnalysis_signalingRole_heatmap(object.list[[i]],font.size = 2,  pattern = "incoming", signaling = pathway.union, title = names(object.list)[i], width = 5, height = 10, color.heatmap = "OrRd")
  gg3[[i]]=netAnalysis_signalingRole_heatmap(object.list[[i]],font.size = 2,  pattern = "all", signaling = pathway.union, title = names(object.list)[i], width = 5, height = 10)
}
pdf("outgong&incoming2.pdf",width = 10)#,height = 15
gg[[1]]+gg[[2]]+gg[[3]]
gg2[[1]]+gg2[[2]]+gg2[[3]]
gg3[[1]]+gg3[[2]]+gg3[[3]]

dev.off()
allcells=colnames(cellchat3@netP$WT$prob[,,1])
pdf("netVisual_bubble0.pdf",height = 10,width = 10)
netVisual_bubble(cellchat3, sources.use = c("neutrophil"),comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis")
netVisual_bubble(cellchat3, targets.use = c("neutrophil"),comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis")
netVisual_bubble(cellchat3, sources.use = c("tissue_mac"),comparison = c(1,2,3), angle.x = 45,color.text=scicolors[1:3],color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("cDC1"),comparison = c(1,2,3), angle.x = 45,color.text=scicolors[1:3],color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("cDC2"),comparison = c(2,3),angle.x = 45,color.text = c(scicolors[2:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("mRegDC"),comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("endo"),targets.use = allcells[1:7],comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("endo"),targets.use = allcells[8:11],comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("endo"),targets.use = allcells[11:15],comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("fibroblast"),targets.use = allcells[1:3],comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("fibroblast"),targets.use = allcells[4:7],comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("fibroblast"),targets.use = allcells[8:10],comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("fibroblast"),targets.use = allcells[11:13],comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 6)
netVisual_bubble(cellchat3, sources.use = c("fibroblast"),targets.use = allcells[14:15],comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, sources.use = c("alveolar_mac"),comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, targets.use = c("cDC2"),comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)
netVisual_bubble(cellchat3, targets.use = c("Tcytotoxic"),comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis",font.size = 8)

dev.off()

# netVisual_bubble(cellchat3,comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),signaling = pathway.union,color.heatmap = "viridis")


pathway3=intersect(intersect(cellchat3@netP$WT$pathways,cellchat3@netP$control$pathways),cellchat3@netP$ATBX$pathways)
# par(mfrow=c(1,3))
pdf("circle_PPP.pdf")
for(p in pathway3){
  pathways.show <- c(p) 
  weight.max <- getMaxWeight(object.list, slot.name = c("netP"), attribute = pathways.show)
  
  gg=list()
  for(i in 1:length(object.list)){
    netVisual_aggregate(object.list[[i]], signaling = pathways.show, layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object.list)[i]))
    
  }
  
}
dev.off()

pathwaydiff_high=intersect(setdiff(cellchat3@netP$control$pathways,cellchat3@netP$WT$pathways),
                      setdiff(cellchat3@netP$control$pathways,cellchat3@netP$ATBX$pathways)
                      )
pdf("circle_NPN.pdf")
i=2
pathways.show=pathwaydiff_high
weight.max=max(cellchat3@netP$control$prob[,,p])
netVisual_aggregate(object.list[[i]], signaling = pathways.show, layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object.list)[i]))
dev.off()
pathwaydiff_low=intersect(setdiff(cellchat3@netP$WT$pathways,cellchat3@netP$control$pathways),
                           setdiff(cellchat3@netP$ATBX$pathways,cellchat3@netP$control$pathways)
)
pdf("circle_PNP.pdf")
for(p in pathwaydiff_low){
  
  pathways.show <- c(p) 
  weight.max <- getMaxWeight(object.list[c(1,3)], slot.name = c("netP"), attribute = pathways.show)
  
  gg=list()
  for(i in c(1,3)){
    netVisual_aggregate(object.list[[i]], signaling = pathways.show, layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object.list)[i]))
    
  }
}
dev.off()
pathwaydiff_high_retain=intersect(setdiff(cellchat3@netP$control$pathways,cellchat3@netP$WT$pathways),cellchat3@netP$ATBX$pathways)
pdf("circle_NPP.pdf")
for(p in pathwaydiff_high_retain){
  
  pathways.show <- c(p) 
  weight.max <- getMaxWeight(object.list[c(2,3)], slot.name = c("netP"), attribute = pathways.show)
  
  gg=list()
  for(i in c(2,3)){
    netVisual_aggregate(object.list[[i]], signaling = pathways.show, layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object.list)[i]))
    
  }
}
dev.off()
pathwaydiff_low_retain=setdiff(setdiff(cellchat3@netP$WT$pathways,cellchat3@netP$control$pathways),cellchat3@netP$ATBX$pathways)
pdf("circle_PNN.pdf")
for(p in pathwaydiff_low_retain){
  
  pathways.show <- c(p) 
  weight.max <- getMaxWeight(object.list[c(1)], slot.name = c("netP"), attribute = pathways.show)
  
  gg=list()
  for(i in c(1)){
    netVisual_aggregate(object.list[[i]], signaling = pathways.show, layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object.list)[i]))
    
  }
}
dev.off()




pathwaytoshow=list(NPN=pathwaydiff_high,PNP=pathwaydiff_low,NPP=pathwaydiff_high_retain,PNN=pathwaydiff_low_retain)
for(i in names(pathwaytoshow)){
  for(j in pathwaytoshow[[i]]){
    print(paste(i,j))
    genes=cellchat3@DB$interaction$interaction_name_2[cellchat3@DB$interaction$pathway_name==j]
    genes=unique(do.call("c",lapply(genes,function(x){x=gsub(" ","",x);x=gsub("[()]","+",x);a=unlist(strsplit(x,"[+-]"));a=a[a!=""];return(a)})))
    genes=sort(genes)
    genes=genes[genes%in%rownames(seuratData)]
    p=VlnPlot(seuratData,features = genes,group.by = "celltype2",pt.size = 0,ncol = 10,split.by="orig")
    tiff(paste0("20221026_pathwaygenes_vln-",i,"-",j,".tiff"),res = 300,width = 20000,height = 1000*(ceiling(length(genes)/10)))
    print(p)
    dev.off()
  }
}


pdf(paste0(dataname0,"netVisual_circle.pdf"))
for(type in names(cellchat_list)){
  cellchat=cellchat_list[[type]]
  cellchat <- aggregateNet(cellchat)
  
  groupSize <- as.numeric(table(cellchat@idents))
  par(mfrow=c(3,3))
  netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = paste(type,"Number of interactions"))
  netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = paste(type,"Interaction weights/strength"))
  
  mat=cellchat@net$weight
  
  for(i in 1:nrow(mat)){
    mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
    mat2[i, ] <- mat[i, ]
    netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
  }
  netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = paste(type,"Number of interactions"))
  netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = paste(type,"Interaction weights/strength"))
  
  for(i  in 1:ncol(mat)){
    mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
    mat2[,i ] <- mat[,i ]
    netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
    
  }
  
}



```
## selected to show
```{r}

fromtoLG=list(from=list(neutrophil=list(tissue_mac=c("Anxa1 - Fpr1","Ccl3 - Ccr5","Entpd1 - Adora2b"),
                                        cDC1=c("Thbs1 - (Itga3+Itgb1)","Thbs1 - Sdc4"),
                                        cDC2=c("Cd52 - Siglecg","Cd80 - Cd28","H2-d1 - Cd8a","H2-k1 - Cd8a","Osm - (Lifr+Il6st)","Nampt - Insr","Spp1 - Cd44"),
                                        mRegDC=c("Thbs1 - Cd36"),
                                        CD4Tnaive=c("Il16 - Cd4"),
                                        CD8Tnaive=c("Itgb2 - Icam1"),
                                        Treg=c("Il16 - Cd4"),
                                        Tresident=c("Cd80 - Cd274","Entpd1 - Adora2a","Itgb2 - Icam1","Spp1 - (Itga4+Itgb1)"),
                                        Tcytotoxic=c("H2-d1 - Cd8a","Itgb2 - Cd226"),
                                        Bcell=c("Selplg - Sell"),
                                        NK=c("Selplg - Sell"),
                                        endo=c("Osm - (Osmr+Il6st)","Osm - (Lifr+Il6st)","Sema4a - (Nrp1+Plxna2)","Spp1 - (Itgav+Itgb1)"),
                                        fibroblast=c("Osm - (Osmr+Il6st)"),
                                        alveolar_mac=c("Vegfa - Vegfr1")
                                        )),
              to=list(neutrophil=list(tissue_mac=c("Bst2 - Pira2","Fn1 - Cd44","Mif - (Cd74+Cxcr2)","Thbs1 - Cd47","Tnf - Tnfrsf1b","Tnf - Tnfrsf1a"),
                                      cDC1=c("Jam1 - (Itgal+Itgb2)","Ccl5 - Ccr1","Icam1 - (Itgax+Itgb2)","Lgals9 - Cd44","Lgals9 - Cd45","Mif - (Cd74+Cxcr2)"),
                                      mRegDC=c("Ccl6 - Ccr1","Cxcl1 - Cxcr2","Cxcl2 - Cxcr2"),
                                      CD4Tnaive=c("Mif - (Cd74+Cxcr2)"),
                                      CD8Tnaive=c("Icam1 - (Itgal+Itgb2)","Icam1 - (Itgam+Itgb2)","Icam1 - (Itgax+Itgb2)","Icam1 - Itgal","Mif - (Cd74+Cd44)","Mif - (Cd74+Cxcr2)","Mif - (Cd74+Cxcr4)"),
                                      Treg=c("Cxcl2 - Cxcr2","Mif - (Cd74+Cxcr4)","Mif - (Cd74+Cxcr2)","Mif - (Cd74+Cd44)"),
                                      Tresident=c("Icam1 - Itgal","Icam1 - (Itgax+Itgb2)","Icam1 - (Itgam+Itgb2)","Icam1 - (Itgal+Itgb2)","Mif - (Cd74+Cxcr4)","Mif - (Cd74+Cxcr2)","Mif - (Cd74+Cd44)"),
                                      Tcytotoxic=c("Mif - (Cd74+Cxcr4)","Mif - (Cd74+Cxcr2)","Mif - (Cd74+Cd44)"),
                                      Bcell=c("Mif - (Cd74+Cxcr2)"),
                                      NK=c("Ccl5 - Ccr1","Mif - (Cd74+Cxcr2)"),
                                      endo=c("Anxa1 - Fpr1","Anxa1 - Fpr2","Cxcl2 - Cxcr2","Lamb2 - Cd44","Lgals9 - Cd44","Lgals9 - Cd45"),
                                      fibroblast=c("Col4a5 - Cd44","Csf1 - Csf1r","Cxcl2 - Cxcr2","Jam3 - (Itgam+Itgb2)"),
                                      alveolar_mac=c("App - Cd74","Fn1 - Cd44","Lgals9 - Cd44","Lgals9 - Cd45")
                                      ),
                      Tcytotoxic=list(neutrophil=c("H2-k1 - Cd8a","H2-d1 - Cd8a","Itgb2 - Cd226"),
                                      tissue_mac=c("Cd80 - Cd28","Fn1 - Cd44","Fn1 - (Itga4+Itgb7)","Fn1 - (Itga4+Itgb1)","H2-d1 - Cd8a","H2-k1 - Cd8a","H2-q4 - Cd8a","H2-q6 - Cd8a","H2-q7 - Cd8a","H2-t23 - Cd8a","Il18 - (Il18r1+Il18rap)","Itgb2 - Cd226"),
                                      cDC1=c("H2-d1 - Cd8a","H2-k1 - Cd8a","H2-q4 - Cd8a","H2-t23 - Cd8a","Itgb2 - Cd226","Pvr - Cd226","Spp1 - (Itga4+Itgb1)"),
                                      cDC2=c("Ccl4 - Ccr5","Ccl5 - Ccr5","Ccl6 - Ccr2","Entpd1 - Adora2a","Itgb2 - Cd226","Pvr - Cd226"),
                                      mRegDC=c("Cd70 - Cd27","H2-t23 - Cd8a","Pvr - Cd226"),
                                      CD4Tnaive=c("Ccl6 - Ccr2"),
                                      CD8Tnaive=c("Ccl6 - Ccr2","Icam1 - (Itgal+Itgb2)","Icam1 - Itgal","Icam1 - Spn","Pvr - Cd226"),
                                      Treg=c("Itgb2 - Cd226","Pvr - Cd226","Icam1 - Itgal","Icam1 - Spn","Pvr - Cd226"),
                                      Tcytotoxic=c("Gzma - F2r","Itgb2 - Cd226","Spp1 - (Itga4+Itgb1)"),
                                      Bcell=c("Ccl6 - Ccr2"),
                                      NK=c("Itgb2 - Cd226","Spp1 - (Itga4+Itgb1)"),
                                      endo=c("Ccl6 - Ccr2","Dll4 - Notch1","Entpd1 - Adora2a"),
                                      fibroblast=c("Col4a5 - Cd44"),
                                      alveolar_mac=c("Fn1 - (Itga4+Itgb1)","Fn1 - (Itga4+Itgb7)","Fn1 - Cd44","Itgb2 - Cd226","Lgals9 - Cd44","Lgals9 - Cd45")
                                      )
                      )
              )


for(i in names(fromtoLG)){
  # print(i)
  for(j in names(fromtoLG[[i]])){
    # print(j)
    print(paste(i,j,length(fromtoLG[[i]][[j]]),sum(names(fromtoLG[[i]][[j]])%in%colnames(cellchat3@netP$WT$prob[,,1]))))
    if(length(fromtoLG[[i]][[j]])!=sum(names(fromtoLG[[i]][[j]])%in%colnames(cellchat3@netP$WT$prob[,,1]))){
      print(setdiff(names(fromtoLG[[i]][[j]]),colnames(cellchat3@netP$WT$prob[,,1])))
    }
  }
}

fromtoLG_path=list()
for(i in names(fromtoLG)){
  for(j in names(fromtoLG[[i]])){
    for(k in names(fromtoLG[[i]][[j]])){
      t=fromtoLG[[i]][[j]][[k]]
      t=gsub(" -","  -",t)
      if(sum(is.na(match(t,cellchat3@DB$interaction$interaction_name_2)))>0){print(paste(i,j,k))}
      fromtoLG_path[[i]][[j]][[k]]=cellchat3@DB$interaction$pathway_name[match(t,cellchat3@DB$interaction$interaction_name_2)]
    }
  }
}

pdf("selected.pathway.bubble.pdf",height = 10,width = 10)
for(i in names(fromtoLG_path$from)){
  ps=unique(do.call("c",fromtoLG_path$from[[i]]))
  p=netVisual_bubble(cellchat3, sources.use = c(i),targets.use = names(fromtoLG_path$from[[i]]),
                   signaling = ps,
                   comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis")
  print(p)
}
for(i in names(fromtoLG_path$to)){
  ps=unique(do.call("c",fromtoLG_path$to[[i]]))
  p=netVisual_bubble(cellchat3, targets.use = c(i),sources.use = names(fromtoLG_path$from[[i]]),
                   signaling = ps,
                   comparison = c(1,2,3),angle.x = 45,color.text = c(scicolors[1:3]),color.heatmap = "viridis")
  print(p)
}
dev.off()

pdf("selected.pathway.circle.pdf",width = 10,height = 10)
for(i in names(fromtoLG_path$from)){
  t1=i
  t2=names(fromtoLG_path$from[[i]])
  ps=unique(do.call("c",fromtoLG_path$from[[i]]))

  for(con in 1:length(cellchat_list)){
    netVisual_chord_gene(cellchat_list[[con]], sources.use = t1, targets.use = t2, 
                         slot.name = 'net', signaling = ps, lab.cex = 0.5, small.gap = 0, 
                         title.name = paste0(names(cellchat_list)[con]))
  }
}
for(i in names(fromtoLG_path$to)){
  t2=i
  t1=names(fromtoLG_path$to[[i]])
  ps=unique(do.call("c",fromtoLG_path$to[[i]]))
  for(con in 1:length(cellchat_list)){
    netVisual_chord_gene(cellchat_list[[con]], sources.use = t1, targets.use = t2, 
                         slot.name = 'net', signaling = ps, lab.cex = 0.5, small.gap = 0, 
                         title.name = paste0(names(cellchat_list)[con]))
  }
}
dev.off()






fromtoLG=list(from=list(neutrophil=c("H2-d1 - Cd8a","Ccl6 - Ccr2")),
              to=list(neutrophil=c("Ccl5 - Ccr1"))
              )
fromtoLG_path=list()
for(i in names(fromtoLG)){
  for(j in names(fromtoLG[[i]])){
      t=fromtoLG[[i]][[j]]
      t=gsub(" -","  -",t)
      t=gsub(" −","  −",t)
      if(sum(is.na(match(t,cellchat3@DB$interaction$interaction_name_2)))>0){print(paste(i,j))}
      fromtoLG_path[[i]][[j]]=cellchat3@DB$interaction$pathway_name[match(t,cellchat3@DB$interaction$interaction_name_2)]
    }
}



pdf("selected.pathway.circle2.pdf",width = 10,height = 10)
for(i in names(fromtoLG_path$from)){
  t1=i
  ps=fromtoLG_path$from[[i]]
  for(j in ps){
    for(con in 1:length(cellchat_list)){
      netVisual_chord_gene(cellchat_list[[con]], sources.use = i,# targets.use = t2, 
                           slot.name = 'net', signaling = j, lab.cex = 0.5, small.gap = 0, 
                           title.name = paste(names(cellchat_list)[con],j))
    }
  }
  
}
for(i in names(fromtoLG_path$to)){
  t2=i
  ps=fromtoLG_path$to[[i]]
  for(j in ps){
    for(con in 1:length(cellchat_list)){
      netVisual_chord_gene(cellchat_list[[con]],  targets.use = t2, 
                           slot.name = 'net', signaling = j, lab.cex = 0.5, small.gap = 0, 
                           title.name = paste(names(cellchat_list)[con],j))
    }
  }
  
}
dev.off()


```



#GMP pseudotime
```{r}
# markers=c("CD81","Cd155","CD29","CD43","CD47","CD71","CD49d","CD49e","CD184","CD205","CD11b","CD88","CD18","CD55","CD11a","Cd115","Ly6c1")
markers=c("CD81","Pvr","Itgb1","Spn","CD47","Tfrc","Itga4","Itga5","Cxcr4","Ly75","Itgam","C5ar1","Itgb2","CD55","Itgal","Csf1r","Ly6c1")
markers=first2big(markers)
GMP=c("Kit","Cd34","Fcgr3","Fcgr2b","Ly6a")
monocytic=c("Irf8", "Ly86", "Csf1r")
Basophilic=c("Gata2","Mcpt8","Prss34")
neutrophilic=c("Elane","Fcnb","Gfi1", "Rgcc", "Vcam1")
Eosinophilic=c("Prg2")

f=c(GMP,monocytic,Basophilic,neutrophilic,markers)
f=f[!duplicated(f)]

p3=FeaturePlot(seuratData,features = c(f),ncol = 10)

tiff(paste0("GMP",".markergene.tiff"),res = 300,width = 10000,height = 1000*4)
print(p3)
dev.off()

p3=FeaturePlot(seuratData_blood10x_extract$neutrophil,features = c(f),ncol = 10)

tiff(paste0("GMP",".markergene-neutrophil.tiff"),res = 300,width = 10000,height = 1000*4)
print(p3)
dev.off()
```

##cellchat

```{r}


```

## pseudotime
```{r}

#!/usr/bin/R
library(Biobase)
library(knitr)
library(reshape2)
library(ggplot2)
library(monocle)
library(RColorBrewer)
library(Seurat)
library(colorRamps)
library(pheatmap)
library(stringr)
library(GSVA)
library(ComplexHeatmap)
library(dendsort)

setwd("/storage/caishangLab/caishang/sharedata/LXQ/singlecellAnalysis/trajectory/neutrophil")
#load("20210902_vprm_stromalseuratDatarm1col1a1.2.RData")
#load("20211030_vp30wttumorrm_stromalseuratDatarm1col1a1.2.RData")
#load("20211031_vp30CD200Prm_stromalseuratDatarm1col1a1.2.RData")
#load("20211103_vp30_muttumorrm_stromalseuratDatarm1col1a1.2.RData")
load("neutrophil.RData")
seuratData=s
rm(s)
#setwd("/storage/caishangLab/caishang/sharedata/LXQ/singlecellAnalysis/trajectory/neutrophil/0")
dataname0="20220929_neutrophil"
#for(mean_exp in seq(0.201,0.4,by=0.001)){
  mean_exp=0.196
  print(mean_exp)
  feature_sheet=data.frame(row.names = rownames(seuratData@assays$RNA@counts),gene_short_name=rownames(seuratData@assays$RNA@counts))
pd <- new("AnnotatedDataFrame", data = seuratData@meta.data)
fd <- new("AnnotatedDataFrame", data =feature_sheet)

HSMM=newCellDataSet(seuratData@assays$RNA@counts,
                    phenoData = pd,
                    featureData = fd,
                    lowerDetectionLimit=0.5,
                    expressionFamily=negbinomial.size())

HSMM <- detectGenes(HSMM, min_expr = 0.1)

# expressed_genes <- row.names(subset(fData(HSMM), num_cells_expressed >= 10))

HSMM <- estimateSizeFactors(HSMM)
HSMM=estimateDispersions(HSMM)

disp_table <- dispersionTable(HSMM)
# HSMM_ordering_genes <- subset(disp_table, mean_expression >= 1 & dispersion_empirical >= 1 * dispersion_fit)$gene_id

HSMM_ordering_genes <- subset(disp_table, mean_expression >= mean_exp & dispersion_empirical >= 1 * dispersion_fit)$gene_id
HSMM <-setOrderingFilter(HSMM,ordering_genes = HSMM_ordering_genes)
# plot_ordering_genes(HSMM)
  
HSMM <-reduceDimension(HSMM,max_components = 2, reduction_method = 'DDRTree',residualModelFormulaStr = "~Size_Factor+num_genes_expressed+percent.mt+S.Score+G2M.Score")#

HSMM <-orderCells(HSMM)#,root_state = 8)

# HSMM <-orderCells(HSMM)
#table(pData(HSMM)[,c('month','State')])
#table(pData(HSMM)[,c('month_version','State')])
p1=plot_cell_trajectory(HSMM, color_by = "State",cell_size = 1,show_tree = F,show_backbone = F)+
  scale_color_manual(values =rainbow(length(unique(pData(HSMM)$State))), name = "State")
# plot_cell_trajectory(HSMM, color_by = "Cluster")+ scale_color_manual(values =rainbow(length(unique(pData(HSMM)$Cluster))), name = "Cluster")
p2=plot_cell_trajectory(HSMM, color_by = "seurat_clusters",cell_size = 1,show_tree = F,show_backbone = F)+ 
  scale_color_manual(values =rainbow(length(unique(pData(HSMM)$seurat_clusters))), name = "seurat_clusters")
p3=plot_cell_trajectory(HSMM, color_by="orig",cell_size = 1,show_tree = F,show_backbone = F)+labs(title =which.max(table(pData(HSMM)[,c('orig','State')])[1,]))+
  scale_color_manual(values =rainbow(length(unique(pData(HSMM)$orig))), name = "orig")
p4=plot_cell_trajectory(HSMM, color_by="stage",cell_size = 1,show_tree = F,show_backbone = F)
  scale_color_manual(values =rainbow(length(unique(pData(HSMM)$orig))), name = "orig")

pdf(paste0(dataname0,"mean_exp",mean_exp,"trajectory.pdf"))
print(p1)
print(p2)
print(p3)
print(p4)
dev.off()
HSMM <-orderCells(HSMM,root_state = which.max(table(pData(HSMM)[,c('orig','State')])[1,]))
p1=ggplot(pData(HSMM),aes(x=Pseudotime,col=Phase))+geom_density(size=1)+scale_color_manual(values = rainbow(length(unique(pData(HSMM)$Phase))))+theme_bw()+theme(plot.title = element_text(hjust = 0.5),axis.title.x = element_blank(),panel.border = element_blank(),panel.grid = element_blank(),axis.line = element_line(size=1,colour = 'black'))

p2=ggplot(pData(HSMM),aes(x=Pseudotime,col=orig))+geom_density(size=1)+scale_color_manual(values = rainbow(length(unique(pData(HSMM)$orig))))+theme_bw()+theme(plot.title = element_text(hjust = 0.5),axis.title.x = element_blank(),panel.border = element_blank(),panel.grid = element_blank(),axis.line = element_line(size=1,colour = 'black'))

p3=ggplot(pData(HSMM),aes(x=Pseudotime,col=orig))+geom_density(col=NA,size=1)+
  scale_color_manual(values = rainbow(length(unique(pData(HSMM)$orig))))+
  theme_bw()+theme(plot.title = element_text(hjust = 0.5),axis.title.x = element_blank(),panel.border = element_blank(),
                   panel.grid = element_blank(),axis.line = element_line(size=1,colour = 'black'))+
  facet_wrap(~orig,nrow = length(unique(pData(HSMM)$orig)),scales = 'free_y')+
  geom_line(stat='density')+
  theme(strip.background = element_rect(fill = NA,size = 0.5),
        strip.text = element_blank())


pdf(paste0(dataname0,"mean_exp",mean_exp,"density.pdf"))
print(p3)
print(p1)
print(p2)
dev.off()

#}

HSMM=HSMM[,order(pData(HSMM)$Pseudotime)]
seuratData_neutrophil=seuratData
HSMM_neutrophil=HSMM
save(seuratData_neutrophil,HSMM_neutrophil,file=paste0(dataname0,"seuratData.HSMM.RData"))
diff_test_res <- differentialGeneTest(HSMM,fullModelFormulaStr = "~sm.ns(Pseudotime)")
save(diff_test_res,file=paste0(dataname0,"HSMM_Pseudotime_diffgene.RData"))
```

```{r}
load("F:/Edata/yaobingqing/20220922_GF/3sample/20220929_pseudotime/20220929_neutrophilseuratData.HSMM.RData")
load("F:/Edata/yaobingqing/20220922_GF/3sample/20220929_pseudotime/20220929_neutrophilHSMM_Pseudotime_diffgene.RData")
xlsx::write.xlsx(diff_test_res,file = paste0(dataname0,"diff_test_res.xlsx"),sheetName = "diff_test_res")

```
```{r}
diff_test_res=diff_test_res[diff_test_res$use_for_ordering,]#591
# diff_test_res=diff_test_res[diff_test_res$pval<0.05,]#428
diff_test_res=diff_test_res[diff_test_res$qval<0.1,]#350

sig_gene_names <- row.names(diff_test_res)

###group into increasing and decreasing gene

pData(HSMM)$month=pData(HSMM)$orig
annotation_col=data.frame(row.names =rownames(pData(HSMM)),month=pData(HSMM)$month,State=pData(HSMM)$State,Phase=pData(HSMM)$Phase)

num=length(unique(pData(HSMM)$month))
num2=length(unique(pData(HSMM)$State))
num3=length(unique(pData(HSMM)$Phase))
anno_col=list(month=rainbow(num)[1:num],State=rainbow(num2)[1:num2],Phase=rainbow(num3)[1:num3])
names(anno_col$month)=levels(pData(HSMM)$month)
names(anno_col$State)=unique(pData(HSMM)$State)
names(anno_col$Phase)=unique(pData(HSMM)$Phase)
```

```{r}
annot_df_col=pData(HSMM)[,c("month","Phase","Pseudotime")]
column_color2=rainbow(length(unique(annot_df_col$month)))
names(column_color2)=levels(annot_df_col$month)
column_color3=rainbow(length(unique(annot_df_col$Phase)))
names(column_color3)=unique(annot_df_col$Phase)

anno_col=HeatmapAnnotation(df=annot_df_col,col=list(month=column_color2,Phase=column_color3,Pseudotime=circlize::colorRamp2(c(0, max(annot_df_col$Pseudotime)), c("white","darkgreen"))))

heatmap_matrix=plot_pseudotime_heatmap2_return_heatmap_matrix(HSMM[sig_gene_names,],cluster_rows = T,
                                                              add_annotation_col = annotation_col,
                                                              return_heatmap = T,hclust_method = 'ward.D2',
                                                              num_clusters=1)
bks <- seq(-3.1, 3.1, by = 0.1)
hmcols <- blue2green2red(length(bks) - 1)

max_ind=apply(heatmap_matrix,1,function(x){a=order(x,decreasing = T)[1:20];xcol=ncol(heatmap_matrix);
s1=sum(sapply(a,function(y){y<0.15*xcol}))
s2=sum(sapply(a,function(y){y>=0.15*xcol&y<0.5*xcol}))
s3=sum(sapply(a,function(y){y>=0.5*xcol&y<0.85*xcol}))
s4=sum(sapply(a,function(y){y>=0.85*xcol}))
return(paste0("cluster",which.max(c(s1,s2,s3,s4))))
})
max_ind=factor(max_ind,levels = paste0("cluster",1:4))
heatmapresult=Heatmap(as.matrix(heatmap_matrix), name = "scaledata", 
                      cluster_rows = T, 
                      cluster_columns = F, 
                      top_annotation = anno_col, use_raster=T,
                      # column_split = annot_df_col$type,
                      row_split= max_ind,
                      # left_annotation = anno_row,
                      show_column_dend = F, show_row_dend = F, 
                      clustering_distance_rows='euclidean', show_row_names = F,show_column_names = F,
                      col=hmcols,
                      #column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                      row_km = 1)
tiff(paste0(dataname0,"heatmap.tiff"))#,width = 1000,height = 800,res = 150
h1.1=draw(heatmapresult)
dev.off()


r.dend=column_dend(h1.1)
rc1.list=row_order(h1.1)
# names(rc1.list)=paste0("cluster",1:15,".")
names(rc1.list)=paste0(names(rc1.list),".")

c.order=do.call("c",rc1.list)

allcell=data.frame(cell=rownames(heatmap_matrix)[c.order],cellcluster=names(c.order))
allcell$cellcluster2=gsub("[.].*","",allcell$cellcluster)
allcell$newcluster=allcell$cellcluster2

heatmap_matrix1=heatmap_matrix[allcell$cell,]
heatmapresult=Heatmap(as.matrix(heatmap_matrix1), name = "scaledata", cluster_rows = T, 
                      cluster_columns = F, 
                      top_annotation = anno_col, use_raster=T,
                      # column_split = annot_df_col$type,
                      row_split= allcell$newcluster,
                      # left_annotation = anno_row,
                      show_column_dend = F, show_row_dend = F, 
                      clustering_distance_rows='euclidean', show_row_names = F,show_column_names = F,
                      # col = c(colorRampPalette(c("grey60", "white"))(10), colorRampPalette(c(RColorBrewer::brewer.pal(9, "YlOrRd")))(60)),
                      col=hmcols,
                      column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                      row_km = 1)
tiff(paste0(dataname0,"heatmap2.tiff"),width = 1000,height = 800,res = 150)#
h1.1=draw(heatmapresult)
dev.off()

r.dend=column_dend(h1.1)
rc1.list=row_order(h1.1)
names(rc1.list)=paste0(names(rc1.list),".")

c.order=do.call("c",rc1.list)
allcell2=data.frame(cell=rownames(heatmap_matrix1)[c.order],cellcluster=names(c.order))
allcell2$cellcluster2=gsub("[.].*","",allcell2$cellcluster)
allcell2$cellcluster2=factor(allcell2$cellcluster2,levels = paste0("cluster",1:4))
heatmap_matrix2=heatmap_matrix1[allcell2$cell,]


heatmapresult=Heatmap(as.matrix(heatmap_matrix2), name = "scaledata", cluster_rows = F, 
                      cluster_columns = F, 
                      top_annotation = anno_col, use_raster=T,
                      # column_split = annot_df_col$type,
                      row_split= allcell2$cellcluster2,
                      # left_annotation = anno_row,
                      show_column_dend = F, show_row_dend = F, 
                      clustering_distance_rows='euclidean', show_row_names = F,show_column_names = F,
                      # col = c(colorRampPalette(c("grey60", "white"))(10), colorRampPalette(c(RColorBrewer::brewer.pal(9, "YlOrRd")))(60)),
                      col=hmcols,
                      column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                      row_km = 1)
tiff(paste0(dataname0,"heatmap3.tiff"),width = 1000,height = 800,res = 150)#
h1.1=draw(heatmapresult)
dev.off()
pdf(paste0(dataname0,"heatmap3.pdf"))
print(heatmapresult)
dev.off()
xlsx::write.xlsx(allcell2,file = paste0(dataname0,"pseudotime_gene_function.xlsx"),sheetName = "gene_cluster",append = T)

```

```{r}

heatmap_matrix=heatmap_matrix2
heatmap_matrix_agg=aggregate(heatmap_matrix,by=list(cluster=allcell2$cellcluster2),"mean")

df_melt=melt(heatmap_matrix_agg)
df_melt$Pseudotime=pData(HSMM)$Pseudotime[match(df_melt$variable,colnames(HSMM))]
ggplot(df_melt,aes(Pseudotime,value,col=cluster))+geom_point()
```
```{r}
colnames(df_melt)[c(3,4)]=c('y',"x")

line_intersection1.2 <- reconPlots::curve_intersect(df_melt[df_melt$cluster=="cluster1"&df_melt$y>0.5,], df_melt[df_melt$cluster=="cluster2"&df_melt$y>0.5,])
line_intersection2.3 <- reconPlots::curve_intersect(df_melt[df_melt$cluster=="cluster2"&df_melt$x>5&df_melt$x<7,], df_melt[df_melt$cluster=="cluster3"&df_melt$x>5&df_melt$x<7,])
line_intersection3.4 <- reconPlots::curve_intersect(df_melt[df_melt$cluster=="cluster3"&df_melt$x>6,], df_melt[df_melt$cluster=="cluster4"&df_melt$x>6,])
labelxy=data.frame(x=c(line_intersection1.2$x,line_intersection2.3$x,line_intersection3.4$x),y=c(line_intersection1.2$y,line_intersection2.3$y,line_intersection3.4$y))
labelxy$label=paste0(signif(labelxy$x,3),", ",signif(labelxy$y,3))
p1=ggplot(mapping=aes(x,y))+geom_line(data=df_melt,mapping=aes(col=cluster))+
  geom_vline(xintercept = labelxy$x)+
  geom_text(data=labelxy,mapping=aes(x=x+1.5,label=label))#+theme_classic()
p1.1=ggplot(mapping=aes(x,y))+geom_point(data=df_melt,mapping=aes(col=cluster))+
  geom_vline(xintercept = labelxy$x)+
  geom_text(data=labelxy,mapping=aes(x=x+1.5,label=label))#+theme_classic()
p1.2=ggplot(df_melt,aes(x,y,col=cluster))+geom_line()+facet_wrap(~cluster,nrow = length(unique(df_melt$cluster)))
p1.3=ggplot(df_melt,aes(x,y,col=cluster))+geom_point()+facet_wrap(~cluster,nrow = length(unique(df_melt$cluster)))
xrange=c(range(pData(HSMM)$Pseudotime)[1],labelxy$x,range(pData(HSMM)$Pseudotime)[2])
xpercent=(xrange[-1]-xrange[-length(xrange)])/xrange[length(xrange)]
xpercent_cum=cumsum(xpercent)
pData(HSMM)$stage=sapply(pData(HSMM)$Pseudotime,function(x){if(x<xrange[2]){return("s1")}else if(x>=xrange[2]&x<xrange[3]){return("s2")}else if(x>=xrange[3]&x<xrange[4]){return("s3")}else if(x>=xrange[4]){return("s4")}else{return("NULL")}})
pData(HSMM)$stage=sapply(pData(HSMM)$Pseudotime,function(x){if(x<2.17){return("s1")}else if(x>=2.17&x<5.11){return("s2")}else if(x>=5.11&x<7.25){return("s3")}else if(x>=7.25){return("s4")}else{return("NULL")}})
seuratData@meta.data$stage=pData(HSMM)$stage[match(colnames(seuratData),colnames(HSMM))]
seuratData@meta.data$Pseudotime=pData(HSMM)$Pseudotime[match(colnames(seuratData),colnames(HSMM))]

data=as.data.frame(table(seuratData@meta.data[,c("Phase","stage")]))
data=data%>%group_by(stage)%>%mutate(percent=Freq/sum(Freq))
p2=ggplot(data,aes(stage,Freq,fill=Phase))+geom_bar(stat="identity")
p3=ggplot(data,aes(stage,percent,fill=Phase))+geom_bar(stat="identity")
p4=ggplot(data)+geom_line(aes(x=stage,y=percent,group=Phase,col=Phase))+scale_y_continuous(limits = c(0.2,0.5))


pdf(paste0(dataname0,"gene_stage_pseudotime_intersecting_point.pdf"),width = 15)
print(p1)
print(p1.1)
print(p1.2)
print(p1.3)

dev.off()
```

```{r}
seuratData$month=seuratData$orig
data=as.data.frame(table(seuratData@meta.data[,c("month","stage")]))
data=data%>%group_by(month)%>%mutate(percent=Freq/sum(Freq))
p1=ggplot(data)+geom_bar(aes(x=month,y=percent,group=month,fill=stage),stat = "identity")



library(RColorBrewer)
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
#处理后有73种差异还比较明显的颜色，基本够用
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))) 

data=data.frame(table(seuratData@meta.data[,c("stage","month")]))

data=data%>%group_by(stage)%>%mutate(perc=Freq/sum(Freq))
data=data.frame(data)
p2=ggplot(data,aes(stage,perc,fill=month))+geom_bar(stat = "identity")+theme_classic()+scale_fill_manual(values = col_vector )

pdf("stage_month_dis.pdf")
print(p1)
print(p2)
dev.off()


```

# function
```{r}
library(clusterProfiler)
library(org.Mm.eg.db)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(ggplot2)
load("~/mmu_gene2id.RData")
gene_cluster=xlsx::read.xlsx("F:/Edata/yaobingqing/20220922_GF/20221010pseudotime_gene_function.xlsx",sheetIndex = 1)
stage_gene_list=lapply(paste0("cluster",1:4),function(x){return(gene_cluster$cell[gene_cluster$cellcluster2==x])})
names(stage_gene_list)=paste0("s",1:4)

enrichfunciton=list()
for(stage in names(stage_gene_list)){
  print(stage)
  genes=stage_gene_list[[stage]]
  geneID=gene2id$ENTREZID[match(genes,gene2id$SYMBOL,nomatch=0)]
  ego=enrichGO(gene = geneID, OrgDb = org.Mm.eg.db, ont = "BP", pAdjustMethod = "BH", qvalueCutoff = 0.05, readable = TRUE)
  ekegg <-enrichKEGG(gene = geneID, organism = 'mmu', pAdjustMethod = "BH", qvalueCutoff = 0.05)
  if(!is.null(ego)){
    ego=ego@result[ego@result$pvalue<0.05,]
    enrichfunciton[[stage]][["ego"]]=ego
    }
  if(!is.null(ekegg)){
    ekegg=ekegg@result[ekegg@result$pvalue<0.05,]
    ekegg$genesymbol=sapply(ekegg$geneID,function(x){a=unlist(strsplit(x,"/"));b=gene2id$SYMBOL[match(a,gene2id$ENTREZID,nomatch = 0)];return(paste0(b,collapse = "/"))})
    enrichfunciton[[stage]][["ekegg"]]=ekegg
  }
}
save(enrichfunciton,file = "enrichfunction_stage.RData")

t=do.call("c",enrichfunciton)
writexl::write_xlsx(t,path =paste0("20221010_stage_enrichfunction",".xlsx") )
```

#show function pattern
```{r}
load("F:/Edata/yaobingqing/20220922_GF/3sample/20220929_pseudotime/20220929_neutrophilseuratData.HSMM.RData")
seuratData=seuratData_neutrophil
HSMM=HSMM_neutrophil

stage_function=list()
for(i in paste0("s",1:4)){
  print(i)
  t=xlsx::read.xlsx("F:/Edata/yaobingqing/20220922_GF/3sample/20220929_pseudotime/20221010_stage_enrichfunction_for_figure.xlsx",sheetName = paste0(i,".ekegg"))
  if(i%in%c("s1","s4")){t=t[t$p.adjust<0.05,]}
  stage_function[[i]]=t
}
p=unique(do.call("c",lapply(stage_function,function(x){x$Description})))
stage_function$con_WT=xlsx::read.xlsx("F:/Edata/yaobingqing/20220922_GF/3sample/20220929_pseudotime/20221010_stage_enrichfunction_for_figure.xlsx",sheetName = paste0("control_vs_ATBX.high",".ekegg"))

load("E:/databases/pathway_genesets/02_kegg_API_download/kegg_api_geneSets20221019.RData")
kegg=mmu_kegg_list_des
names(kegg)=gsub(" - Mus musculus \\(mouse\\)","",names(kegg))
gs=kegg[match(p,names(kegg),nomatch = 0)]

s=AddModuleScore(seuratData,features = gs,seed = 100)
d=s@meta.data[,match("Cluster1",colnames(s@meta.data)):ncol(s@meta.data)]
colnames(d)=names(gs)

  data=t(d)
  data=data[,colnames(HSMM)]

  feature_sheet=data.frame(row.names = rownames(data),gene_short_name=rownames(data))
  pd <- new("AnnotatedDataFrame", data = pData(HSMM))
  fd <- new("AnnotatedDataFrame", data =feature_sheet )


  tmp=newCellDataSet(data,
                     phenoData = pd,
                     featureData = fd,
                     lowerDetectionLimit=-10,
                     expressionFamily=uninormal())

  annotation_col=data.frame(row.names =rownames(pData(HSMM)),month=pData(HSMM)$month,State=pData(HSMM)$State,Phase=pData(HSMM)$Phase)

  tmp=tmp[,colnames(HSMM)]
  #dendsort重新排序
  heatmap_matrix=plot_pseudotime_heatmap2_return_heatmap_matrix(tmp,cluster_rows = T,
                                                                add_annotation_col = annotation_col,
                                                                return_heatmap = T,hclust_method = 'ward.D2',
                                                                num_clusters=1)
  save(heatmap_matrix,file = "neutrophil_pathway_heatmap_matrix.RData")
  load("neutrophil_pathway_heatmap_matrix.RData")
data=as.data.frame(t(heatmap_matrix))
  
  annotation_col=data.frame(row.names =rownames(pData(HSMM)),month=pData(HSMM)$month,State=pData(HSMM)$State,Phase=pData(HSMM)$Phase)
  
  num=length(unique(pData(HSMM)$month))
  num2=length(unique(pData(HSMM)$State))
  num3=length(unique(pData(HSMM)$Phase))
  anno_col=list(month=rainbow(num)[1:num],State=rainbow(num2)[1:num2],Phase=rainbow(num3)[1:num3])
  names(anno_col$month)=levels(pData(HSMM)$month)
  names(anno_col$State)=unique(pData(HSMM)$State)
  names(anno_col$Phase)=levels(pData(HSMM)$Phase)
  
  annot_df_col=pData(HSMM)[,c("month","Phase","Pseudotime")]
  
  column_color2=rainbow(length(unique(annot_df_col$month)))
  names(column_color2)=levels(annot_df_col$month)
  column_color3=rainbow(length(unique(annot_df_col$Phase)))
  names(column_color3)=unique(annot_df_col$Phase)
  
  anno_col=HeatmapAnnotation(df=annot_df_col,col=list(month=column_color2,Phase=column_color3,Pseudotime=circlize::colorRamp2(c(0, max(annot_df_col$Pseudotime)), c("white","darkgreen"))))
  heatmapresult=Heatmap(as.matrix(heatmap_matrix), name = "scaledata", 
                        cluster_rows = T, 
                        cluster_columns = F, 
                        top_annotation = anno_col, use_raster=T,
                        # column_split = annot_df_col$type,
                        # row_split= max_ind,
                        # left_annotation = anno_row,
                        show_column_dend = F, show_row_dend = F, 
                        clustering_distance_rows='euclidean', show_row_names = T,show_column_names = F,
                        # col=hmcols,
                        #column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                        row_km = 1)
max_ind=apply(heatmap_matrix,1,function(x){a=order(x,decreasing = T)[1:20];xcol=ncol(heatmap_matrix);
  s1=sum(sapply(a,function(y){y<0.15*xcol}))
  s2=sum(sapply(a,function(y){y>=0.15*xcol&y<0.5*xcol}))
  s3=sum(sapply(a,function(y){y>=0.5*xcol&y<0.85*xcol}))
  s4=sum(sapply(a,function(y){y>=0.85*xcol}))
  return(paste0("cluster",which.max(c(s1,s2,s3,s4))))
  })
  max_ind=factor(max_ind,levels = paste0("cluster",1:4))
  heatmapresult=Heatmap(as.matrix(heatmap_matrix), name = "scaledata", 
                        cluster_rows = T, 
                        cluster_columns = F, 
                        top_annotation = anno_col, use_raster=T,
                        # column_split = annot_df_col$type,
                        row_split= max_ind,
                        # left_annotation = anno_row,
                        show_column_dend = F, show_row_dend = F, 
                        clustering_distance_rows='euclidean', show_row_names = T,show_column_names = F,
                        # col=hmcols,
                        #column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                        row_km = 1)
  tiff(paste0(dataname0,"heatmap.tiff"))#,width = 1000,height = 800,res = 150
  h1.1=draw(heatmapresult)
  dev.off()
  
  r.dend=column_dend(h1.1)
  rc1.list=row_order(h1.1)
  # names(rc1.list)=paste0("cluster",1:15,".")
  names(rc1.list)=paste0(names(rc1.list),".")
  
  c.order=do.call("c",rc1.list)
  
  allcell=data.frame(cell=rownames(heatmap_matrix)[c.order],cellcluster=names(c.order))
  allcell$cellcluster2=gsub("[.].*","",allcell$cellcluster)
  allcell$newcluster=allcell$cellcluster2
  
  heatmap_matrix1=heatmap_matrix[allcell$cell,]
  heatmapresult=Heatmap(as.matrix(heatmap_matrix1), name = "scaledata", cluster_rows = T, 
                        cluster_columns = F, 
                        top_annotation = anno_col, use_raster=T,
                        # column_split = annot_df_col$type,
                        row_split= allcell$newcluster,
                        # left_annotation = anno_row,
                        show_column_dend = F, show_row_dend = F, 
                        clustering_distance_rows='euclidean', show_row_names = T,show_column_names = F,
                        # col = c(colorRampPalette(c("grey60", "white"))(10), colorRampPalette(c(RColorBrewer::brewer.pal(9, "YlOrRd")))(60)),
                        # col=hmcols,
                        column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                        row_km = 1)
  tiff(paste0(dataname0,"heatmap2.tiff"),width = 1000,height = 800,res = 150)#
  h1.1=draw(heatmapresult)
  dev.off()
  
  r.dend=column_dend(h1.1)
  rc1.list=row_order(h1.1)
  names(rc1.list)=paste0(names(rc1.list),".")
  
  c.order=do.call("c",rc1.list)
  allcell2=data.frame(cell=rownames(heatmap_matrix1)[c.order],cellcluster=names(c.order))
  allcell2$cellcluster2=gsub("[.].*","",allcell2$cellcluster)
  allcell2$cellcluster2=factor(allcell2$cellcluster2,levels = paste0("cluster",1:4))
  heatmap_matrix2=heatmap_matrix1[allcell2$cell,]
  
  heatmapresult=Heatmap(as.matrix(heatmap_matrix2), name = "scaledata", cluster_rows = F, 
                        cluster_columns = F, 
                        top_annotation = anno_col, use_raster=T,
                        # column_split = annot_df_col$type,
                        row_split= allcell2$cellcluster2,
                        # left_annotation = anno_row,
                        show_column_dend = F, show_row_dend = F, 
                        clustering_distance_rows='euclidean', show_row_names = T,show_column_names = F,
                        # col = c(colorRampPalette(c("grey60", "white"))(10), colorRampPalette(c(RColorBrewer::brewer.pal(9, "YlOrRd")))(60)),
                        # col=hmcols,
                        column_dend_height = unit(40, "mm"),row_gap = unit(0, "mm"),
                        row_names_gp = gpar(fontsize = c(10)),
                        row_km = 1)
  tiff(paste0(dataname0,"heatmap3.tiff"),width = 1000,height = 800,res = 150)#
  h1.1=draw(heatmapresult)
  dev.off()
  pdf(paste0(dataname0,"heatmap3-function.pdf"))
  print(heatmapresult)
  dev.off()

pdf("20221105_pathway_score_vln.pdf")  
for(i in rownames(heatmap_matrix2)){
  print(i)
  d=heatmap_matrix2[i,]
  d=cbind(exp=d,seuratData@meta.data[names(d),])
  d$stage=factor(d$stage,levels = paste0("s",1:4))
  p1=ggplot(d,aes(orig,exp,col=orig))+geom_violin()+geom_boxplot(width=.1,outlier.colour = NA)+
    ggpubr::stat_compare_means(comparisons = as.list(data.frame(combn(levels(d$orig),2))))+theme_classic()+
    labs(title = i)+geom_hline(yintercept = median(d$exp[d$orig=="control"]))+
    theme(plot.title = element_text(hjust = 0.5))
  p2=ggplot(d,aes(stage,exp,col=stage))+geom_violin()+geom_boxplot(width=.1,outlier.colour = NA)+
    ggpubr::stat_compare_means(comparisons = as.list(data.frame(combn(levels(d$stage),2))))+theme_classic()+
    # labs(title = i)+geom_hline(yintercept = median(d$exp[d$orig=="control"]))+
    theme(plot.title = element_text(hjust = 0.5))
    print(p1+p2)
    
}
dev.off()
```

```{r}

#qPCR genes

allcell2=xlsx::read.xlsx("F:/Edata/yaobingqing/20220922_GF/3sample/20220929_pseudotime/20221010pseudotime_gene_function.xlsx",sheetIndex = 1)
diff_test_res$stage=allcell2$cellcluster2[match(diff_test_res$gene_short_name,allcell2$cell)]
write.table(diff_test_res,file = "tmp",col.names = T,row.names = F,quote = F,sep = "\t")


genes=diff_test_res$gene_short_name
genes=sort(genes)

p3=VlnPlot(seuratData,features = c(genes),ncol = 10,group.by = "orig")
tiff(paste0("neutrophil_pseudotime_gene.orig.tiff"),res = 300,width = 10000,height = 1000*ceiling(length(genes)/10))
print(p3)
dev.off()

p3=VlnPlot(seuratData,features = c(genes),ncol = 10,group.by = "stage")
tiff(paste0("neutrophil_pseudotime_gene.stage.tiff"),res = 300,width = 10000,height = 1000*ceiling(length(genes)/10))
print(p3)
dev.off()


```

## rm branches
```{r}
load("F:/Edata/yaobingqing/20220922_GF/3sample/20220929_pseudotime/20220929_neutrophilseuratData.HSMM.RData")
HSMM=HSMM_neutrophil[,pData(HSMM_neutrophil)$State%in%c(1,3,5,7)]
HSMM@reducedDimS=HSMM@reducedDimS[,colnames(HSMM)]
plot_cell_trajectory(HSMM,show_tree = F,show_backbone = F,show_branch_points = F)

seuratData=seuratData_neutrophil[,colnames(HSMM)]
DimPlot(seuratData,group.by = "stage")


```

#WT & IBC.001.PDX1.P0
```{r}
load("WT_PDX.myeloid.neutrophil.RData")
markerlist=list()
for(celltype in c(unique(seuratData$celltype))){
  print(celltype)
  s0=seuratData[,seuratData$celltype==celltype]
  Idents(s0)="orig"
  f=FindMarkers(s0,ident.1 = "IBC.001.PDX1.P0",ident.2 = "WT",logfc.threshold=0.1,only.pos=F)
  # f=f[f$p_val<0.05,]
  markerlist[[celltype]]=f
  
}

enrichfunciton=list()
for(celltype in names(markerlist)){
  print(celltype)
  g0=markerlist[[celltype]]
  g0=g0[g0$p_val<0.05,]
  for(j in c("high","low")){
    print(j)
    genes=rownames(g0)[g0$avg_log2FC>0]
    if(j=="low"){
      genes=rownames(g0)[g0$avg_log2FC<0]
    }
    
    geneID=gene2id$ENTREZID[match(genes,gene2id$SYMBOL,nomatch=0)]
    ego=enrichGO(gene = geneID, OrgDb = org.Mm.eg.db, ont = "BP", pAdjustMethod = "BH", qvalueCutoff = 0.05, readable = TRUE)
    ekegg <-enrichKEGG(gene = geneID, organism = 'mmu', pAdjustMethod = "BH", qvalueCutoff = 0.05)
    
    if(!is.null(ego)){
      ego=ego@result[ego@result$pvalue<0.05,]
      enrichfunciton[[celltype]][[j]][["ego"]]=ego
    }
    if(!is.null(ekegg)){
      ekegg=ekegg@result[ekegg@result$pvalue<0.05,]
      ekegg$genesymbol=sapply(ekegg$geneID,function(x){a=unlist(strsplit(x,"/"));b=gene2id$SYMBOL[match(a,gene2id$ENTREZID,nomatch = 0)];return(paste0(b,collapse = "/"))})
      enrichfunciton[[celltype]][[j]][["ekegg"]]=ekegg
      
    }
  }
  
}
markerlist$neutrophil$gene=rownames(markerlist$neutrophil)
t=c(markerlist,do.call("c",lapply(enrichfunciton,function(x){do.call("c",x)})))
writexl::write_xlsx(t,path = "function.xlsx")
gsea_list=list()
for(i in names(markerlist)){
  print(i)
  m=markerlist[[i]]
  genes=m[order(m$avg_log2FC,decreasing = T),]
  geneList=genes$avg_log2FC
  names(geneList)=rownames(genes)
  gsea= GSEA(geneList, TERM2GENE = pathwaylist)
  gsea_list[[i]]=gsea
  
}
xlsx::write.xlsx(gsea_list$neutrophil@result,file = "function.xlsx",append = T,sheetName = "gsea")


```

```{r}
load("WT_PDX.myeloid.neutrophil.RData")
neutrophil_genelist=lapply(neutrophil_genelist,first2big)
s=seuratData
neutrophil_genelist=lapply(neutrophil_genelist,function(x){x[x%in%rownames(s)]})
s=AddModuleScore(s,features = neutrophil_genelist,seed = 100)
colnames(s@meta.data)[match("Cluster1",colnames(s@meta.data)):ncol(s@meta.data)]=names(neutrophil_genelist)
pdf("neutrophil_score-WT_PDX.myeloid.neutrophil.pdf")
for(genelist in names(neutrophil_genelist)){
  print(genelist)
  d=s@meta.data[,c("orig",genelist)]
  colnames(d)[ncol(d)]="exp"
  p1=ggplot(d,aes(orig,exp,col=orig))+geom_violin()+geom_boxplot(width=.1,outlier.colour = NA)+
    ggpubr::stat_compare_means(comparisons = as.list(data.frame(combn(levels(d$orig),2))))+theme_classic()+
    labs(title = genelist)+geom_hline(yintercept = median(d$exp[d$orig=="WT"]))
  print(p1)
}
dev.off()


```

#survival BRCA TCGA&metabric
```{r}

library(survminer)
library(survival)
load("E:/publicdata/TCGA_BRCA/metabric_TCGA_BRCA.exp.clinical.RData")
load("E:/databases/mm10_geneloc/20210422_human2mice.RData")#human2mice gene symbol conversion

load("~/human.gene2id.RData")
load("E:/YaoBingQing/20200709_singlecell/markerlist_bacVScon.RData")
bacmarker=lapply(markerlist,function(x){x[x$p_val_adj<0.05,]})
bacmarker=lapply(bacmarker,function(x){toupper(rownames(x))})
load("F:/Edata/yaobingqing/neutrophil_genelist.RData")
IL17Bsignaling=c("IL17B","IL17RB","ACT1","TRAF3IP2","TRAF6","NFKB1","RELA","MAP3K7")

gs=c(neutrophil_genelist,bacmarker,list(IL17Bsignaling=IL17Bsignaling))
gs=lapply(gs,toupper)

surv_df=matrix(0,ncol = 6)
colnames(surv_df)=c("cancertype","pathway","eventype","pval","tenratehigh","tenratelow")

pdf("BRCA_subtype_survival.pdf")
for(geneset in names(gs)){
  gene=gs[[geneset]]
  print(geneset)
  for(db in names(survival_datalist$exp)){
    # print(db)
    expdata=survival_datalist$exp[[db]]
    clinical=survival_datalist$clinical[[db]]
    genetmp=gene[gene%in%rownames(expdata)]
    clinical$score=apply(expdata[genetmp,],2,mean)
    clinical=clinical[clinical$time>0,]
    
    res.cut=surv_cutpoint(clinical,time = "time",event="event",variables = "score")
    res.cat=surv_categorize(res.cut)
    fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score")), data = res.cat)
    p1=ggsurvplot(
      fit,                     # survfit object with calculated statistics.
      data = res.cat,             # data used to fit survival curves.
      risk.table = TRUE,       # show risk table.
      pval = TRUE,             # show p-value of log-rank test.
      # conf.int = TRUE,         # show confidence intervals for 
      # point estimates of survival curves.
      # xlim = x_limit,        # present narrower X axis, but not affect
      # survival estimates.
      xlab = "Time in years",   # customize X axis label.
      # break.time.by = break.time.by,     # break X axis in time intervals by 500.
      # ggtheme = theme_light(), # customize plot and risk table with a theme.
      # risk.table.y.text.col = T, # colour risk table text annotations.
      risk.table.y.text = FALSE # show bars instead of names in text annotations
      # in legend of risk table
    )+labs(title=paste(geneset,db,sep = "_"))
    
    # print(p1)
    
    p=unicox(surv=res.cat)
    a=p1$data.survplot[p1$data.survplot$score=="high",]
    tenratehigh=a$surv[which.min(abs(a$time-10))]
    a=p1$data.survplot[p1$data.survplot$score=="low",]
    tenratelow=a$surv[which.min(abs(a$time-10))]
    # alleventp=c()
    surv_df=rbind(surv_df,c(db,geneset,"OS",p$pvalue,signif(tenratehigh,3),signif(tenratelow,3)))
    
    for(subtype in levels(clinical$PAM50)){
      t=clinical[clinical$PAM50%in%subtype,]
      res.cut=surv_cutpoint(t,time = "time",event="event",variables = "score")
      res.cat=surv_categorize(res.cut)
      fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score")), data = res.cat)
      p1=ggsurvplot(
        fit,                     # survfit object with calculated statistics.
        data = res.cat,             # data used to fit survival curves.
        risk.table = TRUE,       # show risk table.
        pval = TRUE,             # show p-value of log-rank test.
        # conf.int = TRUE,         # show confidence intervals for 
        # point estimates of survival curves.
        # xlim = x_limit,        # present narrower X axis, but not affect
        # survival estimates.
        xlab = "Time in years",   # customize X axis label.
        # break.time.by = break.time.by,     # break X axis in time intervals by 500.
        # ggtheme = theme_light(), # customize plot and risk table with a theme.
        # risk.table.y.text.col = T, # colour risk table text annotations.
        risk.table.y.text = FALSE # show bars instead of names in text annotations
        # in legend of risk table
      )+labs(title=paste(geneset,db,subtype,sep = "_"))
      
      # print(p1)
      
      p=unicox(surv=res.cat)
      a=p1$data.survplot[p1$data.survplot$score=="high",]
      tenratehigh=a$surv[which.min(abs(a$time-10))]
      a=p1$data.survplot[p1$data.survplot$score=="low",]
      tenratelow=a$surv[which.min(abs(a$time-10))]
      # alleventp=c()
      surv_df=rbind(surv_df,c(paste0(db,"_",subtype),geneset,"OS",p$pvalue,signif(tenratehigh,3),signif(tenratelow,3)))
      
    }
  }
  
}
dev.off()


surv_df=surv_df[-1,]
surv_df=as.data.frame(surv_df)
surv_df$pval=as.numeric(surv_df$pval)
surv_df$tenratehigh=as.numeric(surv_df$tenratehigh)
surv_df$tenratelow=as.numeric(surv_df$tenratelow)
surv_df$highlow=surv_df$tenratehigh>surv_df$tenratelow

t=surv_df
t=t%>%filter(pval<0.05)
t$pathway=factor(t$pathway,levels = names(gs))
t$pathway=droplevels(t$pathway)
p=ggplot(t,aes(pathway,cancertype))+
  geom_point(aes(size=-log(pval),color=highlow))+
  # scale_color_gradient2(low = "blue",high = "red",midpoint = -log(0.05))+
  facet_wrap(~eventype,scales = "free")+
  theme_classic()+#labs(title=gsub("surv_df_|.RData","",i))+
  theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0.5))
pdf("BRCA_surv_dot.pdf")
print(p)
dev.off()

library(ggcorrplot) 
library(psych)
score_list=list()
for(db in names(survival_datalist$exp)){
  print(db)
  expdata=survival_datalist$exp[[db]]
  clinical=clinical=survival_datalist$clinical[[db]]
  enrich.mat <- GSVA::gsva(expr = as.matrix(expdata), gset.idx.list = gs, kcdf = "Gaussian",method = "ssgsea", verbose = TRUE)
  clinical=cbind(clinical,t(enrich.mat))
  score_list[[db]]=clinical
}
save(score_list,file = "score_list.RData")

pdf("geneset_correlation_total.pdf")
for(db in names(score_list)){
  
  d=score_list[[db]]
  t1=c("Lac","Sc","Sk","Staph","Ef")
  t2=setdiff(names(gs),t1)
  
  # cors <- cor(d[,t1],d[,t2])
  t=corr.test(d[,t1],d[,t2])
  t$r[t$p.adj>0.05]=NA
  p=ggcorrplot(t$r)+labs(title = db)
  print(p)
  
  for(subtype in levels(d$PAM50)){
    
    t=corr.test(d[d$PAM50%in%subtype,t1],d[d$PAM50%in%subtype,t2])
    t$r[t$p.adj>0.05]=NA
    p=ggcorrplot(t$r)+labs(title = paste(db,subtype))
    print(p)
    
  }
}
dev.off()

t1=c("Lac","Sc","Sk","Staph","Ef")
t2=setdiff(names(gs),t1)
pdf("geneset_correlation_single.pdf")
for(db in names(score_list)){
  
  d=score_list[[db]]
  
  for(gs1 in t1){
    for(gs2 in t2){
      t=d[,c("PAM50",gs1,gs2)]
      colnames(t)[2:3]=c("gs1","gs2")
      p1=ggplot(t,aes(gs1,gs2))+geom_point()+theme_classic()+
        ggpubr::stat_cor(label.x.npc = "left")+
        stat_smooth(se=F,method="lm",size=1.2)+
        ggpubr::stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")),label.x.npc = "middle")+
        theme_bw()+theme(plot.title = element_text(hjust = 0.5),
                         axis.title.x = element_blank(),panel.border = element_blank(),
                         panel.grid = element_blank(),axis.line = element_line(size=1,colour = 'black'))+
        scale_colour_brewer(palette = "Set1")+labs(title = paste(db,gs1,gs2))
      p2=ggplot(t[!is.na(t$PAM50),],aes(gs1,gs2))+geom_point()+theme_classic()+facet_wrap(~PAM50,scales = "free")+
        ggpubr::stat_cor(label.x.npc = "left")+
        stat_smooth(se=F,method="lm",size=1.2)+
        ggpubr::stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")),label.x.npc = "middle")+
        theme_bw()+theme(plot.title = element_text(hjust = 0.5),
                         axis.title.x = element_blank(),panel.border = element_blank(),
                         panel.grid = element_blank(),axis.line = element_line(size=1,colour = 'black'))+
        scale_colour_brewer(palette = "Set1")+labs(title = paste(db,gs1,gs2))
      print(p1)
      print(p2)
    }
  }
}
dev.off()


pdf("singlegene_correlation_total.pdf")
for(db in names(survival_datalist$exp)){
  
  d=survival_datalist$exp[[db]]
  clinical=survival_datalist$clinical[[db]]
  t1=c(rownames(d)[grep("^IRF|^STING",rownames(d))])
  t2=rownames(d)[grep("^IL17",rownames(d))]
  
  d=t(d[c(t1,t2),])
  d=cbind(clinical,d)
  # cors <- cor(d[,t1],d[,t2])
  t=corr.test(d[,t1],d[,t2])
  t$r[t$p>0.05]=NA
  p=ggcorrplot(t$r)+labs(title = db)
  print(p)
  
  for(subtype in levels(d$PAM50)){
    
    t=corr.test(d[d$PAM50%in%subtype,t1],d[d$PAM50%in%subtype,t2])
    t$r[t$p>0.05]=NA
    p=ggcorrplot(t$r)+labs(title = paste(db,subtype))
    print(p)
    
  }
}
dev.off()

pdf("singlegene_correlation_single.pdf")
for(db in names(survival_datalist$exp)){
  
  d=survival_datalist$exp[[db]]
  clinical=survival_datalist$clinical[[db]]
  t1=c(rownames(d)[grep("^IRF|^STING",rownames(d))])
  t2=rownames(d)[grep("^IL17",rownames(d))]
  
  d=t(d[c(t1,t2),])
  d=cbind(clinical,d)
  
  
  for(gs1 in t1){
    for(gs2 in t2){
      t=d[,c("PAM50",gs1,gs2)]
      colnames(t)[2:3]=c("gs1","gs2")
      p1=ggplot(t,aes(gs1,gs2))+geom_point()+theme_classic()+
        ggpubr::stat_cor(label.x.npc = "left")+
        stat_smooth(se=F,method="lm",size=1.2)+
        ggpubr::stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")),label.x.npc = "middle")+
        theme_bw()+theme(plot.title = element_text(hjust = 0.5),
                         axis.title.x = element_blank(),panel.border = element_blank(),
                         panel.grid = element_blank(),axis.line = element_line(size=1,colour = 'black'))+
        scale_colour_brewer(palette = "Set1")+labs(title = paste(db,gs1,gs2))
      p2=ggplot(t[!is.na(t$PAM50),],aes(gs1,gs2))+geom_point()+theme_classic()+facet_wrap(~PAM50,scales = "free")+
        ggpubr::stat_cor(label.x.npc = "left")+
        stat_smooth(se=F,method="lm",size=1.2)+
        ggpubr::stat_regline_equation(aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~")),label.x.npc = "middle")+
        theme_bw()+theme(plot.title = element_text(hjust = 0.5),
                         axis.title.x = element_blank(),panel.border = element_blank(),
                         panel.grid = element_blank(),axis.line = element_line(size=1,colour = 'black'))+
        scale_colour_brewer(palette = "Set1")+labs(title = paste(db,gs1,gs2))
      print(p1)
      print(p2)
    }
  }
}
dev.off()



```

```{r}
genes=c("IL17B","STING1","IRF3")
gs=as.list(genes)
names(gs)=paste0(genes,"singlegene")
surv_df=matrix(0,ncol = 6)
colnames(surv_df)=c("cancertype","pathway","eventype","pval","tenratehigh","tenratelow")

pdf("singlegene_survival.pdf")
for(geneset in names(gs)){
  gene=gs[[geneset]]
  print(geneset)
  for(db in names(survival_datalist$exp)){
    # print(db)
    expdata=survival_datalist$exp[[db]]
    clinical=survival_datalist$clinical[[db]]
    genetmp=gene[gene%in%rownames(expdata)]
    if(length(genetmp)==0){next()}
    clinical$score=apply(expdata[genetmp,,drop=F],2,mean)
    clinical=clinical[clinical$time>0,]
    
    res.cut=surv_cutpoint(clinical,time = "time",event="event",variables = "score")
    res.cat=surv_categorize(res.cut)
    fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score")), data = res.cat)
    p1=ggsurvplot(
      fit,                     # survfit object with calculated statistics.
      data = res.cat,             # data used to fit survival curves.
      risk.table = TRUE,       # show risk table.
      pval = TRUE,             # show p-value of log-rank test.
      # conf.int = TRUE,         # show confidence intervals for 
      # point estimates of survival curves.
      # xlim = x_limit,        # present narrower X axis, but not affect
      # survival estimates.
      xlab = "Time in years",   # customize X axis label.
      # break.time.by = break.time.by,     # break X axis in time intervals by 500.
      # ggtheme = theme_light(), # customize plot and risk table with a theme.
      # risk.table.y.text.col = T, # colour risk table text annotations.
      risk.table.y.text = FALSE # show bars instead of names in text annotations
      # in legend of risk table
    )+labs(title=paste(geneset,db,sep = "_"))
    
    print(p1)
    
    p=unicox(surv=res.cat)
    a=p1$data.survplot[p1$data.survplot$score=="high",]
    tenratehigh=a$surv[which.min(abs(a$time-10))]
    a=p1$data.survplot[p1$data.survplot$score=="low",]
    tenratelow=a$surv[which.min(abs(a$time-10))]
    # alleventp=c()
    surv_df=rbind(surv_df,c(db,geneset,"OS",p$pvalue,signif(tenratehigh,3),signif(tenratelow,3)))
    
    for(subtype in levels(clinical$PAM50)){
      t=clinical[clinical$PAM50%in%subtype,]
      res.cut=surv_cutpoint(t,time = "time",event="event",variables = "score")
      res.cat=surv_categorize(res.cut)
      fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score")), data = res.cat)
      p1=ggsurvplot(
        fit,                     # survfit object with calculated statistics.
        data = res.cat,             # data used to fit survival curves.
        risk.table = TRUE,       # show risk table.
        pval = TRUE,             # show p-value of log-rank test.
        # conf.int = TRUE,         # show confidence intervals for 
        # point estimates of survival curves.
        # xlim = x_limit,        # present narrower X axis, but not affect
        # survival estimates.
        xlab = "Time in years",   # customize X axis label.
        # break.time.by = break.time.by,     # break X axis in time intervals by 500.
        # ggtheme = theme_light(), # customize plot and risk table with a theme.
        # risk.table.y.text.col = T, # colour risk table text annotations.
        risk.table.y.text = FALSE # show bars instead of names in text annotations
        # in legend of risk table
      )+labs(title=paste(geneset,db,subtype,sep = "_"))
      
      print(p1)
      
      p=unicox(surv=res.cat)
      a=p1$data.survplot[p1$data.survplot$score=="high",]
      tenratehigh=a$surv[which.min(abs(a$time-10))]
      a=p1$data.survplot[p1$data.survplot$score=="low",]
      tenratelow=a$surv[which.min(abs(a$time-10))]
      # alleventp=c()
      surv_df=rbind(surv_df,c(paste0(db,"_",subtype),geneset,"OS",p$pvalue,signif(tenratehigh,3),signif(tenratelow,3)))
      
    }
  }
  
}
dev.off()



```
#singlegene_survival_hormone.pdf singlegene_survival_therapy.pdf
```{r}
pdf("singlegene_survival_therapy.pdf")
for(geneset in names(gs)){
  gene=gs[[geneset]]
  print(geneset)
  for(db in names(survival_datalist$exp)[2]){
    # print(db)
    expdata=survival_datalist$exp[[db]]
    clinical=survival_datalist$clinical[[db]]
    genetmp=gene[gene%in%rownames(expdata)]
    if(length(genetmp)==0){next()}
    clinical$score=apply(expdata[genetmp,,drop=F],2,mean)
    clinical=clinical[clinical$time>0,]
    
    res.cut=surv_cutpoint(clinical,time = "time",event="event",variables = "score")
    res.cat=surv_categorize(res.cut)
    
    for(i in c("Chemotherapy","Hormone.Therapy","Radio.Therapy")){
      print(i)
      res.cat$status=clinical[match(rownames(res.cat),clinical$ID),i]
      fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score+status")), data = res.cat)
      p1=ggsurvplot(
        fit,                     # survfit object with calculated statistics.
        data = res.cat,             # data used to fit survival curves.
        risk.table = TRUE,       # show risk table.
        pval = TRUE,             # show p-value of log-rank test.
        xlab = "Time in years",   # customize X axis label.
        risk.table.y.text = FALSE # show bars instead of names in text annotations
      )+labs(title=paste(geneset,db,i,sep = "_"))
      
      print(p1)
    }
    
    
    for(subtype in levels(clinical$PAM50)){
      t=clinical[clinical$PAM50%in%subtype,]
      res.cut=surv_cutpoint(t,time = "time",event="event",variables = "score")
      res.cat=surv_categorize(res.cut)
      
      for(i in c("Chemotherapy","Hormone.Therapy","Radio.Therapy")){
        print(i)
        res.cat$status=clinical[match(rownames(res.cat),clinical$ID),i]
        fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score+status")), data = res.cat)
        p1=ggsurvplot(
          fit,                     # survfit object with calculated statistics.
          data = res.cat,             # data used to fit survival curves.
          risk.table = TRUE,       # show risk table.
          pval = TRUE,             # show p-value of log-rank test.
          xlab = "Time in years",   # customize X axis label.
          risk.table.y.text = FALSE # show bars instead of names in text annotations
        )+labs(title=paste(geneset,db,subtype,i,sep = "_"))
        
        print(p1)
      }
    }
  }
}
dev.off()

pdf("singlegene_survival_hormone.pdf")
for(geneset in names(gs)){
  gene=gs[[geneset]]
  print(geneset)
  for(db in names(survival_datalist$exp)[2]){
    # print(db)
    expdata=survival_datalist$exp[[db]]
    clinical=survival_datalist$clinical[[db]]
    genetmp=gene[gene%in%rownames(expdata)]
    if(length(genetmp)==0){next()}
    clinical$score=apply(expdata[genetmp,,drop=F],2,mean)
    clinical=clinical[clinical$time>0,]
    
    res.cut=surv_cutpoint(clinical,time = "time",event="event",variables = "score")
    res.cat=surv_categorize(res.cut)
    
    for(i in c("ER.Status","HER2.Status","PR.Status")){
      print(i)
      res.cat$status=clinical[match(rownames(res.cat),clinical$ID),i]
      fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score+status")), data = res.cat)
      p1=ggsurvplot(
        fit,                     # survfit object with calculated statistics.
        data = res.cat,             # data used to fit survival curves.
        risk.table = TRUE,       # show risk table.
        pval = TRUE,             # show p-value of log-rank test.
        xlab = "Time in years",   # customize X axis label.
        risk.table.y.text = FALSE # show bars instead of names in text annotations
      )+labs(title=paste(geneset,db,i,sep = "_"))
      
      print(p1)
    }
    
    
    for(subtype in levels(clinical$PAM50)){
      t=clinical[clinical$PAM50%in%subtype,]
      res.cut=surv_cutpoint(t,time = "time",event="event",variables = "score")
      res.cat=surv_categorize(res.cut)
      
      for(i in c("ER.Status","HER2.Status","PR.Status")){
        print(i)
        res.cat$status=clinical[match(rownames(res.cat),clinical$ID),i]
        fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score+status")), data = res.cat)
        p1=ggsurvplot(
          fit,                     # survfit object with calculated statistics.
          data = res.cat,             # data used to fit survival curves.
          risk.table = TRUE,       # show risk table.
          pval = TRUE,             # show p-value of log-rank test.
          xlab = "Time in years",   # customize X axis label.
          risk.table.y.text = FALSE # show bars instead of names in text annotations
        )+labs(title=paste(geneset,db,subtype,i,sep = "_"))
        
        print(p1)
      }
    }
  }
}
dev.off()


```
#lancet_2005_replase_surv.pdf
```{r}
library(survminer)

lancet_2005=read.table("E:/course/paper/from20200601/survival_prognosis_computing_nature.NEJM/2005.lancet.286patient/exp.txt",
                       header = T,sep = "\t",quote = "")
clinical0=read.table("E:/course/paper/from20200601/survival_prognosis_computing_nature.NEJM/2005.lancet.286patient/clinical.txt",
                     header = T,sep = "\t",quote = "")
GPL96=read.table("E:/SHJ/20200417_array_bone_metastasis/series_matrix/GPL96-57554.txt",header = T,as.is = T,sep = "\t",quote = "")
GPL96=GPL96[GPL96$Gene.Symbol!="",]
GPL96=GPL96[!grepl("[/]",GPL96$Gene.Symbol),]
exp=aggregate(lancet_2005[,2:ncol(lancet_2005)],by=list(gene=GPL96$Gene.Symbol[match(lancet_2005$ID_REF,GPL96$ID)]),max)
rownames(exp)=exp$gene
exp=exp[,-1]
exp=log(exp)
exp=exp[,clinical0$GEO_asscession_number]

pdf("lancet_2005_replase_surv.pdf")
for(gene in c(gs)){#,"CD200","PCSK9","SERPINB9","SLC6A1","BCL11B","VDR","SLC6A2",'TSC22D1'
  print(gene)
  expdata=exp
  clinical=clinical0
  genetmp=gene[gene%in%rownames(expdata)]
  if(!gene%in%rownames(expdata)){next()}
  
  clinical$score=apply(expdata[genetmp,,drop=F],2,mean)
  
  res.cut=surv_cutpoint(clinical,time = "time_to_relapse_or_last_follow.up_months",event="relapse_1_True",variables = "score")
  res.cat=surv_categorize(res.cut)
  fit=survfit(as.formula(paste0('Surv(time_to_relapse_or_last_follow.up_months, relapse_1_True=="1") ~',"score")), data = res.cat)
  p1=ggsurvplot(
    fit,                     # survfit object with calculated statistics.
    data = res.cat,             # data used to fit survival curves.
    risk.table = TRUE,       # show risk table.
    pval = TRUE,             # show p-value of log-rank test.
    xlab = "Time in years",   # customize X axis label.
    risk.table.y.text = FALSE # show bars instead of names in text annotations
  )+labs(title=paste(gene,sep = "_"))
  
  print(p1)
  
  for(type in c("ER-","ER+")){
    t=clinical[clinical$ER_Status%in%type,]
    res.cut=surv_cutpoint(t,time = "time_to_relapse_or_last_follow.up_months",event="relapse_1_True",variables = "score")
    res.cat=surv_categorize(res.cut)
    fit=survfit(as.formula(paste0('Surv(time_to_relapse_or_last_follow.up_months, relapse_1_True=="1") ~',"score")), data = res.cat)
    p1=ggsurvplot(
      fit,                     # survfit object with calculated statistics.
      data = res.cat,             # data used to fit survival curves.
      risk.table = TRUE,       # show risk table.
      pval = TRUE,             # show p-value of log-rank test.
      xlab = "Time in years",   # customize X axis label.
      risk.table.y.text = FALSE # show bars instead of names in text annotations
    )+labs(title=paste(gene,type,sep = "_"))
    
    print(p1)
  }
}
dev.off()


```
#mskcc_bone_MFS_surv_singlegene.pdf
```{r}
library(survminer)

mskcc=read.table("E:/course/paper/from20200601/survival_prognosis_computing_nature.NEJM/2005nature_breast_cancer_metastasis_to_lung/41586_2005_BFnature03799_MOESM11_ESM.txt",
                 header = T,sep = "\t")
clinical=read.table("E:/course/paper/from20200601/survival_prognosis_computing_nature.NEJM/2005nature_breast_cancer_metastasis_to_lung/MSKCC_clinical.txt",
                    header = T,sep="\t",quote = "")
clinical$Sample_Name=gsub("[-]",".",clinical$Sample_Name)
GPL96=read.table("E:/SHJ/20200417_array_bone_metastasis/series_matrix/GPL96-57554.txt",header = T,as.is = T,sep = "\t",quote = "")
GPL96=GPL96[GPL96$Gene.Symbol!="",]
GPL96=GPL96[!grepl("[/]",GPL96$Gene.Symbol),]
exp=aggregate(mskcc[,2:ncol(mskcc)],by=list(gene=GPL96$Gene.Symbol[match(mskcc$ID_REF,GPL96$ID)]),max)
rownames(exp)=exp$gene
exp=exp[,-1]
exp=log(exp)
exp=exp[,clinical$Sample_Name]




pdf("mskcc_bone_MFS_surv_singlegene.pdf")
for(gene in c(gs)){#,"CD200","PCSK9","SERPINB9","SLC6A1","BCL11B","VDR","SLC6A2",'TSC22D1'
  print(gene)
  # expdata=survival_datalist$exp[[db]]
  expdata=exp
  # clinical=survival_datalist$clinical[[db]]
  clinical=clinical
  genetmp=gene[gene%in%rownames(expdata)]
  if(!gene%in%rownames(expdata)){next()}
  # genetmp=TF_target_list_to_hs$encode[[gene]][TF_target_list_to_hs$encode[[gene]]%in%rownames(expdata)]
  
  clinical$score=apply(expdata[genetmp,,drop=F],2,mean)
  
  res.cut=surv_cutpoint(clinical,time = "MFS_YR",event="Met_Event",variables = "score")
  res.cat=surv_categorize(res.cut)
  fit=survfit(as.formula(paste0('Surv(MFS_YR, Met_Event=="1") ~',"score")), data = res.cat)
  p1=ggsurvplot(
    fit,                     # survfit object with calculated statistics.
    data = res.cat,             # data used to fit survival curves.
    risk.table = TRUE,       # show risk table.
    pval = TRUE,             # show p-value of log-rank test.
    xlab = "Time in years",   # customize X axis label.
    risk.table.y.text = FALSE # show bars instead of names in text annotations
  )+labs(title=paste(gene,sep = "_"))
  
  print(p1)
}
dev.off()


```
#brca_mbcproject_wagle_2017.pdf
```{r}
brca_mbcproject_wagle_2017=read.table("E:/publicdata/TCGA_BRCA/cBioPortalData/brca_mbcproject_wagle_2017/data_mrna_seq_v2_rsem.txt",header = T,as.is = T,sep = "\t")
brca_mbcproject_wagle_2017=brca_mbcproject_wagle_2017[!duplicated(brca_mbcproject_wagle_2017$Hugo_Symbol),]
rownames(brca_mbcproject_wagle_2017)=brca_mbcproject_wagle_2017$Hugo_Symbol
colnames(brca_mbcproject_wagle_2017)=gsub("[.]","-",colnames(brca_mbcproject_wagle_2017))
brca_mbcproject_wagle_2017=brca_mbcproject_wagle_2017[,c(-1,-2)]

clinical_brca_mbcproject_wagle_2017=read.table("E:/publicdata/TCGA_BRCA/cBioPortalData/brca_mbcproject_wagle_2017/data_clinical_sample.txt",header = T,as.is = T,quote = '',sep = '\t')
clinical_brca_mbcproject_wagle_2017p=read.table("E:/publicdata/TCGA_BRCA/cBioPortalData/brca_mbcproject_wagle_2017/data_clinical_patient.txt",header = T,as.is = T,quote = '',sep = '\t')
clinical_brca_mbcproject_wagle_2017=clinical_brca_mbcproject_wagle_2017[,c("SAMPLE_ID","PATIENT_ID","SAMPLE_TIMEPOINT","BX_TIME_DAYS","BX_LOCATION","BX_SIDEDNESS_BREAST","BX_ER","BX_PR","BX_HER2OVERALL","TMB_NONSYNONYMOUS")]
clinical_brca_mbcproject_wagle_2017p=clinical_brca_mbcproject_wagle_2017p[,c("PATIENT_ID","AGE_AT_DIAGNOSIS_GROUP","CALC_TIME_TO_METS_DX_DAYS","CALC_TIME_TO_METS_DX_MONTHS","DX_ER","DX_PR","DX_HER2OVERALL","METS_SITES_AT_DX","EVER_METS_SITES")]
clinical_brca_mbcproject_wagle_2017$time2met=clinical_brca_mbcproject_wagle_2017p$CALC_TIME_TO_METS_DX_DAYS[match(clinical_brca_mbcproject_wagle_2017$PATIENT_ID,clinical_brca_mbcproject_wagle_2017p$PATIENT_ID)]
clinical=clinical_brca_mbcproject_wagle_2017
t1=clinical[clinical$BX_LOCATION!="BREAST",]

clinical$time2met[(clinical$PATIENT_ID%in%t1$PATIENT_ID)&(clinical$BX_LOCATION=="BREAST")]=0
colnames(clinical)[match("time2met",colnames(clinical))]="time"

t=intersect(colnames(brca_mbcproject_wagle_2017),clinical$SAMPLE_ID)
expdata=brca_mbcproject_wagle_2017[,t]
clinical=clinical[match(t,clinical$SAMPLE_ID),]
clinical$event=clinical$BX_LOCATION
clinical$event[clinical$event!="BREAST"]="MET"
clinical$event=plyr::mapvalues(clinical$event,from = c("BREAST","MET"),to=c("0","1"))
clinical$event=as.numeric(clinical$event)

pdf("brca_mbcproject_wagle_2017.pdf")
for(gene in c("IL17B","STING1","IRF3")){
  
  genetmp=gene[gene%in%rownames(expdata)]
  if(length(genetmp)==0){next()}
  clinical$score=apply(expdata[genetmp,,drop=F],2,mean)
  
  # clinical=clinical[!is.na(clinical$time),]
  # clinical=clinical[clinical$time>0,]
  
  res.cut=surv_cutpoint(clinical,time = "time",event="event",variables = "score")
  res.cat=surv_categorize(res.cut)
  fit=survfit(as.formula(paste0('Surv(time, event) ~',"score")), data = res.cat)
  p1=ggsurvplot(
    fit,                     # survfit object with calculated statistics.
    data = res.cat,             # data used to fit survival curves.
    risk.table = TRUE,       # show risk table.
    pval = TRUE,             # show p-value of log-rank test.
    xlab = "Time in years",   # customize X axis label.
    risk.table.y.text = FALSE # show bars instead of names in text annotations
  )+labs(title = gene)
  print(p1)
}
dev.off()

```

##scRNA & SHJ & TCGA xcell neutrophil
```{r}
data=xlsx::read.xlsx("E:/YaoBingQing/20200709_singlecell/enriched_pathway/singlecell_enriched_pathway.xlsx",sheetName = "kegg_high")
data=data[data$Cluster=="Staph",]
data$Description=factor(data$Description,levels = data$Description)
p=ggplot(data,aes(-log10(p.adjust),Description,fill=-log10(p.adjust)))+geom_bar(stat = "identity")+theme_classic()+
  scale_fill_continuous(high="red",low="blue")
pdf("staph_kegg.pdf")
print(p)
dev.off()

load("E:/databases/pathway_genesets/02_kegg_API_download/kegg_api_geneSets20221019.RData")
match(levels(data$Description),names(mmu_kegg_list_des))

gs=mmu_kegg_list_des[names(mmu_kegg_list_des)%in%data$Description]
t=match(names(gs),levels(data$Description))
gs=gs[order(t)]
gs=lapply(gs,function(x){x[x%in%rownames(seuratData)]})
s=AddModuleScore(seuratData,features = gs,seed = 100)
colnames(s@meta.data)[match("Cluster1",colnames(s@meta.data)):ncol(s@meta.data)]=names(gs)
s=s[,s$orig%in%c("Con2","Staph")]
pdf("neutrophil_score-WT_PDX.myeloid.neutrophil.pdf")
for(genelist in names(gs)){
  print(genelist)
  d=s@meta.data[,c("orig",genelist)]
  colnames(d)[ncol(d)]="exp"
  p1=ggplot(d,aes(orig,exp,col=orig))+geom_violin()+geom_boxplot(width=.1,outlier.colour = NA)+
    ggpubr::stat_compare_means(comparisons = as.list(data.frame(combn(unique(as.character(d$orig)),2))))+
    theme_classic()+
    labs(title = genelist)+geom_hline(yintercept = median(d$exp[d$orig=="WT"]))
  print(p1)
}
dev.off()

p1=VlnPlot(s,c(rownames(seuratData)[grep("Il17",rownames(seuratData))]),group.by = "orig")

d=data.frame(t(s@assays$RNA@data[c(rownames(seuratData)[grep("Il17",rownames(seuratData))]),]),orig=s$orig)
df=reshape2::melt(d)
head(df)
p2=ggplot(df,aes(orig,value,fill=orig))+geom_boxplot()+theme_classic()+
  facet_wrap(~variable)+ggpubr::stat_compare_means(comparisons = list(c("Con2","Staph")))
pdf("scRNA.IL17.pdf")
p1
p2
dev.off()

load("E:/SHJ/20200406_mm10_10x/clusterRdata/20201027_seuratData_tumor_tmp.RData")
s=seuratData_tumor_tmp
p1=VlnPlot(s,features = c(rownames(seuratData_tumor_tmp)[grep("Il17",rownames(seuratData_tumor_tmp))]),group.by = "orig",pt.size = 0)
d=data.frame(t(s@assays$RNA@data[c(rownames(seuratData_tumor_tmp)[grep("Il17",rownames(seuratData_tumor_tmp))]),]),orig=s$orig)
df=reshape2::melt(d)
head(df)
p2=ggplot(df,aes(orig,value,fill=orig))+geom_boxplot(outlier.colour = NA)+theme_classic()+
  facet_wrap(~variable)+ggpubr::stat_compare_means(comparisons = list(c("Primary1","Lung1")))
d$expIl17b=d$Il17b>0.001
table(d[,c("orig","expIl17b")])
d=as.data.frame(table(d[,c("orig","expIl17b")]))
d=d%>%group_by(orig)%>%mutate(perc=Freq/sum(Freq))
p3=ggplot(d,aes(orig,perc,fill=expIl17b))+geom_bar(stat = "identity")+theme_classic()+
  geom_text(aes(label=signif(perc,digits = 3)))+labs(title = "Il17b")

pdf("SHJ.tumor.IL17.pdf")
p1
p2
p3
dev.off()

d=as.data.frame(table(seuratData@meta.data[,c("orig","celltype")]))
d=d%>%group_by(orig)%>%mutate(perc=Freq/sum(Freq))
d$perc=signif(d$perc,digits = 3)
d$orig_celltype=paste0(d$orig,"_",d$celltype)
d$orig_celltype=factor(d$orig_celltype,levels = paste0(rep(c("Primary1","Lung1"),each=9),"_",unique(d$celltype)))
pdf("SHJ.celltype.perc.pdf")
ggplot(d,aes(orig,perc,fill=celltype))+geom_bar(stat = "identity")+theme_classic()
ggplot(d,aes(orig,perc,fill=celltype))+geom_bar(stat = "identity",position = "dodge")+theme_classic()
ggplot(d,aes(orig_celltype,perc,fill=celltype))+geom_bar(stat="identity")+theme_classic()+
  geom_text(aes(label=perc))+
  theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0.5))
dev.off()


load("F:/Edata/publicdata/agingdata/code/TCGA_fpkm_list.RData")
load("~/human.gene2id.RData")
library(immunedeconv)
a=gene2id$SYMBOL[match(substr(rownames(TCGA_fpkm_list2$STAD),1,15),gene2id$ENSEMBL)]
t1=c(which(is.na(a)),which(a%in%""))
a=a[-t1]
for(i in names(TCGA_fpkm_list2)){
  print(i)
  exp=TCGA_fpkm_list2[[i]]
  exp=exp[-t1,]
  rownames(exp)=a
  TCGA_fpkm_list2[[i]]=exp
}

mcp_list=list()
for(i in "BRCA"){
  print(i)
  exp=TCGA_fpkm_list2[[i]]
  res_mcp = as.data.frame(deconvolute(exp, "xcell"))#mcp_counter,xcell,epic
  mcp_list[[i]]=res_mcp
  
}
pdf("celltype_IL17_cor.pdf")
for(i in "BRCA"){
  print(i)
  mcp=mcp_list[[i]]
  mcp$cell_type=gsub(" ","_",mcp$cell_type)
  rownames(mcp)=mcp$cell_type
  mcp=mcp[,-1]
  exp=TCGA_fpkm_list2[[i]]
  genes=rownames(exp)[grep("^IL17|^IRF|^STING",rownames(exp))]
  genes=sort(genes)
  a=substr(colnames(mcp),14,15)
  
  t=psych::corr.test(t(exp[genes,a%in%c(paste0("0",1:9))]),t(mcp[,a%in%c(paste0("0",1:9))]))
  t$r[t$p>0.05]=NA
  p=ggcorrplot::ggcorrplot(t$r)+labs(title = i)
  print(p)
  
  
}
dev.off()

survival_data=openxlsx::read.xlsx("F:/Edata/publicdata/agingdata/code/survival_data_all.xlsx")
survival_data$OS.time=survival_data$OS.time/365.25
survival_data$DSS.time=survival_data$DSS.time/365.25
survival_data$DFI.time=survival_data$DFI.time/365.25
survival_data$PFI.time=survival_data$PFI.time/365.25
eventype=list(OS=c("OS","OS.time"),DSS=c("DSS","DSS.time"),DFI=c("DFI","DFI.time"),PFI=c("PFI","PFI.time"))

pdf("neutrophi.survival.pdf")
for(cancertype in c("BRCA")){
  print(cancertype)
  exp=TCGA_fpkm_list2[[cancertype]]
  exp=exp[,substr(colnames(exp),14,15)%in%c(paste0("0",1:9))]#01 06
  cancer=substr(colnames(exp),1,12)
  exp=exp[,!duplicated(cancer)] ###rm duplicates
  colnames(exp)=substr(colnames(exp),1,12)
  
  s=survival_data[survival_data$type==cancertype,]
  t=intersect(colnames(exp),s$bcr_patient_barcode)
  s0=s[match(t,s$bcr_patient_barcode),]
  
  mcp=mcp_list[[cancertype]]
  mcp$cell_type=gsub(" ","_",mcp$cell_type)
  rownames(mcp)=mcp$cell_type
  mcp=mcp[,-1]
  s0$neutrophil=as.numeric(mcp["Neutrophil",match(s0$bcr_patient_barcode,substr(colnames(mcp),1,12))])
  for(et in names(eventype)){
    print(et)
    s1=s0[,c(eventype[[et]],"neutrophil")]
    colnames(s1)[1:2]=c("event","time")
    
    if(sum(is.na(s1$event))/nrow(s1)>0.5){next()}
    
    
    # s1$score="high"
    # s1$score[t0<median(t0)]="low"
    
    s1$score=s1$neutrophil
    # s1=s1[!is.na(s1$time),]
    # s1=s1[s1$time>0,]
    
    res.cut=surv_cutpoint(s1,time = "time",event="event",variables = "score")
    s1=surv_categorize(res.cut)
    
    
    p=unicox(surv=s1)
    p1=p$plot+labs(title = paste0(cancertype,"-","-",et))
    
    print(p1)
    
    
  } 
  
}
dev.off()

```
## Lai guan Ng
```{r}
Laiguan=list(
  T3=c("Atf3","Ccl3","Ccl4","Cd274","Cstb","Cxcl3","Hcar2","Hilpda","Hk2","Hmox1","Ier3","Jun","Plin2","Spp1","Tgif1","Tnfrsf23","Vegfa","Zeb2","Ldha","Mif"),
  T2=c("Cd300ld","Cxcr2","Dusp1","Gbp2","Ifitm1","Il1b","Isg15","Jaml","Junb","Msrb1","Osm","S100a6","Selplg","Slpi"),
  T1=c("Ltc4s","Mmp8","Mmp9","Ppia","Prr13","Ptma","Retnlg")
)
gs=Laiguan
s=seuratData[,seuratData$celltype=="neutrophil"]
s=AddModuleScore(s,features = gs,seed = 100)

data=s@meta.data
colnames(data)[match("Cluster1",colnames(data)):ncol(data)]=c(names(gs))


pdf("Laiguan_geneset_exp.in_GF.pdf")
d=aggregate(data[,c(names(gs))],by=list(data$orig),FUN=median)
rownames(d)=d[,1]
d=d[,-1]
# d=scale(d)
# Heatmap(d)
# d=t(scale(t(d)))
Heatmap(d)
for(i in names(gs)){
  print(i)
  d=data[,c("orig",i)]
  colnames(d)[2]="exp"
  p=ggplot(d,aes(orig,exp,fill=orig))+geom_violin()+geom_boxplot(width=.1,outlier.colour = NA)+
    ggpubr::stat_compare_means(comparisons = as.list(data.frame(combn(levels(d$orig),2))))+theme_classic()+
    labs(title = i)+geom_hline(yintercept = median(d$exp[d$orig=="control"]))+
    theme(plot.title = element_text(hjust = 0.5))
  print(p)
}
dev.off()




```