---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(factoextra)
library(FactoMineR)
library(Rtsne)
library(DESeq2)
library(edgeR)
mm10_gene_length=read.table("E:/databases/hg38_geneloc/hg38_geneloc_rmdp_exonlength",as.is = T,header = T)
mm10_gene_length$ENSG=substr(mm10_gene_length$ENSG,1,15)
getwd()
```
#combine htseqcount result
```{r}
dir0="X:/analyze_data/FEdata/yaobingqing/20250324_HumanNeu/countdata/cellwithneutrophil/"
gzfile=list.files(path = dir0,pattern = 'gene.counts')
countdata=list()
for(i in gzfile){
  countdata[[gsub('.gene.counts','',i)]]=read.table(paste0(dir0,i),header = F,row.names=1)
}

dir0="X:/analyze_data/FEdata/yaobingqing/20250324_HumanNeu/countdata/supernatantwithneutrophil/"
gzfile=list.files(path = dir0,pattern = 'gene.counts')
for(i in gzfile){
  countdata[[gsub('.gene.counts','',i)]]=read.table(paste0(dir0,i),header = F,row.names=1)
}


data0=do.call("cbind",countdata)
colnames(data0)=names(countdata)

df=data.frame(expgene=rownames(data0))
df$symbol=mm10_gene_length$symbol[match(df$expgene,mm10_gene_length$ENSG)]
df=df[df$expgene%in%mm10_gene_length$ENSG,]
data0=data0[df$expgene,]


rownames(data0)=df$symbol

fileinfo=read.table("X:/analyze_data/FEdata/yaobingqing/20250324_HumanNeu/sampleinfo.txt",header = T)
sampleinfo=fileinfo[match(colnames(data0),fileinfo$sample),]
sampleinfo$LiveDead=plyr::mapvalues(sampleinfo$LiveDead,from = c("pure_L","pure_D","D","L"),to=c("pureLive","pureDead","Dead","Live"))
sampleinfo$group=paste0(sampleinfo$withTumor,".",sampleinfo$LiveDead)
sampleinfo$group=factor(sampleinfo$group)

data00=data0
sampleinfo0=sampleinfo




savelist=list()
for(cellline in c("cellwithneutrophil","supernatantwithneutrophil","supernatantbac")){
  print(cellline)
  
  if(cellline%in%c("cellwithneutrophil")){
    sampleinfo=sampleinfo0[sampleinfo0$group%in%c("T.Dead","T.Live","T.wo"),]
  }else if(cellline%in%c("supernatantbac")){
    sampleinfo=sampleinfo0[sampleinfo0$group%in%c("S.Dead","S.Live","S.wo"),]
  }else{
    sampleinfo=sampleinfo0[sampleinfo0$group%in%c("M.medium","S.Dead","S.Live","S.pureDead","S.pureLive","S.wo"),]
  }
  data0=data00[,sampleinfo$sample]
  
  sampleinfo$group=droplevels(sampleinfo$group)
  
  
  library(limma)
  library(Glimma)
  library(edgeR)
  library(Mus.musculus)
  
  
  data=data0
  savelist[[cellline]][["data0"]]=data0
  data[is.na(data)]=0
  x=DGEList(data)
  group=sampleinfo$group
  
  x$samples$group=group
  genesym=rownames(x)
  # genes=select(Mus.musculus,keys=genesym,columns = c("ENTREZID","TXCHROM"),keytype = "SYMBOL")
  # genes=genes[!duplicated(genes$SYMBOL),]
  # x$genes=genes
  
  # x=x[,x$samples$group%in%c("CTRL1","CMC","N12")]
  # group=x$samples$group
  cpm <- cpm(x)
  lcpm <- cpm(x, log=TRUE, prior.count=2)
  
  L <- mean(x$samples$lib.size) * 1e-6
  M <- median(x$samples$lib.size) * 1e-6
  c(L, M)
  
  table(rowSums(x$counts==0)==5)
  
  keep.exprs <- filterByExpr(x, group=group)
  x <- x[keep.exprs,, keep.lib.sizes=FALSE]
  
  lcpm.cutoff <- log2(10/M + 2/L)
  library(RColorBrewer)
  nsamples <- ncol(x)
  col <- brewer.pal(nsamples, "Paired")
  par(mfrow=c(1,2))
  plot(density(lcpm[,1]), col=col[1], lwd=2, ylim=c(0,0.26), las=2, main="", xlab="")
  title(main="A. Raw data", xlab="Log-cpm")
  abline(v=lcpm.cutoff, lty=3)
  for (i in 2:nsamples){
    den <- density(lcpm[,i])
    lines(den$x, den$y, col=col[i], lwd=2)
  }
  legend("topright", colnames(x), text.col=col, bty="n")
  lcpm <- cpm(x, log=TRUE,prior.count=2)
  plot(density(lcpm[,1]), col=col[1], lwd=2, ylim=c(0,0.26), las=2, main="", xlab="")
  title(main="B. Filtered data", xlab="Log-cpm")
  abline(v=lcpm.cutoff, lty=3)
  for (i in 2:nsamples){
    den <- density(lcpm[,i])
    lines(den$x, den$y, col=col[i], lwd=2)
  }
  legend("topright", colnames(x), text.col=col, bty="n")
  
  #归一化基因表达分布
  x <- calcNormFactors(x, method = "TMM")
  xexp=x
  save(xexp,file = "xexp.RData")
  lcpm <- cpm(x, log=TRUE,prior.count=2)
  pdf(paste0(cellline,".normalized.pdf"))
  boxplot(lcpm, las=2, col=col, main="")
  title(main="Normalised data",ylab="Log-cpm")
  
  dev.off()
  # save(lcpm,file = "lcpm.RData")
  
  lcpm <- cpm(x, log=TRUE)
  savelist[[cellline]][["lcpm"]]=lcpm
  col.group <- group
  levels(col.group) <-  brewer.pal(nlevels(col.group), "Set3")
  col.group <- as.character(col.group)
  plotMDS(lcpm, labels=group, col=col.group,pch = 21,gene.selection = "common")
  t=plotMDS(lcpm, labels=colnames(x), col=col.group,pch = 21,gene.selection = "common")
  title(main="Sample groups")
  
  xy = cbind(x=t$x,y=t$y, x$samples)
  xy$sample=rownames(xy)
  
  pdf(paste0(dataname0,"MDS.",cellline,".pdf"))
  p=ggplot(xy, aes(x=x,y=y))+geom_point(aes(colour = group ))+
    geom_text(aes(label=sample),size=3)+
    theme_classic()
  print(p)
  dev.off()
  
  ##差异分析
  design=model.matrix(~0+group)
  colnames(design)=gsub("group","",colnames(design))
  
  if(cellline=="cellwithneutrophil"){

    contr.matrix <- makeContrasts(
      T.Live_vs_T.Dead=T.Live-T.Dead,
      T.Live_vs_T.wo=T.Live-T.wo,
      T.Dead_vs_T.wo=T.Dead-T.wo,
      
      levels = colnames(design))
    
  }else if(cellline%in%c("supernatantbac")){
    
    contr.matrix <- makeContrasts(
      
      S.Live_vs_S.Dead=S.Live-S.Dead,
      S.Live_vs_S.wo=S.Live-S.wo,
      S.Dead_vs_S.wo=S.Dead-S.wo,
      levels = colnames(design))
  }else{
    
    contr.matrix <- makeContrasts(
      
      S.Live_vs_S.Dead=S.Live-S.Dead,
      S.Live_vs_S.wo=S.Live-S.wo,
      S.Dead_vs_S.wo=S.Dead-S.wo,
      S.pureLive_vs_S.pureDead=S.pureLive-S.pureDead,
      S.Live_vs_S.pureLive=S.Live-S.pureLive,
      S.Dead_vs_S.pureDead=S.Dead-S.pureDead,
      
      S.Live_vs_M.medium=S.Live-M.medium,
      S.Dead_vs_M.medium=S.Dead-M.medium,
      S.wo_vs_M.medium=S.wo-M.medium,
      S.pureLive_vs_M.medium=S.pureLive-M.medium,
      S.pureDead_vs_M.medium=S.pureDead-M.medium,
      
      
      levels = colnames(design))
    
  }
  
  # contr.matrix <- makeContrasts(
  #   CMCvsCTRL1=CMC-CTRL1,
  #   N12vsCTR1L=N12-CTRL1,
  #   levels = colnames(design))
  # par(mfrow=c(1,2))
  v <- voom(x, design, plot=TRUE)
  vfit <- lmFit(v, design)
  vfit <- contrasts.fit(vfit, contrasts=contr.matrix)
  efit <- eBayes(vfit)#topTable
  
  
  
  plotSA(efit, main="Final model: Mean-variance trend")
  
  tfit <- treat(vfit, lfc=1)
  dt <- decideTests(tfit,adjust.method = "none")
  summary(dt)
  
  
  DR_list=list()
  for(i in colnames(tfit$coefficients)){
    t=topTreat(tfit,coef = i,n=Inf)
    DR_list[[i]]=t
    # write.table(t,file = paste0(i,".txt"),sep = "\t",row.names = T,col.names = T,quote = F)
    # xlsx::write.xlsx(t,file = "20220913_diffgene.xlsx",sheetName = i,append = T)
  }
  DR_list=lapply(DR_list,function(x){x$gene=rownames(x);return(x)})
  
  savelist[[cellline]][["gene"]]=DR_list
  
  
}
save(savelist,file = "single.paired.comp.RData")


DR_list=lapply(savelist,function(x){x$gene})
# names(DR_list)=gsub("with.*","",names(DR_list))
names(DR_list)=c("","","bac")
DR_list=do.call("c",DR_list)


pdf(paste0(dataname0,"diffgene_valcano",".pdf"))
for(i in c(names(DR_list))){
  print(i)
  # d=data.frame(p=p[,i],fc=fc[,i],updown=updown[,i])
  t=DR_list[[i]]
  d=data.frame(p=t$P.Value,fc=t$logFC,gene=t$gene)
  
  d$updown=0
  d$updown[d$p<0.05&d$fc>0]=1
  d$updown[d$p<0.05&d$fc<0]=-1
  d$updown=factor(d$updown,levels = c(-1,0,1))
  # d2=d[-log10(d$p)>10|abs(d$fc)>8,]
  d2=d[d$p<0.05,]
  # d2=d2[d2$gene%in%rownames(updown)[updown[,i]!=0],]
  d2=d2[order(d2$fc),]
  d2=rbind(head(d2,10),tail(d2,10))
  t=ggplot(d,aes(fc,-log10(p),col=updown))+geom_point()+
    # geom_point(d2,col="green")+
    geom_text(data=d2,aes(label=gene),size=3)+
    scale_color_manual(values = c("blue","grey","red"))+labs(title=i)
  print(t)
}
dev.off()






markerlist=DR_list

load("~/human.gene2id.RData")


library(biomaRt)
library(clusterProfiler)
library(org.Hs.eg.db)
library(org.Mm.eg.db)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(ggplot2)

load("E:/databases/pathway_genesets/01_msigdb/20220428_msigdblist_human&mice.RData")#msigdb
load("E:/databases/pathway_genesets/02_kegg_API_download/kegg_api_geneSets20221019.RData") #hsa_kegg_list_des
# used_pathway=c(hsa_kegg_list_des,msigdblist$c2_01_all_canonical_pathway)
used_pathway=hsa_kegg_list_des
names(used_pathway)=gsub(" - Homo sapiens \\(human\\)","",names(used_pathway))
pathwaylist=data.frame(pathway=rep(names(used_pathway),sapply(used_pathway,length)),gene=do.call("c",used_pathway),stringsAsFactors = F)


enrichfunciton=list()
gsea_list=list()

for(celltype in names(markerlist)){
  print(celltype)
  
  m=markerlist[[celltype]]
  genes=m[order(m$logFC,decreasing = T),]
  geneList=genes$logFC
  names(geneList)=rownames(genes)
  gsea= GSEA(geneList, TERM2GENE = pathwaylist,pvalueCutoff = 1,pAdjustMethod = "BH")
  # j="IL-17 signaling pathway"
  # enrichplot::gseaplot2(gsea,geneSetID = j,title =j,color = 1,subplots = c(1,2,3),rel_heights = c(10,1,2),pvalue_table = T)
  
  
  gsea_list[[celltype]]=gsea
  
  m=m[m$P.Value<0.05,]
  
  for(j in c("high","low")){
    print(j)
    genes=rownames(m)[m$logFC>0]
    if(j=="low"){
      genes=rownames(m)[m$logFC<0]
    }
    
    geneID=gene2id$ENTREZID[match(genes,gene2id$SYMBOL,nomatch=0)]
    ego=enrichGO(gene = geneID, OrgDb = org.Hs.eg.db, ont = "BP", pAdjustMethod = "BH", qvalueCutoff = 0.05, readable = TRUE)
    ekegg <-enrichKEGG(gene = geneID, organism = 'hsa', pAdjustMethod = "BH", qvalueCutoff = 0.05)
    
    if(!is.null(ego)){
      ego=ego@result[ego@result$pvalue<0.05,]
      enrichfunciton[[celltype]][[j]][["ego"]]=ego
    }
    if(!is.null(ekegg)){
      ekegg=ekegg@result[ekegg@result$pvalue<0.05,]
      ekegg$genesymbol=sapply(ekegg$geneID,function(x){a=unlist(strsplit(x,"/"));b=gene2id$SYMBOL[match(a,gene2id$ENTREZID,nomatch = 0)];return(paste0(b,collapse = "/"))})
      enrichfunciton[[celltype]][[j]][["ekegg"]]=ekegg
      
    }
  }
}


functionlist=list(fun=enrichfunciton,gsea=gsea_list,gene=markerlist)
save(functionlist,file = "functionlist.RData")

t1=do.call("c",do.call("c",enrichfunciton))
t2=lapply(gsea_list,function(x){x@result})
names(t2)=paste0(names(t2),".gsea")
t3=markerlist
names(t3)=paste0(names(t3),".gene")
t3=lapply(t3,function(x){x$gene=rownames(x);return(x)})
t=c(t1,t2,t3)
names(t)=gsub("Live","L",names(t))
names(t)=gsub("Dead","D",names(t))

writexl::write_xlsx(t,path = "function.xlsx")




```


```{r}

savelist_nofilter=list()
for(cellline in c("cellwithneutrophil","supernatantwithneutrophil","supernatantbac")){
  print(cellline)
  
  if(cellline%in%c("cellwithneutrophil")){
    sampleinfo=sampleinfo0[sampleinfo0$group%in%c("T.Dead","T.Live","T.wo"),]
  }else if(cellline%in%c("supernatantbac")){
    sampleinfo=sampleinfo0[sampleinfo0$group%in%c("S.Dead","S.Live","S.wo"),]
  }else{
    sampleinfo=sampleinfo0[sampleinfo0$group%in%c("M.medium","S.Dead","S.Live","S.pureDead","S.pureLive","S.wo"),]
  }
  data0=data00[,sampleinfo$sample]
  
  sampleinfo$group=droplevels(sampleinfo$group)
  
  
  library(limma)
  library(Glimma)
  library(edgeR)
  library(Mus.musculus)
  
  
  data=data0
  savelist_nofilter[[cellline]][["data0"]]=data0
  data[is.na(data)]=0
  x=DGEList(data)
  group=sampleinfo$group
  
  x$samples$group=group
  genesym=rownames(x)
  # genes=select(Mus.musculus,keys=genesym,columns = c("ENTREZID","TXCHROM"),keytype = "SYMBOL")
  # genes=genes[!duplicated(genes$SYMBOL),]
  # x$genes=genes
  
  # x=x[,x$samples$group%in%c("CTRL1","CMC","N12")]
  # group=x$samples$group
  cpm <- cpm(x)
  lcpm <- cpm(x, log=TRUE, prior.count=2)
  
  L <- mean(x$samples$lib.size) * 1e-6
  M <- median(x$samples$lib.size) * 1e-6
  c(L, M)
  
  table(rowSums(x$counts==0)==5)
  
  # keep.exprs <- filterByExpr(x, group=group)
  # x <- x[keep.exprs,, keep.lib.sizes=FALSE]
  
  lcpm.cutoff <- log2(10/M + 2/L)
  library(RColorBrewer)
  nsamples <- ncol(x)
  col <- brewer.pal(nsamples, "Paired")
  par(mfrow=c(1,2))
  plot(density(lcpm[,1]), col=col[1], lwd=2, ylim=c(0,0.26), las=2, main="", xlab="")
  title(main="A. Raw data", xlab="Log-cpm")
  abline(v=lcpm.cutoff, lty=3)
  for (i in 2:nsamples){
    den <- density(lcpm[,i])
    lines(den$x, den$y, col=col[i], lwd=2)
  }
  legend("topright", colnames(x), text.col=col, bty="n")
  lcpm <- cpm(x, log=TRUE,prior.count=2)
  plot(density(lcpm[,1]), col=col[1], lwd=2, ylim=c(0,0.26), las=2, main="", xlab="")
  title(main="B. Filtered data", xlab="Log-cpm")
  abline(v=lcpm.cutoff, lty=3)
  for (i in 2:nsamples){
    den <- density(lcpm[,i])
    lines(den$x, den$y, col=col[i], lwd=2)
  }
  legend("topright", colnames(x), text.col=col, bty="n")
  
  #归一化基因表达分布
  x <- calcNormFactors(x, method = "TMM")
  xexp=x
  save(xexp,file = "xexp.RData")
  lcpm <- cpm(x, log=TRUE,prior.count=2)
  pdf(paste0(cellline,".normalized.pdf"))
  boxplot(lcpm, las=2, col=col, main="")
  title(main="Normalised data",ylab="Log-cpm")
  
  dev.off()
  # save(lcpm,file = "lcpm.RData")
  
  lcpm <- cpm(x, log=TRUE)
  savelist_nofilter[[cellline]][["lcpm"]]=lcpm
  
}
save(savelist_nofilter,file = "savelist_nofilter.RData")


```

```{r}


selectPath=list(GO=c("autophagy of nucleus","autophagosome assembly","autophagosome organization","response to starvation","endoplasmic reticulum unfolded protein response","autophagy of mitochondrion","regulation of G2/M transition of mitotic cell cycle","regulation of I-kappaB kinase/NF-kappaB signaling","replicative senescence","PERK-mediated unfolded protein response","vesicle coating","negative regulation of calcium ion transport into cytosol","vesicle budding from membrane","regulation of response to endoplasmic reticulum stress","regulation of lipid metabolic process","cholesterol homeostasis","immune response-regulating signaling pathway","regulation of lymphocyte differentiation","negative regulation of toll-like receptor signaling pathway","negative regulation of leukocyte activation","purine nucleoside monophosphate metabolic process","T cell apoptotic process","negative regulation of immune system process","negative regulation of cell activation","regulation of T cell activation"),
                gsea.immuneActivate=c("Yersinia infection","Epstein-Barr virus infection","Salmonella infection","Shigellosis","NOD-like receptor signaling pathway","Fc gamma R-mediated phagocytosis","Natural killer cell mediated cytotoxicity","Phagosome","Endocytosis","Toll-like receptor signaling pathway","Legionellosis","Neutrophil extracellular trap formation","Antigen processing and presentation","Staphylococcus aureus infection","Lysosome"),
                gsea.immuneSuppress=c("Pathways in cancer","Protein processing in endoplasmic reticulum","IL-17 signaling pathway","Apoptosis","JAK-STAT signaling pathway","Cellular senescence","T cell receptor signaling pathway","PD-L1 expression and PD-1 checkpoint pathway in cancer","HIF-1 signaling pathway","Ferroptosis"),
                gsea.tumorcell=c("Epithelial cell signaling in Helicobacter pylori infection","Estrogen signaling pathway","Bacterial invasion of epithelial cells","Breast cancer","Signaling pathways regulating pluripotency of stem cells")
                )
pathwaydf=data.frame(pathway=do.call("c",selectPath),type=rep(names(selectPath),sapply(selectPath,length)))
pathwaydf$pathway[pathwaydf$type%in%"GO"]=paste0("GO_",toupper(gsub(" ","_",pathwaydf$pathway[pathwaydf$type%in%"GO"])))

load("E:/databases/pathway_genesets/02_kegg_API_download/kegg_api_geneSets20221019.RData")
gs1=hsa_kegg_list_des[names(hsa_kegg_list_des)%in%do.call("c",selectPath[grep("gsea",names(selectPath))])]

load("E:/databases/pathway_genesets/01_msigdb/20220428_msigdblist_human&mice.RData")#msigdb
showpath=paste0("GO_",toupper(gsub(" ","_",selectPath$GO)))
showpath=showpath[!is.na(match(showpath,names(msigdblist$c5_06_GO_BP)))]
gs_GO=msigdblist$c5_06_GO_BP[showpath]

gs=c(gs1,gs_GO)

load("single.paired.comp.RData")


for(cellline in c("cellwithneutrophil","supernatantwithneutrophil")){
  print(cellline)
  
  data0=savelist[[cellline]]$data0
  
  
  enrich.mat <- GSVA::gsva(expr = as.matrix(data0), gset.idx.list = gs, kcdf = "Poisson",method = "gsva", verbose = TRUE)
  enrich.mat0=enrich.mat
  library(ComplexHeatmap)
  annot_df_col=sampleinfo0[match(colnames(data0),sampleinfo0$sample),c("LiveDead","group")]
  annot_df_col$group=droplevels(annot_df_col$group)
  column_color1=rainbow(length(levels(annot_df_col$group)))
  names(column_color1)=levels(annot_df_col$group)
  
  annot_df_row=pathwaydf[match(rownames(enrich.mat0),pathwaydf$pathway),"type",drop=F]
  # column_color2=rainbow(length(unique(annot_df_col$type2)))
  # names(column_color2)=unique(annot_df_col$type2)
  # column_color3=rainbow(length(unique(annot_df_col$mice)))
  # names(column_color3)=unique(annot_df_col$mice)
  
  anno_col=HeatmapAnnotation(df=annot_df_col,col=list(group=column_color1))
  anno_row=rowAnnotation(df=annot_df_row)
  
  t=enrich.mat0
  t=t(scale(t(t)))
  # t=scale(t)
  heatmap_matrix=Heatmap(as.matrix(t), #name = "scaledata", 
                         cluster_rows = T, 
                         cluster_columns = T, 
                         top_annotation = anno_col,# use_raster=T,
                         column_split = annot_df_col$group ,
                         left_annotation = anno_row,
                         row_split = annot_df_row$type,
                         # show_column_dend = F, show_row_dend = F, 
                         clustering_distance_rows='euclidean', show_row_names = T,show_column_names = T,
                         # col=hmcols,
                         #column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                         row_km = 1)
  pdf(paste0(dataname0,"pathway.activity.",cellline,".pdf"),width = 10,height = 10)
  print(heatmap_matrix)
  
  dev.off()
  
}

load("single.paired.comp.RData")
load("savelist_nofilter.RData")
selectGene=list(immuneSuppress=c(paste0("IL17R",LETTERS[1:5]),"IL6R-AS1","PIK3CD-AS1","TNFRSF8","ATG7","MAPKAPK3","CISH","IL10RB","GSN-AS1","TRAF1","ETS2","CD274","FAM83C-AS1","IL6R-AS1","BDH1","ERCC1","DDX21","TRAF1","ATG7","MVD","HIVEP1","LAMB3","CSF1","IL1B","IL4R","NOS2","CD84","CXCL1","CXCR1","CXCL2","CXCR2","CXCL8","TREM1","CD274","CYBB","PDCD1LG2","SLC27A2"),
                immuneActivate=c("TYROBP","ORAI3","TRPM2","PLD1","FLNA","GAB2","KIRREL2","SH2D3A","TLR2","NCF4","LCP2","TRAF1")
)

genedf=data.frame(gene=do.call("c",selectGene),type=rep(names(selectGene),sapply(selectGene,length)))
genedf=genedf[!duplicated(genedf$gene),]
for(cellline in c("cellwithneutrophil","supernatantwithneutrophil")){
  print(cellline)
  
  # data0=savelist[[cellline]]$lcpm
  data0=savelist_nofilter[[cellline]]$lcpm
  data0=data0[match(genedf$gene,rownames(data0),nomatch = 0),]
  
  library(ComplexHeatmap)
  annot_df_col=sampleinfo0[match(colnames(data0),sampleinfo0$sample),c("LiveDead","group")]
  annot_df_col$group=droplevels(annot_df_col$group)
  column_color1=rainbow(length(levels(annot_df_col$group)))
  names(column_color1)=levels(annot_df_col$group)
  
  annot_df_row=genedf[match(rownames(data0),genedf$gene),"type",drop=F]
  # column_color2=rainbow(length(unique(annot_df_col$type2)))
  # names(column_color2)=unique(annot_df_col$type2)
  # column_color3=rainbow(length(unique(annot_df_col$mice)))
  # names(column_color3)=unique(annot_df_col$mice)
  
  anno_col=HeatmapAnnotation(df=annot_df_col,col=list(group=column_color1))
  anno_row=rowAnnotation(df=annot_df_row)
  
  t=data0
  t=t(scale(t(t)))
  # t=scale(t)
  heatmap_matrix=Heatmap(as.matrix(t), #name = "scaledata", 
                         cluster_rows = T, 
                         cluster_columns = T, 
                         top_annotation = anno_col,# use_raster=T,
                         column_split = annot_df_col$group ,
                         left_annotation = anno_row,
                         row_split = annot_df_row$type,
                         # show_column_dend = F, show_row_dend = F, 
                         clustering_distance_rows='euclidean', show_row_names = T,show_column_names = T,
                         # col=hmcols,
                         #column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                         row_km = 1)
  pdf(paste0(dataname0,"gene.activity.nofilter.",cellline,".pdf"),width = 10,height = 10)
  print(heatmap_matrix)
  
  dev.off()
  
}


```

```{r}

showgenelist=list(tumorgene=c("EGFR","WNT7B","TGFA","MELK","MMP2","MMP9","CD44","ITGB1","BCL2L12","GLUT1","GLUT3","HIF1A","MKI67","CLDN3","CLDN4"),
                  Neugene=c("MPO","ELANE","CXCR1","CXCR2","S100A8","S100A9")
                  )
genedf=data.frame(gene=do.call("c",showgenelist),type=rep(names(showgenelist),sapply(showgenelist,length)))


load("X:/analyze_data/FEdata/yaobingqing/20250324_HumanNeu/01_together/lcpm.RData")
data0=lcpm[match(tumorcellgene,rownames(lcpm),nomatch = 0),]

library(ComplexHeatmap)
annot_df_col=sampleinfo0[match(colnames(data0),sampleinfo0$sample),c("LiveDead","group")]
annot_df_col$group=droplevels(annot_df_col$group)
column_color1=rainbow(length(levels(annot_df_col$group)))
names(column_color1)=levels(annot_df_col$group)

annot_df_row=genedf[match(rownames(data0),genedf$gene),"type",drop=F]
# column_color2=rainbow(length(unique(annot_df_col$type2)))
# names(column_color2)=unique(annot_df_col$type2)
# column_color3=rainbow(length(unique(annot_df_col$mice)))
# names(column_color3)=unique(annot_df_col$mice)

anno_col=HeatmapAnnotation(df=annot_df_col,col=list(group=column_color1))
anno_row=rowAnnotation(df=annot_df_row)

t=data0
t=t(scale(t(t)))
# t=scale(t)
heatmap_matrix=Heatmap(as.matrix(t), #name = "scaledata", 
                       cluster_rows = T, 
                       cluster_columns = T, 
                       top_annotation = anno_col,# use_raster=T,
                       column_split = annot_df_col$group ,
                       left_annotation = anno_row,
                       row_split = annot_df_row$type,
                       # show_column_dend = F, show_row_dend = F, 
                       clustering_distance_rows='euclidean', show_row_names = T,show_column_names = T,
                       # col=hmcols,
                       #column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                       row_km = 1)
pdf(paste0(dataname0,"gene.activity.tumorgene.",".pdf"),width = 15)
print(heatmap_matrix)

dev.off()


```

```{r}
##20250403


go=c("regulation of lipid metabolic process","negative regulation of leukocyte activation","purine nucleoside monophosphate metabolic process")
kegg=c("Pathways in cancer","Protein processing in endoplasmic reticulum","IL-17 signaling pathway","Apoptosis","JAK-STAT signaling pathway","T cell receptor signaling pathway","PD-L1 expression and PD-1 checkpoint pathway in cancer","HIF-1 signaling pathway","Ferroptosis")

load("E:/databases/pathway_genesets/02_kegg_API_download/kegg_api_geneSets20221019.RData")
gs1=hsa_kegg_list_des[names(hsa_kegg_list_des)%in%kegg]

load("E:/databases/pathway_genesets/01_msigdb/20220428_msigdblist_human&mice.RData")#msigdb
showpath=paste0("GO_",toupper(gsub(" ","_",go)))
showpath=showpath[!is.na(match(showpath,names(msigdblist$c5_06_GO_BP)))]
gs_GO=msigdblist$c5_06_GO_BP[showpath]

gs=c(gs1,gs_GO)


for(cellline in c("cellwithneutrophil","supernatantwithneutrophil")){
  print(cellline)
  
  data0=savelist[[cellline]]$data0
  data0=data0[,grep("Hu2_|Hu1_",colnames(data0),invert = T)]#
  
  enrich.mat <- GSVA::gsva(expr = as.matrix(data0), gset.idx.list = gs, kcdf = "Poisson",method = "gsva", verbose = TRUE)
  enrich.mat0=enrich.mat
  library(ComplexHeatmap)
  annot_df_col=sampleinfo0[match(colnames(data0),sampleinfo0$sample),c("LiveDead","group")]
  annot_df_col$group=droplevels(annot_df_col$group)
  column_color1=rainbow(length(levels(annot_df_col$group)))
  names(column_color1)=levels(annot_df_col$group)
  
  # annot_df_row=pathwaydf[match(rownames(enrich.mat0),pathwaydf$pathway),"type",drop=F]
  # column_color2=rainbow(length(unique(annot_df_col$type2)))
  # names(column_color2)=unique(annot_df_col$type2)
  # column_color3=rainbow(length(unique(annot_df_col$mice)))
  # names(column_color3)=unique(annot_df_col$mice)
  
  anno_col=HeatmapAnnotation(df=annot_df_col,col=list(group=column_color1))
  # anno_row=rowAnnotation(df=annot_df_row)
  
  t=enrich.mat0
  t=t(scale(t(t)))
  # t=scale(t)
  heatmap_matrix=Heatmap(as.matrix(t), #name = "scaledata", 
                         cluster_rows = T, 
                         cluster_columns = T, 
                         top_annotation = anno_col,# use_raster=T,
                         column_split = annot_df_col$group ,
                         # left_annotation = anno_row,
                         # row_split = annot_df_row$type,
                         # show_column_dend = F, show_row_dend = F, 
                         clustering_distance_rows='euclidean', show_row_names = T,show_column_names = T,
                         # col=hmcols,
                         #column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                         row_km = 1)
  pdf(paste0(dataname0,"pathway.activity2.",cellline,".pdf"),width = 10,height = 10)
  print(heatmap_matrix)
  dev.off()
  
}

genes=c("BDH1","IL10RB","TNFRSF8","CYBB","CXCL2","TRAF1","LAMB3","CSF1","HIVEP1","TREM1","CXCL8","IL1B","MVD","CD274","IL4R","IL17RA","CXCL1","CISH")
load("savelist_nofilter.RData")
for(cellline in c("cellwithneutrophil","supernatantwithneutrophil")){
  print(cellline)
  
  # data0=savelist[[cellline]]$lcpm
  data0=savelist_nofilter[[cellline]]$lcpm
  data0=data0[,grep("Hu2_|Hu1_",colnames(data0),invert = T)]#
  data0=data0[match(genes,rownames(data0),nomatch = 0),]
  # data0=data0[match(genedf$gene,rownames(data0),nomatch = 0),]
  
  library(ComplexHeatmap)
  annot_df_col=sampleinfo0[match(colnames(data0),sampleinfo0$sample),c("LiveDead","group")]
  annot_df_col$group=droplevels(annot_df_col$group)
  column_color1=rainbow(length(levels(annot_df_col$group)))
  names(column_color1)=levels(annot_df_col$group)
  
  # annot_df_row=genedf[match(rownames(data0),genedf$gene),"type",drop=F]
  # column_color2=rainbow(length(unique(annot_df_col$type2)))
  # names(column_color2)=unique(annot_df_col$type2)
  # column_color3=rainbow(length(unique(annot_df_col$mice)))
  # names(column_color3)=unique(annot_df_col$mice)
  
  anno_col=HeatmapAnnotation(df=annot_df_col,col=list(group=column_color1))
  # anno_row=rowAnnotation(df=annot_df_row)
  
  t=data0
  t=t(scale(t(t)))
  # t=scale(t)
  heatmap_matrix=Heatmap(as.matrix(t), #name = "scaledata", 
                         cluster_rows = T, 
                         cluster_columns = T, 
                         top_annotation = anno_col,# use_raster=T,
                         column_split = annot_df_col$group ,
                         # left_annotation = anno_row,
                         # row_split = annot_df_row$type,
                         # show_column_dend = F, show_row_dend = F, 
                         clustering_distance_rows='euclidean', show_row_names = T,show_column_names = T,
                         # col=hmcols,
                         #column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                         row_km = 1)
  pdf(paste0(dataname0,"gene.activity.nofilter2.",cellline,".pdf"),width = 10,height = 10)
  print(heatmap_matrix)
  
  dev.off()
  
}



```

##20250409
```{r}

go=c("PERK-mediated unfolded protein response","regulation of response to endoplasmic reticulum stress","regulation of lipid metabolic process","negative regulation of leukocyte activation","purine nucleoside monophosphate metabolic process")
kegg=c("Pathways in cancer","Protein processing in endoplasmic reticulum","IL-17 signaling pathway","Apoptosis","JAK-STAT signaling pathway","PD-L1 expression and PD-1 checkpoint pathway in cancer","HIF-1 signaling pathway","Ferroptosis")

load("E:/databases/pathway_genesets/02_kegg_API_download/kegg_api_geneSets20221019.RData")
gs1=hsa_kegg_list_des[names(hsa_kegg_list_des)%in%kegg]

load("E:/databases/pathway_genesets/01_msigdb/20220428_msigdblist_human&mice.RData")#msigdb
showpath=paste0("GO_",toupper(gsub(" ","_",go)))
showpath=gsub("-","_",showpath)
showpath=showpath[!is.na(match(showpath,names(msigdblist$c5_06_GO_BP)))]
# showpath=showpath[!is.na(match(showpath,names(msigdblist$c5.ontology)))]
gs_GO=msigdblist$c5_06_GO_BP[showpath]

gs=c(gs1,gs_GO)


for(cellline in c("cellwithneutrophil","supernatantwithneutrophil")){
  print(cellline)
  
  data0=savelist[[cellline]]$data0
  data0=data0[,grep("Hu2_",colnames(data0),invert = T)]#|Hu1_|Hu10_
  # data0=data0[,grep("Hu2_|Hu1_",colnames(data0),invert = T)]#
  
  enrich.mat <- GSVA::gsva(expr = as.matrix(data0), gset.idx.list = gs, kcdf = "Poisson",method = "gsva", verbose = TRUE)
  enrich.mat0=enrich.mat
  library(ComplexHeatmap)
  annot_df_col=sampleinfo0[match(colnames(data0),sampleinfo0$sample),c("LiveDead","group")]
  annot_df_col$group=droplevels(annot_df_col$group)
  column_color1=rainbow(length(levels(annot_df_col$group)))
  names(column_color1)=levels(annot_df_col$group)
  
  # annot_df_row=pathwaydf[match(rownames(enrich.mat0),pathwaydf$pathway),"type",drop=F]
  # column_color2=rainbow(length(unique(annot_df_col$type2)))
  # names(column_color2)=unique(annot_df_col$type2)
  # column_color3=rainbow(length(unique(annot_df_col$mice)))
  # names(column_color3)=unique(annot_df_col$mice)
  
  anno_col=HeatmapAnnotation(df=annot_df_col,col=list(group=column_color1))
  # anno_row=rowAnnotation(df=annot_df_row)
  
  t=enrich.mat0
  t=t(scale(t(t)))
  # t=scale(t)
  heatmap_matrix=Heatmap(as.matrix(t), #name = "scaledata", 
                         cluster_rows = T, 
                         cluster_columns = T, 
                         top_annotation = anno_col,# use_raster=T,
                         column_split = annot_df_col$group ,
                         # left_annotation = anno_row,
                         # row_split = annot_df_row$type,
                         # show_column_dend = F, show_row_dend = F, 
                         clustering_distance_rows='euclidean', show_row_names = T,show_column_names = T,
                         # col=hmcols,
                         #column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                         row_km = 1)
  pdf(paste0(dataname0,"pathway.activity4.",cellline,".pdf"),width = 10,height = 10)
  print(heatmap_matrix)
  dev.off()
  
}

genes=c("AQP9","BCL2A1","CCL18","CCL20","CD300E","CTSC","CXCL1","CXCL2","CXCL3","CXCL5","CXCL6","CXCR1","CXCR2","EMR2","EMR3","FGR","FPR1","FPR2","GPR97","HCK","HK3","ICAM1","IL15RA","IL1B","IL1R2","IL2RA","IL6","IL8","LILRA3","LILRA5","LILRA6","LIRRB2","LIRRB3","LYN","MEFV","MMP12","NCF2","OSM","PLAUR","PTAFR","S100A12","S100A8","S100A9","SLC11A1","TREM1","CD300ld","CD300Id","CD84","LOX-1","Arg1","Arg2","FATP2","IL6","NOX2","PTGS2","TNFA","MMP8","BDH1","IL10RB","TNFRSF8","CYBB","CXCL2","TRAF1","LAMB3","CSF1","HIVEP1","TREM1","CXCL8","IL1B","MVD","CD274","IL4R","IL17RA","CXCL1","CISH")
genes=unique(genes)
genes=setdiff(genes,c("IL6","IL15RA","CXCL6","CXCL3","CCL20","IL2RA","LILRA6","LYN","LILRA5","HK3"))
load("savelist_nofilter.RData")
for(cellline in c("cellwithneutrophil","supernatantwithneutrophil")){
  print(cellline)
  
  # data0=savelist[[cellline]]$lcpm
  data0=savelist_nofilter[[cellline]]$lcpm
  data0=data0[,grep("Hu2_|Hu1_|Hu10_",colnames(data0),invert = T)]#
  data0=data0[match(genes,rownames(data0),nomatch = 0),]
  # data0=data0[match(genedf$gene,rownames(data0),nomatch = 0),]
  
  library(ComplexHeatmap)
  annot_df_col=sampleinfo0[match(colnames(data0),sampleinfo0$sample),c("LiveDead","group")]
  annot_df_col$group=droplevels(annot_df_col$group)
  column_color1=rainbow(length(levels(annot_df_col$group)))
  names(column_color1)=levels(annot_df_col$group)
  
  # annot_df_row=genedf[match(rownames(data0),genedf$gene),"type",drop=F]
  # column_color2=rainbow(length(unique(annot_df_col$type2)))
  # names(column_color2)=unique(annot_df_col$type2)
  # column_color3=rainbow(length(unique(annot_df_col$mice)))
  # names(column_color3)=unique(annot_df_col$mice)
  
  anno_col=HeatmapAnnotation(df=annot_df_col,col=list(group=column_color1))
  # anno_row=rowAnnotation(df=annot_df_row)
  
  t=data0
  t=t(scale(t(t)))
  t=t[!is.na(t[,1]),]
  # t=scale(t)
  heatmap_matrix=Heatmap(as.matrix(t), #name = "scaledata", 
                         cluster_rows = T, 
                         cluster_columns = T, 
                         top_annotation = anno_col,# use_raster=T,
                         column_split = annot_df_col$group ,
                         # left_annotation = anno_row,
                         # row_split = annot_df_row$type,
                         # show_column_dend = F, show_row_dend = F, 
                         clustering_distance_rows='euclidean', show_row_names = T,show_column_names = T,
                         # col=hmcols,
                         #column_dend_height = unit(40, "mm"),#row_names_gp = gpar(fontsize=10),
                         row_km = 1)
  pdf(paste0(dataname0,"gene.activity.nofilter.rm1.2.10",cellline,".pdf"),width = 10,height = 10)
  print(heatmap_matrix)
  
  dev.off()
  
}




##lancet



library(survminer)

lancet_2005=read.table("E:/course/paper/from20200601/survival_prognosis_computing_nature.NEJM/2005.lancet.286patient/exp.txt",
                       header = T,sep = "\t",quote = "")
clinical0=read.table("E:/course/paper/from20200601/survival_prognosis_computing_nature.NEJM/2005.lancet.286patient/clinical.txt",
                     header = T,sep = "\t",quote = "")
clinical0$time_to_relapse_or_last_follow.up_months=clinical0$time_to_relapse_or_last_follow.up_months/12

GPL96=read.table("E:/SHJ/20200417_array_bone_metastasis/series_matrix/GPL96-57554.txt",header = T,as.is = T,sep = "\t",quote = "")
GPL96=GPL96[GPL96$Gene.Symbol!="",]
GPL96=GPL96[!grepl("[/]",GPL96$Gene.Symbol),]
exp=aggregate(lancet_2005[,2:ncol(lancet_2005)],by=list(gene=GPL96$Gene.Symbol[match(lancet_2005$ID_REF,GPL96$ID)]),max)
rownames(exp)=exp$gene
exp=exp[,-1]
exp=log(exp)
exp=exp[,clinical0$GEO_asscession_number]


TF_target_list_to_hs=readRDS("E:/databases/TF_target/20210615_TF_target_list_human.rds")




library(survminer)

lancet_2005=read.table("E:/course/paper/from20200601/survival_prognosis_computing_nature.NEJM/2005.lancet.286patient/exp.txt",
                       header = T,sep = "\t",quote = "")
clinical0=read.table("E:/course/paper/from20200601/survival_prognosis_computing_nature.NEJM/2005.lancet.286patient/clinical.txt",
                     header = T,sep = "\t",quote = "")
GPL96=read.table("E:/SHJ/20200417_array_bone_metastasis/series_matrix/GPL96-57554.txt",header = T,as.is = T,sep = "\t",quote = "")
GPL96=GPL96[GPL96$Gene.Symbol!="",]
GPL96=GPL96[!grepl("[/]",GPL96$Gene.Symbol),]
exp=aggregate(lancet_2005[,2:ncol(lancet_2005)],by=list(gene=GPL96$Gene.Symbol[match(lancet_2005$ID_REF,GPL96$ID)]),max)
rownames(exp)=exp$gene
exp=exp[,-1]
exp=log(exp)
exp=exp[,clinical0$GEO_asscession_number]



pdf("lancet_2005_replase_surv.pdf")
for(gene0 in c(names(gs))){#,"CD200","PCSK9","SERPINB9","SLC6A1","BCL11B","VDR","SLC6A2",'TSC22D1'
  print(gene0)
  gene=gs[[gene0]]
  expdata=exp
  clinical=clinical0
  genetmp=gene[gene%in%rownames(expdata)]
  if(sum(genetmp%in%rownames(expdata))==0){next()}
  # genetmp=TF_target_list_to_hs$encode[[gene]][TF_target_list_to_hs$encode[[gene]]%in%rownames(expdata)]
  
  clinical$score=apply(expdata[genetmp,,drop=F],2,mean)
  # clinical=clinical[clinical$ER_Status=="ER+",]
  
  res.cut=surv_cutpoint(clinical,time = "time_to_relapse_or_last_follow.up_months",event="relapse_1_True",variables = "score")
  res.cat=surv_categorize(res.cut)
  fit=survfit(as.formula(paste0('Surv(time_to_relapse_or_last_follow.up_months, relapse_1_True=="1") ~',"score")), data = res.cat)
  p1=ggsurvplot(
    fit,                     # survfit object with calculated statistics.
    data = res.cat,             # data used to fit survival curves.
    risk.table = TRUE,       # show risk table.
    pval = TRUE,             # show p-value of log-rank test.
    # conf.int = TRUE,         # show confidence intervals for 
    # point estimates of survival curves.
    # xlim = x_limit,        # present narrower X axis, but not affect
    # survival estimates.
    xlab = "Time in years",   # customize X axis label.
    # break.time.by = break.time.by,     # break X axis in time intervals by 500.
    # ggtheme = theme_light(), # customize plot and risk table with a theme.
    # risk.table.y.text.col = T, # colour risk table text annotations.
    risk.table.y.text = FALSE # show bars instead of names in text annotations
    # in legend of risk table
  )+labs(title=paste(gene0,sep = "_"))
  
  print(p1)
  
  for(type in c("ER-","ER+")){
    t=clinical[clinical$ER_Status%in%type,]
    res.cut=surv_cutpoint(t,time = "time_to_relapse_or_last_follow.up_months",event="relapse_1_True",variables = "score")
    res.cat=surv_categorize(res.cut)
    fit=survfit(as.formula(paste0('Surv(time_to_relapse_or_last_follow.up_months, relapse_1_True=="1") ~',"score")), data = res.cat)
    p1=ggsurvplot(
      fit,                     # survfit object with calculated statistics.
      data = res.cat,             # data used to fit survival curves.
      risk.table = TRUE,       # show risk table.
      pval = TRUE,             # show p-value of log-rank test.
      # conf.int = TRUE,         # show confidence intervals for 
      # point estimates of survival curves.
      # xlim = x_limit,        # present narrower X axis, but not affect
      # survival estimates.
      xlab = "Time in years",   # customize X axis label.
      # break.time.by = break.time.by,     # break X axis in time intervals by 500.
      # ggtheme = theme_light(), # customize plot and risk table with a theme.
      # risk.table.y.text.col = T, # colour risk table text annotations.
      risk.table.y.text = FALSE # show bars instead of names in text annotations
      # in legend of risk table
    )+labs(title=paste(gene0,type,sep = "_"))
    
    print(p1)
  }
}
dev.off()


```


```{r}
# siggene_list=lapply(names(DR_list),function(x){a=updown[,x];b=rownames(updown)[a!=0];DR_list[[x]][b,]})
siggene_list=lapply(DR_list,function(x){x[x$P.Value<0.05,]})
names(siggene_list)=names(DR_list)
siggene_list=lapply(siggene_list,function(x){x$gene=rownames(x);return(x)})
writexl::write_xlsx(siggene_list,path = "siggene_list.xlsx")
up1=siggene_list$N12_vs_CTRL23[siggene_list$N12_vs_CTRL23$logFC>0,]
down1=siggene_list$N12_vs_CTRL23[siggene_list$N12_vs_CTRL23$logFC<0,]

# up1=siggene_list$PCDH_vs_CTRL23[siggene_list$PCDH_vs_CTRL23$logFC>0,]
# down1=siggene_list$PCDH_vs_CTRL23[siggene_list$PCDH_vs_CTRL23$logFC<0,]

up2=siggene_list$NFK_vs_PCDH[siggene_list$NFK_vs_PCDH$logFC>0,]
down2=siggene_list$NFK_vs_PCDH[siggene_list$NFK_vs_PCDH$logFC<0,]

rescued=list(up=intersect(rownames(up1),rownames(down2)),down=intersect(rownames(down1),rownames(up2)))
genes_to_show=do.call("c",rescued)

pdf("rescued_gene.pdf")
for(gene in c(genes_to_show)){
  df=data.frame(exp=lcpm[gene,],groups=group)
  p=ggplot(df,aes(group,exp,fill=groups))+geom_boxplot()+geom_point()+
    ggpubr::stat_compare_means(comparisons = as.list(data.frame(combn(levels(df$groups),2))),method = "t.test")+
    theme_classic()+labs(title = gene)
  print(p)
}
dev.off()

d=lcpm[genes_to_show,]
d=t(scale(t(d)))


annot_df_col=data.frame(celltype=sampleinfo$type)
annot_df_col$celltype=factor(annot_df_col$celltype,levels = c(levels(group)))
# column_col=colpalettes[1:3]
annol_col=HeatmapAnnotation(df=annot_df_col)
pdf("rescued.gene.pdf")
Heatmap(d, name = "scaledata", cluster_rows = T, cluster_columns = F, use_raster=T,
        column_split = annot_df_col$celltype,
        top_annotation = annol_col,
        show_column_dend = T, show_row_dend = F,
        clustering_distance_rows='euclidean', show_row_names = T,show_column_names = F,
        # col =col_fun,
        column_dend_height = unit(40, "mm"),
        row_km = 1)
dev.off()


```


```{r}
library(gplots)
g=unique(do.call("c",lapply(DR_list[1:7],function(x){head(rownames(x),10)})))
mycol <- colorpanel(1000,"blue","white","red")
pdf(paste0(dataname0,"heatmap",".pdf"))
heatmap.2(lcpm[g,], scale="row",
          # labRow=v$genes$SYMBOL[g],
          labCol=group,
          col=mycol, trace="none", density.info="none", 
          # margin=c(8,6), lhei=c(2,10),
          dendrogram="column")


library(ComplexHeatmap)
annot_df_col=x$samples
column_color2=rainbow(length(unique(annot_df_col$group)))
names(column_color2)=levels(annot_df_col$group)
anno_col=HeatmapAnnotation(df=annot_df_col,col=list(group=column_color2))

t=t(scale(t(lcpm[g,])))
heatmapresult=Heatmap(t, name = "scaledata", 
                      cluster_rows = T, 
                      cluster_columns = T, 
                      top_annotation = anno_col, use_raster=T,
                      # column_split = annot_df_col$type,
                      # row_split= max_ind,
                      # left_annotation = anno_row,
                      show_column_dend = F, show_row_dend = F, 
                      clustering_distance_rows='euclidean', show_row_names = T,show_column_names = T,
                      # col=hmcols,
                      #column_dend_height = unit(40, "mm"),#
                      row_names_gp = gpar(fontsize=3),
                      row_km = 1)
heatmapresult
dev.off()
```

```{r}
load("E:/databases/pathway_genesets/Msigdb/20220414_Msigdb.RData")
idx=ids2indices(Mm.Msigdb_sym$c2,rownames(v))
cam_list=list()
for(i in 1){
  
  cam=camera(v,idx,design = design,contrast = contr.matrix[,i])
  cam=cbind(data.frame(path=rownames(cam)),cam)
  cam_list[[colnames(contr.matrix)[i]]]=cam
  # write.table(cam,file =paste0(dataname0,"pathway",colnames(contr.matrix)[i],".txt"),sep = "\t",col.names = T,row.names = T,quote = F)
}
writexl::write_xlsx(cam_list,path = '20230814_pathway_gsea.xlsx')
save(cam_list,file = "cam_list.RData")
# load("cam_list.RData")
cam_list=lapply(cam_list,function(x){x$path=tolower(x$path);return(x)})
writexl::write_xlsx(cam_list,path = '20230814_pathway_gsea.xlsx')

barcodeplot(efit$t[,1], index=idx$KEGG_GLYCOLYSIS_GLUCONEOGENESIS, 
            main=colnames(contr.matrix)[1])


```

```{r}
geneset=do.call("c",Mm.Msigdb_sym)
pathwaylist=data.frame(pathway=rep(names(geneset),sapply(geneset,length)),gene=do.call("c",geneset),stringsAsFactors = F)
gsea_list=list()

load("DR_list.RData")
for(i in names(DR_list)){
  print(i)
  gene=DR_list[[i]]
  
  geneList=gene$logFC
  names(geneList)=rownames(gene)
  geneList=sort(geneList,decreasing = T)
  gsea= GSEA(geneList, TERM2GENE = pathwaylist)
  gsea_list[[i]]=gsea
}

save(gsea_list,file = "gsea_list.RData")

t=lapply(gsea_list,function(x){x@result})
t=lapply(t,function(x){x$Description=tolower(x$Description);return(x)})
writexl::write_xlsx(t,path = "gsea.xlsx")

```

#enrich kegg go
```{r}
load("DR_list.RData")
markerlist=lapply(DR_list,function(x){x[x$P.Value<0.05,]})
enrichfunciton=list()
for(i in names(markerlist)){
  print(i)
  
  for(j in c("high","low")){
    print(j)
    t=markerlist[[i]]
    genes=rownames(t)[t$logFC>0]
    if(j=="low"){
      genes=rownames(t)[t$logFC<0]
    }
    
    geneID=gene2id$ENTREZID[match(genes,gene2id$SYMBOL,nomatch=0)]
    ego=enrichGO(gene = geneID, OrgDb = org.Mm.eg.db, ont = "BP", pAdjustMethod = "BH", qvalueCutoff = 0.05, readable = TRUE)
    ekegg <-enrichKEGG(gene = geneID, organism = 'mmu', pAdjustMethod = "BH", qvalueCutoff = 0.05)
    
    if(!is.null(ego)){
      ego=ego@result[ego@result$pvalue<0.05,]
      enrichfunciton[[i]][[j]][["ego"]]=ego
    }
    if(!is.null(ekegg)){
      ekegg=ekegg@result[ekegg@result$pvalue<0.05,]
      ekegg$genesymbol=sapply(ekegg$geneID,function(x){a=unlist(strsplit(x,"/"));b=gene2id$SYMBOL[match(a,gene2id$ENTREZID,nomatch = 0)];return(paste0(b,collapse = "/"))})
      enrichfunciton[[i]][[j]][["ekegg"]]=ekegg
      
    }
  }
}
save(enrichfunciton,file = "enrichfunciton.RData")
t=do.call("c",lapply(enrichfunciton,function(x){do.call("c",x)}))
writexl::write_xlsx(t,path = "enrichfunction.xlsx")


```

```{r}
ah = AnnotationHub::AnnotationHub()
# sort(unique(ah$dataprovider))
# sort(unique(ah$species))
# head(unique(ah$species))
mdb=query(ah,c("KEGG","Mus musculus"))
mdb2=query(ah,c("GeneOntology"))
df=mcols(mdb)
df2=mcols(mdb2)
meta_graph_info_mmu=mdb[["AH60905"]]
pmgi_mmu_genes=mdb[["AH60910"]]
pmgi_mmu_GO=mdb[["AH60911"]]
go_bp_frame_v2=mdb2[["AH69109"]]
allkegg=unique(meta_graph_info_mmu$all.labelids[,"path.name"])
allGOBP=unique(go_bp_frame_v2$name)
load("F:/Edata/dongtingting/20220913_RNAseq/enrichfunciton.RData")
for(db in c("go","kegg")){
  print(db)
  for(type in c("high","low")){
    
    single=xlsx::read.xlsx("E:/YaoBingQing/20200709_singlecell/enriched_pathway/singlecell_enriched_pathway.xlsx",sheetName = paste0(db,"_",type))
    single=single[single$Cluster%in%"Staph",]
    
    for(comp in names(enrichfunciton)){
      crispr=enrichfunciton[[comp]][[type]][[paste0("e",db)]]
      a=intersect(single$Description,crispr$Description)
      # xlsx::write.xlsx(a,file = "F:/Edata/dongtingting/20220913_RNAseq/20221026_single_crispr_intersection.xlsx",sheetName =paste0(db,"_single_",type,"_",comp) ,append = T)
      
      p=1
      if(db=="kegg"){
        p=hyperTest(gene1 = single$Description,gene2 = crispr$Description,backgroundGenes = union(allkegg,c(single$Description,crispr$Description)))
        
      }else{
        p=hyperTest(gene1 = single$Description,gene2 = crispr$Description,backgroundGenes = union(allGOBP,c(single$Description,crispr$Description)))
        
      }
      VennDiagram::venn.diagram(
        x = list(single=single$Description,crispr=crispr$Description),
        # category.names = c(LETTERS[1:4]),
        filename = paste0(paste0(db,"_single_",type,"_",comp),".tiff"),
        output=TRUE,
        fill = c("brown","darkblue"),
        main=paste0(db,"_single_",type,"_",comp),
        main.cex = 0.8,
        sub=paste0("pvalue: ",round(p,digits = 3)),
        sub.cex = 0.7
      )
    }
    
    
  }
  
}


```

pdf("function.pdf")
for(i in c("CTRL1 VS CTRL","Nfkb2KO VS CTRL")){
  print(i)
  t=xlsx::read.xlsx("F:/Edata/dongtingting/20220913_RNAseq/20221122-selected signaling pathway analyze.xlsx",sheetName = i)
  t$P2=-log10(t$PValue)
  t$P2[t$Direction=="Down"]=-1*t$P2[t$Direction=="Down"]
  t=t[order(t$Direction,t$P2),]
  t$path=paste0(t$path," (",t$NGenes,")")
  
  t$path=factor(t$path,levels = t$path[!duplicated(t$path)])
  p1=ggplot(t,aes(P2,path,fill=NGenes))+geom_bar(stat = "identity")+theme_classic()+
    labs(x="log10Pvalue",y="",title = i)+
    scale_fill_continuous(high="red",low="blue")
  p2=ggplot(t,aes(P2,path))+geom_bar(stat = "identity")+theme_classic()+
    labs(x="log10Pvalue",y="",title = i)
  
  t$NGenes[t$Direction=="Down"]=-1*t$NGenes[t$Direction=="Down"]
  t=t[order(t$Direction,t$NGenes),]
  t$P2=abs(t$P2)
  t$path=factor(as.character(t$path),levels = as.character(t$path[!duplicated(t$path)]))
  p3=ggplot(t,aes(NGenes,path,fill=P2))+geom_bar(stat = "identity")+theme_classic()+
    labs(x="NGenes",y="",title = i)+
    scale_fill_continuous(high="red",low="blue")
  p4=ggplot(t,aes(NGenes,path))+geom_bar(stat = "identity")+theme_classic()+
    labs(x="NGenes",y="",title = i)
  print(p1)
  print(p2)
  print(p3)
  print(p4)
}
dev.off()




library(survminer)
library(survival)
load("E:/publicdata/TCGA_BRCA/metabric_TCGA_BRCA.exp.clinical.RData")
load("E:/databases/mm10_geneloc/20210422_human2mice.RData")#human2mice gene symbol conversion

load("~/human.gene2id.RData")
load("F:/Edata/dongtingting/20220913_RNAseq/DR_list.RData")
gs=list(NFKB2singlegene="NFKB2",N12vsCTRL=DR_list$N12vsCTRL)
load("F:/Edata/dongtingting/20220913_RNAseq/ctrl1/DR_list.RData")
gs$N12vsCTRL1=DR_list$N12vsCTR1L
gs0=gs
gs=lapply(gs0,function(x){if(class(x)=="data.frame"){x$SYMBOL[x$adj.P.Val<0.05]}else{return(x)}})#adj.P.Val,P.Value

gs=lapply(gs,toupper)
gs$intersection=intersect(gs$N12vsCTRL1,gs$N12vsCTRL)


surv_df=matrix(0,ncol = 6)
colnames(surv_df)=c("cancertype","pathway","eventype","pval","tenratehigh","tenratelow")

pdf("BRCA_NFKB2_survival.pdf")
for(geneset in names(gs)){
  gene=gs[[geneset]]
  print(geneset)
  for(db in names(survival_datalist$exp)){
    print(db)
    expdata=survival_datalist$exp[[db]]
    clinical=survival_datalist$clinical[[db]]
    genetmp=gene[gene%in%rownames(expdata)]
    clinical$score=apply(expdata[genetmp,,drop=F],2,mean)
    clinical=clinical[clinical$time>0,]
    
    res.cut=surv_cutpoint(clinical,time = "time",event="event",variables = "score")
    res.cat=surv_categorize(res.cut)
    fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score")), data = res.cat)
    p1=ggsurvplot(
      fit,                     # survfit object with calculated statistics.
      data = res.cat,             # data used to fit survival curves.
      risk.table = TRUE,       # show risk table.
      pval = TRUE,             # show p-value of log-rank test.
      xlab = "Time in years",   # customize X axis label.
      risk.table.y.text = FALSE # show bars instead of names in text annotations
    )+labs(title=paste(geneset,db,sep = "_"))
    
    print(p1)
    
    p=unicox(surv=res.cat)
    a=p1$data.survplot[p1$data.survplot$score=="high",]
    tenratehigh=a$surv[which.min(abs(a$time-10))]
    a=p1$data.survplot[p1$data.survplot$score=="low",]
    tenratelow=a$surv[which.min(abs(a$time-10))]
    # alleventp=c()
    surv_df=rbind(surv_df,c(db,geneset,"OS",p$pvalue,signif(tenratehigh,3),signif(tenratelow,3)))
    
    for(subtype in levels(clinical$PAM50)){
      t=clinical[clinical$PAM50%in%subtype,]
      res.cut=surv_cutpoint(t,time = "time",event="event",variables = "score")
      res.cat=surv_categorize(res.cut)
      fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score")), data = res.cat)
      p1=ggsurvplot(
        fit,                     # survfit object with calculated statistics.
        data = res.cat,             # data used to fit survival curves.
        risk.table = TRUE,       # show risk table.
        pval = TRUE,             # show p-value of log-rank test.
        xlab = "Time in years",   # customize X axis label.
        risk.table.y.text = FALSE # show bars instead of names in text annotations
      )+labs(title=paste(geneset,db,subtype,sep = "_"))
      
      print(p1)
      
      p=unicox(surv=res.cat)
      a=p1$data.survplot[p1$data.survplot$score=="high",]
      tenratehigh=a$surv[which.min(abs(a$time-10))]
      a=p1$data.survplot[p1$data.survplot$score=="low",]
      tenratelow=a$surv[which.min(abs(a$time-10))]
      # alleventp=c()
      surv_df=rbind(surv_df,c(paste0(db,"_",subtype),geneset,"OS",p$pvalue,signif(tenratehigh,3),signif(tenratelow,3)))
      
    }
  }
  
}
dev.off()

surv_df=surv_df[-1,]
surv_df=as.data.frame(surv_df)
surv_df$pval=as.numeric(surv_df$pval)
surv_df$tenratehigh=as.numeric(surv_df$tenratehigh)
surv_df$tenratelow=as.numeric(surv_df$tenratelow)
surv_df$highlow=surv_df$tenratehigh>surv_df$tenratelow

t=surv_df
t=t%>%filter(pval<0.05)
t$pathway=factor(t$pathway,levels = names(gs))
t$pathway=droplevels(t$pathway)
p=ggplot(t,aes(pathway,cancertype))+
  geom_point(aes(size=-log(pval),color=highlow))+
  # scale_color_gradient2(low = "blue",high = "red",midpoint = -log(0.05))+
  facet_wrap(~eventype,scales = "free")+
  theme_classic()+#labs(title=gsub("surv_df_|.RData","",i))+
  theme(axis.text.x = element_text(angle = 90,hjust = 1,vjust = 0.5))
pdf("BRCA_surv_dot.pdf")
print(p)
dev.off()



library(tidyr)
library(UpSetR)

crispr=read.table("E:/DongTingTing/20210923/gene_sig_2_sgRNA_filter_20220105.txt",header = T)

single=read.table("E:/YaoBingQing/20200709_singlecell/staph.gene.txt",header = T)
single=single[single$p_val<0.05,]

gs=list(schigh=rownames(single)[single$avg_log2FC>0],sclow=rownames(single)[single$avg_log2FC<0],
        pos=crispr$gene[crispr$log2FoldChange>0],neg=crispr$gene[crispr$log2FoldChange<0]
)
VennDiagram::venn.diagram(gs,filename = "tmp.jpg")
b=fromList(gs)
pdf("gene.intersection.pdf")
upset(b,sets=colnames(b),set_size.show = T)
dev.off()
t=do.call("c",gs)[!duplicated(do.call("c",gs))]
rownames(b)=t
b$entry=rownames(b)


data_with_intersection <- b %>%
  unite(col = "intersection", -c("entry"), sep = "")

data_with_intersection %>%
  group_by(intersection) %>%
  summarise(list = list(entry)) %>%
  mutate(list = setNames(list, intersection)) %>%
  pull(list)

for(i in names(table(data_with_intersection$intersection))){
  t=colnames(b)[1:4][unlist(strsplit(i,""))=="1"]
  print(t)
  xlsx::write.xlsx(data.frame(gene=data_with_intersection$entry[data_with_intersection$intersection==i]),file = "scRNA_intersect_crispr.xlsx",sheetName = paste(t,collapse = "_"),append = T)
}

genesummary_show=crispr

pdf("crispr.single.intersection.pdf")
for(i in names(table(data_with_intersection$intersection))){
  print(i)
  t=colnames(b)[1:4][unlist(strsplit(i,""))=="1"]
  if(length(t)==1){next()}
  print(t)
  gene=data_with_intersection$entry[data_with_intersection$intersection==i]
  p1=ggplot(genesummary_show,aes(x=log2FoldChange,y=-log10(pvalue)))+geom_point()+
    geom_point(data=genesummary_show[genesummary_show$gene%in%gene,],col="red")+
    theme_classic()+labs(title = paste(t,collapse = "_"))
  p2=ggplot(genesummary_show,aes(x=log2FoldChange,y=-log10(pvalue)))+geom_point(col="grey")+
    geom_point(data=genesummary_show[genesummary_show$gene%in%gene,],col="red")+
    geom_text(data=genesummary_show[genesummary_show$gene%in%c(gene),],aes(label=gene))+
    theme_classic()+labs(title = paste(t,collapse = "_"))
  print(p1)
  print(p2)
}
dev.off()


load("E:/databases/pathway_genesets/02_kegg_API_download/kegg_api_geneSets20221019.RData")
virus_list=mmu_kegg_list_des[grep("virus",names(mmu_kegg_list_des),ignore.case = T)]

gene=unique(do.call("c",virus_list))
d=genesummary_show[genesummary_show$gene%in%gene,]
p1=ggplot(d,aes(x=log2FoldChange,y=-log10(pvalue)))+geom_point(col="grey")+
  geom_point()+
  geom_text(aes(label=gene))+
  theme_classic()+labs(title = paste(t,collapse = "_"))
p2=ggplot(d,aes(x=log2FoldChange,y=-log10(pvalue)))+geom_point(col="grey")+
  geom_point()+
  # geom_text(aes(label=gene))+
  theme_classic()+labs(title = paste(t,collapse = "_"))
pdf("virus_gene_intersect_crispr.pdf")
p1
p2
dev.off()


crispr=read.table("E:/DongTingTing/20210923/gene_sig_2_sgRNA_filter_20220105.txt",header = T)
genesummary_show=crispr
crisprpathway=xlsx::read.xlsx("E:/DongTingTing/20210923/20220105_gene_enrich_function_2.xlsx",sheetName = "neg_kegg")
crisprpathway=crisprpathway[crisprpathway$Count>2,]
load("E:/databases/pathway_genesets/02_kegg_API_download/kegg_api_geneSets20221019.RData")
virus_list=mmu_kegg_list_des[grep("virus",names(mmu_kegg_list_des),ignore.case = T)]
virus_list=virus_list[names(virus_list)%in%crisprpathway$Description]
gene=unique(do.call("c",virus_list))
crisprgene=genesummary_show$gene[genesummary_show$gene%in%gene]
crisprgeneneg=genesummary_show$gene[genesummary_show$gene%in%gene&genesummary_show$log2FoldChange<0]
crisprgenepos=genesummary_show$gene[genesummary_show$gene%in%gene&genesummary_show$log2FoldChange>0]

genelist=c(virus_list,list(crisprgene=crisprgene,crisprgeneneg=crisprgeneneg,crisprgenepos=crisprgenepos))
genelist=lapply(genelist,toupper)


library(survminer)
library(survival)
load("E:/publicdata/TCGA_BRCA/metabric_TCGA_BRCA.exp.clinical.RData")
load("E:/databases/mm10_geneloc/20210422_human2mice.RData")#human2mice gene symbol conversion

load("~/human.gene2id.RData")

surv_df=matrix(0,ncol = 6)
colnames(surv_df)=c("cancertype","pathway","eventype","pval","tenratehigh","tenratelow")

pdf("virus_survival.pdf")
for(geneset in names(genelist)){
  gene=genelist[[geneset]]
  print(geneset)
  for(db in names(survival_datalist$exp)){
    print(db)
    expdata=survival_datalist$exp[[db]]
    clinical=survival_datalist$clinical[[db]]
    genetmp=gene[gene%in%rownames(expdata)]
    clinical$score=apply(expdata[genetmp,,drop=F],2,mean)
    clinical=clinical[clinical$time>0,]
    
    res.cut=surv_cutpoint(clinical,time = "time",event="event",variables = "score")
    res.cat=surv_categorize(res.cut)
    fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score")), data = res.cat)
    p1=ggsurvplot(
      fit,                     # survfit object with calculated statistics.
      data = res.cat,             # data used to fit survival curves.
      risk.table = TRUE,       # show risk table.
      pval = TRUE,             # show p-value of log-rank test.
      xlab = "Time in years",   # customize X axis label.
      risk.table.y.text = FALSE # show bars instead of names in text annotations
    )+labs(title=paste(geneset,db,sep = "_"))
    
    print(p1)
    
    p=as.numeric(gsub("p = ","",survminer::surv_pvalue(fit)$pval.txt))
    a=p1$data.survplot[p1$data.survplot$score=="high",]
    tenratehigh=a$surv[which.min(abs(a$time-10))]
    a=p1$data.survplot[p1$data.survplot$score=="low",]
    tenratelow=a$surv[which.min(abs(a$time-10))]
    # alleventp=c()
    surv_df=rbind(surv_df,c(db,geneset,"OS",p,signif(tenratehigh,3),signif(tenratelow,3)))
    
    for(subtype in levels(clinical$PAM50)){
      t=clinical[clinical$PAM50%in%subtype,]
      res.cut=surv_cutpoint(t,time = "time",event="event",variables = "score")
      res.cat=surv_categorize(res.cut)
      fit=survfit(as.formula(paste0('Surv(time, event=="1") ~',"score")), data = res.cat)
      p1=ggsurvplot(
        fit,                     # survfit object with calculated statistics.
        data = res.cat,             # data used to fit survival curves.
        risk.table = TRUE,       # show risk table.
        pval = TRUE,             # show p-value of log-rank test.
        xlab = "Time in years",   # customize X axis label.
        risk.table.y.text = FALSE # show bars instead of names in text annotations
      )+labs(title=paste(geneset,db,subtype,sep = "_"))
      
      print(p1)
      
      p=as.numeric(gsub("p = ","",survminer::surv_pvalue(fit)$pval.txt))
      a=p1$data.survplot[p1$data.survplot$score=="high",]
      tenratehigh=a$surv[which.min(abs(a$time-10))]
      a=p1$data.survplot[p1$data.survplot$score=="low",]
      tenratelow=a$surv[which.min(abs(a$time-10))]
      # alleventp=c()
      surv_df=rbind(surv_df,c(paste0(db,"_",subtype),geneset,"OS",p,signif(tenratehigh,3),signif(tenratelow,3)))
      
    }
  }
  
}
dev.off()

